### 2014-08-30 paper/sections/syntax-and-semantics.tex
The syntax and op. semantics section

\section{The target language---syntax and semantics}
\label{sec:defs}
We let $x$, $y$, $z$ range over variables, $f$ over user-functions and
$p$ over  primitive functions ($\CONS$, $\PRIM$ etc.).   The syntax of
our language  is shown  in Figure~\ref{fig:lang-syntax}; it  has eager
semantics and  restricts programs to be in  Administrative Normal Form
(ANF)~\cite{chakravarty03perspective}  where all actual  parameters to
functions   are   variables.   This   restriction   does  not   affect
expressibility (and indeed we feel  free to ignore it in examples when
inessential). In addition, this  form also explicates the formation of
closures.  We further  require  that  each variable  in  a program  is
distinct, so  that no scope shadowing  occurs---this simplifies proofs
of soundness.  In this  formulation expressions: either perform a test
($\SIF$), make a computation step ($\LET$) or return ($\RETURN$).  The
$\RETURN$ keyword is logically redundant, but we find it clarifies the
semantics and analysis.

The body of the program is
the  expression denoted  by  $e_\mainpgm$; for analysis purposes
it is convenient
to regard $e_\mainpgm$ as part of a function definition
$(\DEFINE\ ({\tt main})\ e_\mainpgm)$ as in C\@.
% As  mentioned earlier,
We write  $\pi\!:\!e$ to  associate the label  $\pi$ (not part  of the
language syntax) with the program point just before expression $e$.

% \paragraph{Inlining.}
In spite of the ANF restrictions it is still possible
to inline non-recursive functions (a fact we use to prove the safety
of liveness analysis).  A user-function call
$(\LET\,\, x \leftarrow (f\,y_1\,\ldots\,y_n) \,\, \IN\,\, e)$
to a function defined
(after renaming its formals and locals to be disjoint from existing variables)
by $(\DEFINE\ (f\ z_1\ \ldots\ z_n)\ e_f)$
is replaced by a sequence of $\LET$'s of the form
$z_i\leftarrow (\ID\,\,y_i)$ followed by
the body $e_f$ but with its $(\RETURN\,\,w)$ expressions replaced by
$(\LET\,\, x \leftarrow (\ID\,\, w) \,\, \IN\,\, e)$.

\begin{figure}[t!]
\footnotesize
\begin{eqnarray*}
   p \in \mathit{Prog} & ::= & d_1 \ldots d_n \,\, e_\mainpgm
    \hspace{7em} \mbox{\em --- program}\\
    d \in Fdef & ::= & (\DEFINE\,\, (f\,\, x_1 \,\, \ldots \,\,x_n)\,\,
    e) 
    \hspace{1.7em} \mbox{\em --- function definition} \\
e \in \mathit{Expr} & ::= &
\left\{\begin{array}{ll@{\hspace{4em}}l}
       (\SIF\,\, x\,\, e_1\,\, e_2) && \mbox{\em --- conditional} \\ 
       (\LET\,\, x \leftarrow s\,\, \IN\,\, e) && \mbox{\em --- let binding} \\
       (\RETURN\,\, x) && \mbox{\em --- return from function}
    \end{array}\right. \\
s \in \mathit{Stmt} & ::= &
\left\{\begin{array}{lr@{\hspace{1em}}l}
       k && \mbox{\em --- constant (numeric or $\NIL$)}\\
       (\CONS\,\, x_1\,\, x_2) && \mbox{\em --- constructor} \\ 
       (\CAR\,\, x) &  (\CDR\,\, x) & \mbox{\em --- selectors} \\ 
       (\NULLQ\,\, x) & (\PRIM\,\, x_1\,\, x_2) & \mbox{\em ---  tester and generic arithmetic} \\ 
%        (\NULLQ\,\, x) && \mbox{\em ---  tester} \\ 
       (\ID\,\, x) && \mbox{\em ---  identity function (for inlining)} \\ 
%        (\PRIM\,\, x_1\,\, x_2) && \mbox{\em --- generic arithmetic} \\ 
       \multicolumn{2}{l}{(f\,\, x_1\,\,\ldots\,\, x_n)} 
            & \mbox{\em --- function application} 
    \end{array}\right.
\end{eqnarray*}
  \caption{The syntax of our language}\label{fig:lang-syntax}
\figrule
\normalsize
\end{figure}

\subsubsection{Semantics}
We now give an operational
semantics for  our language.  Later, a refinement of
the  operational semantics,  which we  call minefield  semantics, will
serve to prove liveness analysis correct.
We give a small-step semantics because, unlike big-step semantics,
correctness for non-terminating programs does not need special treatment.
We start with the domains used by the semantics:
\[
\begin{array}{rlcl@{\hspace{2em}}l}
\rho: & \mathit{Env} &=&\mathit{Var} \rightarrow {Loc} & \mbox{-- Environment} \\ 
v:   & \mathit{Val} &=& \mathbb{N} + \{\NIL\} + \mathit{Val \times
  Val}& \mbox{-- Values}\\
c:   & \mathit{Clo} &=& \mathit{(Exp \times Env)}& \mbox{-- Closures}\\
\heap: & \mathit{Heap} & =&\mathit{Loc} \rightarrow (\mathit{Val} +
\mathit{Clo}) & \mbox{-- Heap}
\end{array}
\]
%Note that in the eager case, if a cons had a numeric argument, this
%fact was known when the cons was created. Thus you could have unboxed
%numbers. This is not possible now. For instance, the argument of a
%cons could be a closure which evaluates to a number.
%Decision - numbers are boxed
Here $\mathit{Loc}$ is countable set of locations which hold $\CONS$ cells.
A value is either a number, the empty list $\NIL$, or a location  $\ell$.
Our liveness analysis does not track numeric values, and thus is neutral
as to whether these are boxed or represented as immediates.
An environment is a finite mapping from variables to values, and a heap a finite mapping
from locations to pairs of values.
% Given  a program,  an  environment is  a
% mapping from  the set  of variables of  the program  $\mathit{Var}$ to
% values in $\mathit{Val}$.    The heap is a mapping
% from  locations  to pairs of  values.
Finally, $\stk$ is a stack
(using $\bullet$ for push and $[\,]$ for empty stack)
of frames of unfinished function calls.  A
frame is a triple $(e,x,\rho)$ representing the call site
$
(\LET\,\, x \leftarrow (f\,y_1\,\ldots\,y_n) \,\, \IN\,\, e)
$
being evaluated in environment $\rho$.
Frames can also be viewed as continuations, in this view
the ({\sc ord-return}) 
 rule in the small-step operational semantics (Figure~\ref{fig:lang-opsem-small}) invokes them.

The semantics 
of statements $s$ are given by the judgement form
$\rho, \heap, s \rightsquigarrow \heap', v$
and those for expressions $e$ by the form
$\rho, \stk, \heap, e \rightarrow \rho', \stk', \heap', e'$.
The start state is $(\{\},[\,], \{\}, e_\mainpgm)$ and the
program terminates successfully with result value $\rho(x)$
on reaching the halt state $(\rho,[\,], \heap, (\RETURN\ x))$

Notation:  we write $\rho[x \mapsto v]$ for the environment which is
as $\rho$ but has value $v$ at $x$.
We also write $[\myvec{x} \mapsto \myvec{v}]$ which respectively
has values $v_1, \ldots, v_n$ at $x_1, \ldots, x_n$ and write
$[\myvec{x} \mapsto \rho(\myvec{y})]$ when 
$v_1, \ldots, v_n$ are $\rho(y_1), \ldots, \rho(y_n)$.

\paragraph{Stuck states.} Note that certain forms of $e$ do not reduce
with $\rightarrow$ (perhaps because $\rightsquigarrow$ could not reduce
a contained $s$).  Some of these we eliminate syntactically, e.g.\
ensuring all variables and functions are defined and are called with
the correct number of parameters.
Others include $(\CDR\ \NIL), (\CAR\ 3), (+\ \NIL\ 4)$ and $(\SIF\ \NIL\ e_1\ e_2)$.
All but the first can be eliminated with a static type system but,
treating our program as dynamically typed, we regard all these
as stuck states.

% A program terminates successfully if it reaches the halt state
% in Figure~\ref{fig:lang-opsem-small}.
% Note that $\rho,[\,], \heap, \pp{\pi}{(\RETURN\ x)}$
% is the halt state (with result value $\rho(x)$).

\begin{figure}[t!]  
% (Random text to ensure proper line-spacing (problem with infrule macro?).)

\footnotesize
\begin{minipage}{0.335\textwidth}
\infrule[ord-const]
        {}
        {\rho, \stk, \heap,   \pp{\pi}{k}
          \rightsquigarrow \rho, \stk, \heap, k}
\end{minipage}\begin{minipage}{0.655\textwidth}
\infrule[ord-cons]
      {
%%\mbox{$\ell \not\in \mbox{dom}(\heap)$ is a fresh location}
}
      {\rho, \ell \bullet \stk,\heap, \pp{\pi}{(\CONS\,\, x\ y)}
        \rightsquigarrow \rho, \stk, \heap[\ell\mapsto(\rho(x),\rho(y))],
        \ell}
\end{minipage} \\[0.5ex]

\begin{minipage}{0.395\textwidth}
\infrule[ord-car]
        {\heap(\rho(x)) = (e_1, e_2)}
        {\rho, \stk, \heap, \pp{\pi}{(\CAR\,\, x)}
          \rightsquigarrow \rho, \stk, \heap, e_1}
\end{minipage}
\begin{minipage}{0.395\textwidth}
\infrule[ord-car]
        {\heap(\rho(x)) = e}
        {\rho, \stk, \heap, \pp{\pi}{(\CAR\,\, x)}
          \rightsquigarrow \rho, (\CAR \bullet):\stk, \heap, v_1}
\end{minipage}\\[0.5ex]

\begin{minipage}{0.595\textwidth}
\infrule[ord-cdr]
        {\heap(\rho(x)) = (v_1, v_2)}
        {\rho, \heap, \pp{\pi}{(\CDR\,\, x)}
          \rightsquigarrow \heap, v_2}
\end{minipage} \\[0.5ex]

\begin{minipage}{0.395\textwidth}
\infrule[ord-id]
      {
}
      {\rho, \heap, \pp{\pi}{(\ID\,\, x)}
        \rightsquigarrow \heap, \rho(x)}
\end{minipage}\begin{minipage}{0.595\textwidth}
\infrule[ord-prim]
      {\rho(x) \in \mathbb{N} \andalso \rho(y)\in \mathbb{N}}
      {\rho, \heap, \pp{\pi}{(\PRIM\,\, x\ y)}
        \rightsquigarrow \heap, \rho(x) + \rho(y)}
\end{minipage} \\[0.5ex]

\begin{minipage}{0.395\textwidth}
% \infrule[ord-null-false]
\infrule[]
        {\rho(x)\not= \NIL
        }{\rho, \heap, \pp{\pi}{(\NULLQ\,\, x)}
          \rightsquigarrow \heap, 0}
\end{minipage}\begin{minipage}{0.595\textwidth}
% \infrule[ord-null-true]
\infrule[ord-null]
        {\rho(x)= \NIL
        }{\rho, \heap, \pp{\pi}{(\NULLQ\,\, x)}
          \rightsquigarrow \heap, 1}
\end{minipage} \\[0.5ex]

\begin{minipage}{0.395\textwidth}
% \infrule[ord-if-true]{
\infrule[]{
  \rho(x) \in \mathbb{N}\setminus\{0\}
}{
  \rho,\stk, \heap,
  \pp{\pi}{(\SIF\ x\ \pp{\pi_2}{e_1}\ \pp{\pi_3}{e_2})}
  \longrightarrow \rho,\stk, \heap, e_1}
\end{minipage}\begin{minipage}{0.595\textwidth}
% \infrule[ord-if-false]{
\infrule[ord-if]{
  \rho(x) = 0
}{
  \rho,\stk, \heap,
  \pp{\pi}{(\SIF\ x\ \pp{\pi_2}{e_1}\ \pp{\pi_3}{e_2})}
  \longrightarrow \rho,\stk, \heap, e_2}
\end{minipage} \\[0.5ex]

\begin{minipage}{0.990\textwidth}
\infrule[ord-let-nonfn]{
%  \rho,\stk, \heap, \pp{\pi_1}{s} \longrightarrow \rho,\stk, \heap', v_1 
  \rho, \heap, \pp{\pi_1}{s} \rightsquigarrow \heap', v 
  \andalso
  \mbox{$s$ is not
$\pp{\pi}{(f\,\, \pp{\pi_1}{y_1} \ldots \pp{\pi_n}{y_n})}$
       }
}{
  \rho,\stk, \heap,  \pp{\pi}{(\LET\,\, x
    \leftarrow  \pp{\pi_1}{s}\,\, \IN\,\,
    \pp{\pi_2}{e})}
  \longrightarrow
  \rho[x \mapsto v], \stk, \heap', \pp{\pi_2}{e}}
\end{minipage} \\[0.5ex]

\begin{minipage}{0.990\textwidth}
\infrule[ord-let-fncall]{
  \mbox{$s$ is 
$\pp{\pi}{(f\,\, \pp{\pi_1}{y_1} \ldots \pp{\pi_n}{y_n})}$
       }
\andalso
\mbox{$f$ defined as $(\DEFINE\ (f\ z_1\ \ldots\ z_n)\ \pp{\pi_f}{e_f})$}
}{
  \rho,\stk, \heap,  \pp{\pi}{(\LET\,\, x
    \leftarrow  \pp{\pi_1}{s}\,\, \IN\,\,
    \pp{\pi_2}{e})}
  \longrightarrow
% [z_1\mapsto \rho(y_1), \ldots, z_n\mapsto \rho(y_n)], 
[\myvec{z} \mapsto \rho(\myvec{y})],
  \,(\rho,x,e)\bullet\stk,\, \heap, e_f
}
\end{minipage} \\[0.5ex]

\begin{minipage}{0.990\textwidth}
\infrule[ord-return]{
}{
  \rho, \,(\rho',x',e')\bullet\stk,\, \heap,
  \pp{\pi}{(\RETURN\ x)}
  \longrightarrow \rho'[x' \mapsto \rho(x)], \stk, \heap, e'}
\end{minipage}

% \normalsize
% Note that $\rho,[\,], \heap, \pp{\pi}{(\RETURN\ x)}$
% is the halt state (with result value $\rho(x)$).

  \caption{The small-step operational semantics}
  \label{fig:lang-opsem-small}
\figrule
\normalsize
\end{figure}

%==============================================================
\renewcommand{\pp}[2]{\ensuremath{#1\!\!:\!#2}} % prog point