### 2014-09-10 paper/sections/liveness.tex
Just started

\section{Liveness}\label{sec:liveness}

In classical liveness analysis a  variable is either `live' (its value
may be  used in future  computation) or `dead' (definitely  not used).
Semantically, a variable is dead at a given program point if arbitrary
changes to its value have no effect on the computation.  Later we will
use $\bot$ to represent a value  which `explodes' when it is used in a
computation; dead variables can  safely have their value replaced with
$\bot$.  For heap-allocated data we need a richer model of liveness in
that both variables  and fields of $\CONS$ cells may  be dead or live.
Using  $\acar$,  $\acdr$  to  represent access  using  $\CAR$,  $\CDR$
respectively, liveness of the structure reachable from a variable is a
set  of  {\em  access  paths}  which  we  represent  as  a  subset  of
$\{\acar,\acdr\}^\ast$, and  use conventional grammar  notation.  Thus
the liveness of $x$ being $\{\acdr\acar, \acdr\acdr\acar\}$ means that
future computation can  only refer to the second  and third members of
$x$   considered  as   a   list.   Semantically,   access  paths   are
prefix-closed, as  accessing a field requires accessing  all the paths
from  the variable  to  the field,  and  hence the  above liveness  is
properly   written   $\{\epsilon,   \acdr,   \acdr\acar,   \acdr\acdr,
\acdr\acdr\acar\}$.  The classical notions  of a scalar variable being
live or dead correspond to $\{\epsilon\}$ and $\{\}$.

The  overall liveness  (also  written {\em  liveness environment}  for
emphasis) at a program point  is conceptually a mapping from variables
to subsets of $\{\acar,\acdr\}^\ast$, but we often abuse notation, for
example writing $\{x.\acar\acdr, x.\acdr, y.\epsilon\}$ instead of the
map           $[x          \mapsto\{\epsilon,\acar,\acar\acdr,\acdr\},
  y\mapsto\{\epsilon\},  z\mapsto\{\}]$.    Analogously  to  classical
liveness,  the liveness  at  program  point $\pi$  in  $\pi:e$ is  the
liveness just before executing $e$.

A complementary notion to liveness is {\em demand}.  The demand for
expression $e$ is again an access path---that subset of of $\{\acar,\acdr\}^\ast$ which the
context of $e$ may explore of $e$'s result.  So, for example given a
demand $\sigma$ and the expression $\pi: (\RETURN\ x)$, the liveness at
$\pi$ is exactly $x.\sigma$.
The classical analogy of this is in {\em strong liveness}, where
an assignment node $n: x:=y+z$ causes $y$ and $z$ to be live on entry
to $n$ if (and only if) $x$ is live at exit of $n$---the liveness
of $x$ at exit from $n$ becomes the demand on $y+z$.
Note that, for an operation like division which
may raise an exception, the assignment $n: x:=y/z$ makes $y$ and $z$ live
regardless of the liveness of $x$.

We use $\sigma$  to range over demands, $\alpha$  to range over access
paths  and $\Lv$  to range  over liveness  environments.  The notation
$\sigma_1\sigma_2$  denotes  the  set $\lbrace  \alpha_1\alpha_2  \mid
\alpha_1 \in  \sigma_1, \alpha_2 \in \sigma_2\rbrace$.  Often we shall
abuse notation to  juxtapose an edge label and a  set of access paths:
$\acar\sigma$   is   a   shorthand  for   $\lbrace\acar\rbrace\sigma$.
Finally,  we use $\Lfonly$  to range  over {\em  demand transformers}.
These transform demands on a function call into demands for its formal
parameters:    given    a    user    function    $f$,    defined    by
$(\DEFINE\  (f\ x_1\  \ldots\ x_n)\  \  e_f)$ and  called with  demand
$\sigma$, then $\Lf{f}{i}{\sigma}$ is the liveness of $x_i$ at $e_f$.

Note that liveness refers to  variables and fields, and not to $\CONS$
cells  (i.e.\  to  edges  in   the  memory  graph,  not  to  locations
themselves).   Hence liveness  of $\{x.\epsilon,x.\acar\}$  means that
future computation may refer to  the value $\ell$ of variable $x$, and
also to the $\CAR$ field of  location $\ell$.  In the absence of other
pointers to heap location $\ell$, we are certain that the $\CDR$ field
of  $\ell$  will  not  be   referenced  and  may  hence  be  corrupted
arbitrarily.   Note therefore,  that  while $\ell$  cannot be  garbage
collected, any location  $\ell'$ stored in the $\CDR$  field of $\ell$
would  be  garbage (again  provided  there  are  no other  aliases  to
$\ell$).

\subsection{Liveness Analysis}
First recall the classical formulation of liveness (as sets of simple variables)
on three-address instructions,
$
\mathit{live}_\mathit{in}(I) = \mathit{live}_\mathit{out}(I)
     \setminus \mathit{def\/}(I)
     \cup   \mathit{ref\/}(I)
$,
and then note that {\em strong liveness} needs, when $I$ is the instruction $z:=x+y$,
that $\mathit{ref}(I)$ be refined to
$\{x,y\}$ if $z \in\mathit{live}_\mathit{out}(I)$ and $\{\}$ otherwise.

Our liveness analysis formulated in Figure~\ref{fig:live-judge}
is analogous.
% has three analogous aspects.
% \begin{itemize}
% \item
Firstly, the function $\mathit{ref}$, when given a statement $s$,
returns the incremental liveness {\em generated} by $s$.
Because we generalize {\em strong} liveness, $\mathit{ref}$ needs a second parameter,
specifying the demand $\sigma$ on the result of $s$,
to determine which access paths of its free variables are made live.
The cases for $(\ID\ x)$ and $(\PRIM\ x\ y)$ exemplify this.
% The cases for
% $(\CAR\ x)$, $(\CAR\ x)$ and $(\CONS\ x\ y)$ show how demand for path $\alpha \in \sigma$
% is transformed into a deeper or shallower path in the demand for (or liveness of) $x$ or $y$.
% The   operator \PRIM\ dereferences its arguments  resulting in the liveness shown.
A demand  of  $\sigma$  on  (\CAR~$x$)  is  transformed  to  the  demand
$\acar\sigma$  on  $x$.   In  addition, \CAR\  always  dereferences  its
argument (even if its result is never used).
This generates the liveness  $\{x.\epsilon\} \cup  x.\acar\sigma$
(note $\sigma$ may be $\{\}$).
In the opposite  sense, the  demand  of $\acar\sigma$  on
(\CONS~$x$~$y$) is  transformed to the  demand $\sigma$ on  $x$.  Note
that \CONS\ does not, by itself, dereference its arguments.
Thirdly, for the case of a user-function call, a third parameter $\Lfonly$ to 
$\mathit{ref}$
expresses how the demand $\sigma$ on the result is transformed into demands on its parameters.
Constants generate no liveness.

% \item
The  function  $\mathcal{L}$ now  gives  the  (total)  liveness of  an
expression $e$.   The cases $\RETURN$ and  $\SIF$ are straightforward,
but note the liveness $x.\epsilon$  generated by the latter.  The case
$(\LET\   z\leftarrow   s\  \IN\   e')$   resembles  a   three-address
instruction:  the liveness  of $e$  is given  by taking  the liveness,
$\Lv$, of $e'$, killing any liveness of $z$ and adding any incremental
liveness from  $s$.  The main subtlety  is how the liveness  of $z$ in
$\Lv$  is converted  to a  demand  $\Lv(z)$ to  be placed  on $s$  via
$\mathit{ref}(s,\Lv(z),\Lfonly)$.

% \item
%%%%%%\marginpar{Say $\mathcal{L}$ is monotonic?}
Finally, the  judgement form $\mathit{Prog}  \len \Lfonly$ is  used to
determine  $\Lfonly$\@.   Analogously   to  classical  liveness  being
computed  as  a  solution  of  dataflow  equations,  we  require,  via
inference   rule  ({\sc  live-define}),   $\Lfonly$  to   satisfy  the
fixed-point property that:  when we assume $\Lfonly$ to  be the family
of demand transformers for the program then the calculated liveness of
each function  body $\mathcal{L}(e_f,\sigma,\Lfonly)$ agrees  with the
assumed $\Lfonly$.   As usual, there  are often multiple  solutions to
$\Lfonly$;  all are  safe (see  Section~\ref{sec:correctness})  but we
prefer  the  least  one  as  giving  the  least  liveness  subject  to
safety---and hence greatest amount of garbage collected.
% \end{itemize}

We make three observations: firstly the rule ({\sc live-define}) has a
least  solution  as  $\mathcal{L}(\cdot)$  is monotonic  in  $\sigma$;
secondly  that  ({\sc  live-define})   resembles  the  rule  for  type
inference of mutually recursive  function definitions, and thirdly the
asymmetry of  demand and liveness (compared to  post- and pre-liveness
classically) is due to the functional formulation here.

Section~\ref{sec:computing} shows how the demand transformers $\Lfonly$ for a program
(representing a fully context-sensitive analysis)
can be safely approximated, for each function, by a {\em procedure summary}
(unifying the contexts in the style of 0-CFA)\@.
The summary consists of a pair of a single demand and, for this demand,
the corresponding tuple of demands the function makes on its arguments.

%==============================================================
\begin{figure}[t]  
% \footnotesize
\begin{eqnarray*}
\mathit{ref\/}(\kappa,\sigma,\Lfonly)
          &=& \{\,\} \mbox{, for $\kappa$ a constant, including $\NIL$}\\
\mathit{ref\/}((\CONS~x~y),\sigma,\Lfonly)
          &=& \{x.\alpha \mid \acar\alpha \in \sigma\} \cup \{y.\alpha \mid \acdr\alpha \in \sigma\} \\
\mathit{ref\/}((\CAR~x),\sigma,\Lfonly)
          &=& \{x.\epsilon\} \cup \{x.\acar\alpha \mid \alpha \in \sigma\} \\
\mathit{ref\/}((\CDR~x),\sigma,\Lfonly)
          &=& \{x.\epsilon\} \cup \{x.\acdr\alpha \mid \alpha \in \sigma\} \\
\mathit{ref\/}((\ID~x),\sigma,\Lfonly)
          &=& \{x.\sigma\} \\
\mathit{ref\/}((\PRIM~x~y),\sigma,\Lfonly)
          &=& \{x.\epsilon,y.\epsilon\} \\
\mathit{ref\/}((\NULLQ~x),\sigma,\Lfonly)
          &=& \{x.\epsilon\} \\
\mathit{ref\/}((f~y_1~\cdots~y_n),\sigma,\Lfonly)
%          &=& \bigcup_{i=1}^n y_i.\Lf{f}{i}{\sigma}
          &=&  \begin{array}{@{}l}  % to discourage \displaystyle
               \bigcup_{i=1}^n y_i.\Lf{f}{i}{\sigma}
               \end{array}
%          &=& \bigcup \{y_i.\Lf{f}{i}{\sigma} \mid i=1,\ldots, n\}
\\[1ex]
\mathcal{L}((\RETURN~x),\sigma,\Lfonly) &=& x.\sigma \\
\mathcal{L}((\SIF~x~e_1~e_2),\sigma,\Lfonly) &=&
        \mathcal{L}(e_1,\sigma,\Lfonly) \cup
        \mathcal{L}(e_2,\sigma,\Lfonly) \cup
        \{x.\epsilon\} \\
\mathcal{L}((\LET~x \leftarrow  s~\IN~e),\sigma,\Lfonly) &=&
        \Lv
           \setminus \{x.*\}
           \cup \mathit{ref\/}(s,\Lv(x),\Lfonly)
\mbox{ where } \Lv = \mathcal{L}(e,\sigma,\Lfonly)
\end{eqnarray*}

\begin{minipage}{0.85\textwidth}
\infrule[live-define]
        {\mathcal{L}(e_f,\sigma,\Lfonly) = 
           \bigcup_{i=1}^n z_i.\Lf{f}{i}{\sigma}
              \mbox{ for each $f$ and $\sigma$}
        }
        { d_1 \ldots d_k \len \Lfonly
\\ \makebox[0mm]{where
     $(\DEFINE\ (f\ z_1\ \ldots\ z_n)\ \ e_f)$ is a member of $d_1 \ldots d_k$}}
\end{minipage}

\normalsize
  \caption{Liveness equations and judgement rule}\label{fig:live-judge}
\end{figure}