### 2015-10-26 paper/proof.tex
fixing the proof

\documentclass[9pt]{sigplanconf}
\usepackage{etex}
\usepackage{float}
\usepackage{mathrsfs}
\usepackage{booktabs}
\usepackage{boxedminipage}
\usepackage[T1]{fontenc}
\usepackage{epsfig}
\usepackage{multirow}
\usepackage{url}
\usepackage[normalem]{ulem}
\usepackage{schemepgm}
\usepackage{graphicx}
\usepackage{graphpap}
\usepackage{tabularx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{pstricks}
\usepackage{pst-text}
\usepackage{pst-node}
\usepackage{pst-tree}
\usepackage{pst-rel-points}
\usepackage{bcprules}
\usepackage{subfigure}
\usepackage[boxed]{algorithm2e}
\usepackage{makecell}
\usepackage{balance}

\input{fh_defs}
\def\drawplusplus#1#2#3{\hbox to 0pt{\hbox to #1{\hfill\vrule height #3 depth
      0pt width #2\hfill\vrule height #3 depth 0pt width #2\hfill
      }}\vbox to #3{\vfill\hrule height #2 depth 0pt width
      #1 \vfill}}
      %Poor man's typography
\def\concat{\mathrel{\drawplusplus {12pt}{0.4pt}{5pt}}}
      %It would be better to specify these in font-relative measures, but it
      %probably doesn't scale anyway.
%\newcommand{\mycomment}[1]{}
\definecolor{Myblue}{rgb}{.2,0,1}
\newcommand{\comment}[1]{{\color{Myblue}{#1}}}
%\newcommand{\cblue}[1]{{\color{blue}{#1}}}
%\newcommand{\blankout}[1]{}

%\newcommand{\deltacall}[3]{\delta_{#1}({#2},{#3})}
\newcommand{\scmin} {\mbox{\sf\em in}}
\newcommand{\scmnull}{\mbox{\sf\em null?}}
\newcommand{\scmpair}{\mbox{\sf\em pair?}}
\newcommand{\scmprim}{\ensuremath{\mathsf{+}}}
\newcommand{\sh}[1]{{\colorbox{gray!20}{\framebox{$#1$}}}}
\def\myvec{\mathaccent"017E } 
\newcommand{\stk}{\mbox{S}}   
\newcommand{\ID}{\mbox{$\mathbf{ id}$}}
\newcommand{\bang}{\mbox{\sc bang}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{lemma}[theorem]{Lemma}
\begin{document}
% --- Author Metadata here ---
%\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
%\CopyrightYear{2007} % Allows default copyright year (20XX) to be
%over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data
%(0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

\title{Liveness-Based Garbage Collection for Lazy Languages}

% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.
\cmt{{
\numberofauthors{3} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors
section.
%
\author{
% 1st. author
\alignauthor K. Prasanna Kumar \\
       \affaddr{IIT Bombay,}\\
       \affaddr{Mumbai 400076, India}
       \email{prasannak@cse.iitb.ac.in}
% 2nd. author
\alignauthor Amey Karkare\\
       \affaddr{IIT Kanpur}\\
       \affaddr{Kanpur 208016, India}
       \email{karkare@cse.iitk.ac.in}
% 3rd. author
\alignauthor Amitabha Sanyal \\
       \affaddr{IIT Bombay}\\
       \affaddr{Mumbai 400076, India}\\
       \email{as@cse.iitb.ac.in}
}
}}
\authorinfo{Double Blind Review}{}{}
\maketitle

\begin{abstract}
We consider  the problem of reducing  the memory required to  run lazy
first-order functional programs.  To do this, we  analyze programs for
liveness of  heap-allocated data.  The accompanying  garbage collector
is modified  to use the result  of the analysis to  preserve only live
data---a subset  of reachable data---during garbage  collection.  This
results  in an  increase  in the  garbage  reclaimed and  consequently
reduces the  memory footprint of  programs.  While this  technique has
already been shown to yield  benefits for eager first-order languages,
the lack of a statically determinable execution order and the presence
of closures at runtime pose additional challenges for lazy languages.
This requires changes  both in the earlier liveness  analysis and also
in the  garbage collection scheme  required to reclaim memory  that is
not live.

The implementation  of our  analysis annotates each  potential garbage
collection  point  in   the  program  with  a   set  of  deterministic
finite-state automata (DFA) describing the  liveness at that point. We
also implement a copying collector that  uses the DFA to preserve live
objects.   Our  experiments  confirm  that  liveness-based  collection
always increases  the garbage  reclaimed, consequently  decreasing the
number of collections.  In addition,  it reduces the execution time of
several programs.  Experiments  also reveal that the  best results are
obtained  for   a  mixed  garbage   collection  scheme  that   does  a
liveness-based    collection   for    fully    evaluated   data    and
reachability-based   collection  for   closures.   On our benchmark
programs, we obtain a decrease of up to 45\% in the number of
collections, and a decrease of up to 88\% in the number of dragged
cells.  Even though liveness based collector uses an expensive
traversal scheme, the execution time is competitive (0.2X - 2.1X) when
compared to a reachability based collector.
\end{abstract}

\category{D.3.4}{Programming Languages}{Processors}[Memory Management
  (Garbage Collection), Optimizations]
\category{F.3.2}{Logic and Meanings Of Programs}{Semantics of
  Programming Languages}[Program Analysis]

\terms{Algorithms, Languages, Theory}

\keywords{Heap Analysis, Liveness Analysis, Memory Management, Garbage
  Collection, Lazy Languages}

\section{Introduction}
\label{sec:intro}

Functional programs use dynamically allocated memory extensively.  The
allocation  is either  explicit (while using  constructors, for example) or
implicit (as while creating  closures).  Programs  written in  lazy
functional languages put additional  demands on memory as they require
closures to  be carried  from the  point of creation  to the  point of
evaluation.

While  the runtime  system  of most  functional  languages includes  a
garbage  collector  (GC)  to  efficiently  reclaim  memory,  empirical
studies on Scheme~\cite{karkare06effectiveness} and, more importantly,
on  Haskell~\cite{rojemo96lag}  programs  have shown  that  GCs  leave
uncollected a  large number of  memory objects that are  reachable but
not live (here {\em live} means  the object can potentially be used by
the program  at a  later stage).  This  results in  unnecessary memory
retention.

In this paper,  we propose the use of liveness  analysis of heap cells
for garbage collection in a lazy first-order functional language.  The
central notion in our analysis  is a generalization of liveness called
{\em  demand}---the  pattern  of  future  uses  of  the  value  of  an
expression.   The analysis  has two  parts.  First,  each function  is
summarized in a context-sensitive manner as a {\em demand transformer}
that transforms a {\em symbolic} demand  on its body to demands on its
arguments.  This summary  is then used to step  through function calls
during analysis.   Second, the concrete  demand on a function  body is
obtained through a 0-CFA-like conservative approximation that combines
the  demands on  all the  calls to  the function.   The result  of the
analysis  is an  annotation of  each program  point with  finite-state
automata capturing the liveness of  variables at that point. Depending
on the program point where  garbage collection has been triggered, the
GC consults a set of  automata to restrict reachability during marking.
This results in an increase  in the garbage reclaimed and consequently
in fewer garbage collections.

While this idea has been shown to be effective~\cite{asati14lgc} for a
first-order {\em  eager} language, a straightforward  extension of the
earlier  technique  is  not  possible  for  lazy  languages,  where  a
heap-allocated   object   may   be  an   unevaluated   expression   or
closure. Since  data is made  live by  evaluation of closures,  and in
lazy languages  the place in  the program where this  evaluation takes
place cannot  be statically determined, laziness  complicates liveness
analysis itself.   In addition, apart  from evaluated data,  we extend
liveness-based collection  to closures---closures  with no  future use
should also  be collected.  This  involves tracking closures  to their
points of  creation at  runtime from where  we obtain  their liveness.
This requires significant modification to the earlier GC scheme.

Experiments    with   a    single    generation   copying    collector
(Section~\ref{sec:experiments})   confirm  the   expected  performance
benefits.  Liveness-based collection results in an increase in garbage
reclaimed.  As  a consequence, there  is a  decrease in the  number of
collections (up to 45\%) and a decrease  (up to 88\%) in the number of
{\em  dragged}  cells  (cells  that  survive a  GC  after  their  last
use). Besides, there is a reduction  in the overall execution time for
some programs.  Our  experiments show that a mixed  strategy, where we
do liveness based GC up to the closure boundary and reachability based
GC for the arguments (free variables) of the closure, gives us similar
benefits as  a fully  liveness based  GC (where  the arguments  of the
closure  are  also  collected  using liveness)  with  lesser  overhead
(Section~\ref{sec:strategies}).

%------------------------------------------------------------%
\subsection{Motivating Example}
\label{sec:motiv}
\begin{figure}[t!]
  \psset{unit=1mm}
  \begin{pspicture}(90,52)(0,-52)
    %\psframe(90,48)(0,-48)
    \renewcommand{\arraystretch}{1}
    \begin{tabular}{@{}c@{}}
      %\psframebox
          {\sf
	\renewcommand{\arraystretch}{1}{
	  \begin{uprogram}
            \UFL\
            \UNL{0} (\DEFINE\ (\length\  \pl)
	    \UNL{1}  (\SIF~(\NULLQ \ \pl) $0$
            (\PRIM\ 1\ (\length (\CDR\  \pl)))))
            \UNL{0}
	    \UNL{0}  (\DEFINE\ (\append\  \lista\ \listb)
	    \UNL{1}  (\SIF~(\NULLQ \ \lista)
	    \listb
	    \UNL{2} \hspace*{-0.1cm}(\CONS\ (\CAR\  \lista) (\append\
            (\CDR\  \lista)
            \listb))))
            \UNL{0}
            \UNL{0} (\LET\ \px\
            $\leftarrow$\ (\CONS\ $5$
            (\CONS\ (\CONS\ $6$ \NIL) \NIL) \IN
	    \UNL{1} (\LET\ \py\   $\leftarrow$\  (\CONS\ $3$ \NIL) \IN
	    \UNL{2}
            (\LET\ \pz\  $\leftarrow$\  (\append\ \px\  \py)\ \IN\
            \UNL{3} (\SIF~(\NULLQ~(\CAR~\pz))~$0$~$\pi$:\
            (\length\ \pz))))))
	  \end{uprogram}
      }}
      \\ \\
      (a) Example program. \\  \\
%      \includegraphics[width=.45\textwidth]{motiv-example}
      \input{mem-graph} \\ \\
      \renewcommand{\arraystretch}{.9}
      \begin{tabular}[t]{p{0.9\columnwidth}}
        (b) Memory graph at $\pi$.  \scalebox{.7}{\TwoCellsAD{a1}{a2}}
        denotes a closure. Thick edges denote live links. Traversal
        stops at edges marked $\times$ during garbage collection for
        an ideal collector.
      \end{tabular}
    \end{tabular} 
  \end{pspicture}
%  \vspace*{-3ex}
  \caption{Example Program and its Memory Graph}\label{fig:mot-example}
\end{figure}

Figure~\ref{fig:mot-example} shows an example program and the state of
the heap at  the program point $\pi$, i.e. just  before the evaluation
of $(\length\  \pz)$.  The heap is  represented by a graph  in which a
node either  represents atomic  values ($\NIL$,  integers etc.),  or a
\CONS\ cell containing $\CAR$ and $\CDR$  fields, or a chunk of memory
representing  an  unevaluated  value   also  called  a  {\em  closure}
(represented  by  hatched  boxes).   Edges   in  the  graph  are  {\em
  references} and emanate from variables  or fields.  The figure shows
the  lists \px\  and  \pz\ partially  evaluated.   The evaluation  was
triggered    by   the    test   (\NULLQ~(\CAR~\pz))    in   the
\SIF\ expression.

The edges  shown by thick  arrows are those  which are live  at $\pi$.
Ideally, a cell should be  preserved during garbage collection only if
it is reachable from  the root set through a path  of live edges.  All
other  cells can  be reclaimed.   Thus if  a garbage  collection takes
place at $\pi$ with the heap shown in Figure~\ref{fig:mot-example}(b),
an ideal liveness-based  collector (LGC) will preserve  only the cells
referenced by  $\pz$, $(\CDR~ \pz)$  and (the cells  constituting) the
closure   referenced   by   $(\CDR~(\CDR~\pz))$.    In   contrast,   a
reachability-based collector (RGC) would have preserved all cells.

While an ideal garbage collector is  impractical, in this work we show
that  static analysis  of  heap data  can help  practical garbage
collectors in reclaiming more  garbage.  The specific contributions of
this paper are
\begin{itemize}
\item  We  formulate  a  liveness  analysis for  a  lazy  first  order
  functional language  and prove its correctness.  We show  how to
  compute  function  summaries  in  a context  independent  manner  to
  achieve  a  context  sensitive  analysis. We  further  describe  the
  transformation of the  liveness equations to grammar rules  in a way
  that the language  accepted by the resulting  grammar describes live
  heap locations.
\item We prove  the undecidability of the grammar  that results from
  the  analysis.   To overcome  this,  we  present an  algorithm  to
  transform grammars  to finite  state automata  (DFA) such  that the
  language accepted by the  DFA over-approximates the liveness (and
  hence a safe approximation).
\item We have implemented a garbage  collector that can use the DFA
  to  retain live  cells that  also include  closures. We  have done
  extensive  experiments with  this garbage  collector and  obtained
  empirical results  that show  the effectiveness  of liveness-based
  garbage collection.
\end{itemize}

\subsection{Organization of the paper}

Section~\ref{sec:defs}  introduces  the   syntax  of  the  programming
language  considered  in our  analysis  and  a small-step  operational
semantics  for it.   We present  our liveness  analysis for  this lazy
language  in Section~\ref{sec:liveness}.   Section~\ref{sec:computing}
describes  the  formulation of  the  liveness  as  a grammar  and  its
approximation  to   DFA.  It  also   contains  a  proof   showing  the
undecidability  of  the  precise  liveness in  our  formulation  which
necessitates  the approximation  to DFA.   Section~\ref{sec:GC-scheme}
discusses different ways of handling closures and their pros and cons.
We  report our  experimental results  in Section~\ref{sec:experiments}
along with some observations.  Section~\ref{sec:relatedwork} discusses
previous  work   related  to  garbage  collection   and  liveness  and
Section~\ref{sec:conclusion}   discusses   possible   extensions   and
concludes the paper.
\section{The target language---syntax and semantics}
\label{sec:defs}
Figure~\ref{fig:lang-syntax} describes the syntax  of our language. It
is a first order language with lazy semantics. Programs are restricted
to        be        in        Administrative        Normal        Form
(ANF)~\cite{chakravarty03perspective} where  all actual  parameters to
functions  are  variables.  While  this  restriction  does not  affect
expressibility,  this form  has  the benefit  of  making explicit  the
creation of  closures through  the $\LET$  construct.  $\LET$s  in our
language  are lazy;  in the  expression $\LET\,\,  x \leftarrow  s\,\,
\IN\,\, e$, $x$ may occur in  $s$. This enables creation of graph-like
structures in  a pure language. The  restriction of \LET\ to  a single
definition  is for  ease  of  exposition---generalization to  multiple
definitions does not add conceptual difficulties.  We further restrict
each variable in a program to  be distinct, so that no scope shadowing
occurs---this  simplifies  the  reasoning   about  the  program.   The
$\SRETURN$  and   \SIF\  expressions  and  strict   operators  trigger
evaluation of closures.

The body  of a function ${\mathit  f}$ is denoted  as $e_{\mathit f}$.
We   shall   extend   this   convention   to   the   main   expression
$e_\mainpgm$. We assume that each program has a distinguished function
\mainpgm\      with      the      definition     $(\DEFINE\      ({\tt
  \mainpgm})\  e_\mainpgm)$ and  the execution  of the  program starts
with the call to \mainpgm.  We write $\pi\!:\!e$ to associate the label
$\pi$ (not  part of the language  syntax) with the  program point just
before expression $e$.

\begin{figure}[t]\footnotesize
\renewcommand{\arraystretch}{1}
\begin{eqnarray*}
   p \in \mathit{Prog} & \!\!\!::=\!\!\! & d_1 \ldots d_n \,\, e_\mainpgm
    \hspace{5em} \mbox{\em --- program}\\
    \mathit{df} \in Fdef & \!\!\!::=\!\!\! & (\DEFINE\,\, (f\,\, x_1 \,\, \ldots
\,\,x_n)\,\,
    e)
    \hspace{0.2em} \ \ \ \ \mbox{\em --- function def} \\
e \in \mathit{Expr} & \!\!\!::=\!\!\! &
\left\{\begin{array}{@{}ll@{\hspace{1em}}l}
       (\SIF\,\, x\,\, e_1\,\, e_2) && \!\!\!\mbox{\em --- conditional} \\
       (\LET\,\, x \leftarrow s\,\, \IN\,\, e) &&\!\!\! \mbox{\em --- let
binding} \\
       (\SRETURN\,\, x) && \!\!\!\mbox{\em --- return from function}
    \end{array}\right. \\
s \in \mathit{App} & \!\!\!::=\!\!\!  &
\left\{\begin{array}{@{}l@{\hspace{1.2em}}l}
       k & \mbox{\em --- constant (numeric or $\NIL$)}\\
       (\CONS\,\, x_1\,\, x_2) & \mbox{\em --- constructor} \\
       (\CAR\,\, x) & \\
       (\CDR\,\, x) & \mbox{\em --- selectors} \\
       (\NULLQ\,\, x) & \\
       (\PRIM\,\, x_1\,\, x_2) & \mbox{\em ---  tester/generic arithmetic} \\
%%       (\ID\,\, x) & \mbox{\em ---  identity function (for inlining)} \\
       (f\,\, x_1\,\,\ldots\,\, x_n) & \mbox{\em --- function application}
    \end{array}\right.
\end{eqnarray*}
  \caption{The syntax of our language}\label{fig:lang-syntax}
\figrule
\normalsize
\end{figure}

\begin{figure*}[t!]
\comment{

\begin{center}\footnotesize
\renewcommand{\arraystretch}{1.5}
\begin{tabular}{|c|c|c|}
\hline
Premise & Transition & Rule name \\
\hline
\hline 
          & $\rho,\, (\rho', \ell, e)\!:\!S,\, H,\, \kappa
  \rightsquigarrow \rho',\, S,\, H[\ell := \kappa],\, e$    &  \sc{const}
\\
\hline
          & {$\rho, \,(\rho', \ell, e)\!:\!S,\, H,\, (\CONS~x~y)
\rightsquigarrow
$  $\rho',\, S,\, H[\ell := (\rho(x), \rho(y))],\, e$}     &  \sc{cons} \\
\hline

$H(\rho(x)) \mbox{ is } \langle s, \rho'\rangle$ & $\rho,\, S,\, H,\, (\CAR~x)
\rightsquigarrow
\rho',\, (\rho, x, (\CAR~x))\!:\!S, \,H,\, s$      &
\sc{car-clo}
\\
\hline

$H(\rho(x)) \mbox{ is } (\langle s, \rho'\rangle, d)$ & $\rho,\, S,\,  H,\,
(\CAR~x)  \rightsquigarrow \rho', \,(\rho, addr(\langle s, \rho,\rangle), (\CAR~x))\!:\!S,\, H,\, s$      &
\sc{car-1-clo} \\
\hline

$H(\rho(x)) \mbox{ is } (v, d)$ & $\rho,\, (\rho', \ell, e)\!:\!S,\, H,\,
(\CAR~x)  \rightsquigarrow \rho',\, S,\, H[\ell := v],\, e$      &
\sc{car-select} \\
\hline

$H(\rho(x)), H(\rho(y)) \in \mathbb{N}$
 & {$\rho,\, (\rho', \ell, e)\!:\!S,\, H,\, (+~x~y)  \rightsquigarrow$
$\rho', \,S,\, H[\ell := H(\rho'(x)) + H(\rho'(y))],\, e$}      &
\sc{prim-add} \\
\hline

$H(\rho(x)) \mbox{ is } \langle s, \rho'\rangle$ & $\rho,\, S,\, H,\, (+~x~y)
\rightsquigarrow
\rho',\, (\rho, \rho(x), (+~x~y))\!:\!S,\, H,\, s$      &
\sc{prim-1-clo} \\
\hline
$H(\rho(y)) \mbox{ is } \langle s, \rho'\rangle $ & $\rho,\, S,\, H,\, (+~x~y)
\rightsquigarrow
\rho', (\rho,\, \rho(y),\, (+~x~y))\!:\!S,\, H,\, s$      &
\sc{prim-2-clo} \\
\hline
{$\mathit{f}~\mbox{defined as}$
$~(\DEFINE~(f~\myvec{y})~e_{\mathit{f}})$}  & $\rho,\, S,\, H,\,
(f~\myvec{x})  \rightsquigarrow
[\myvec{y} \mapsto \rho(\myvec{x})],\, S,\, H,\, e_{\mathit{f}}$      &
\sc{funcall} \\
\hline
$\ell$ is a new location& {$\rho,\, S,\, H,\, (\LET~x\leftarrow s~\IN~e)
  \rightsquigarrow$
$\rho\oplus[x \mapsto \ell],\, S,\, H[\ell \mapsto \langle s,
    \lfloor\rho\rfloor_{FV(s)}  \oplus [x \mapsto
  \ell]\rangle],\, e$} &
\sc{let} \\
\hline
$H(\rho(x)) \ne 0$ & $\rho,\, S,\, H,\, (\SIF~x~e_1~e_2)   \rightsquigarrow
\rho,\, S,\, H,\,  e_1$ & \sc{if-true} \\
\hline
$H(\rho(x)) = 0$ & $\rho,\, S,\, H,\, (\SIF~x~e_1~e_2)   \rightsquigarrow
\rho,\, S, \,H, \, e_2$ & \sc{if-false} \\
\hline
$H(\rho(x)) = \langle s, \rho' \rangle $ & {$\rho,\, S,\, H,\,
  (\SIF~x~e_1~e_2)   \rightsquigarrow
\rho',\, (\rho, \rho(x), (\SIF~x~e_1~e_2))\!:\!S,\, H, \, s$}
&
\sc{if-clo} \\
\hline
{$H(\rho(x))~\mbox{is}$ $\mbox{WHNF with value}~v$}& $\rho,\, (\rho', \ell,
e)\!:\!S,\, H,\,
(\SRETURN~x)  \rightsquigarrow \rho',\, S,\, H[\ell := v],\, e$ &
\sc{return-whnf}\\
\hline
$H(\rho(x)) = \langle s, \rho' \rangle $ & {$\rho,\, S,\, H,\, (\SRETURN~x)
  \rightsquigarrow$
$\rho',\, (\rho, \rho(x), (\SRETURN~x))\!:\!S,\, H,\,  s$} &
\sc{return-clo} \\
\hline
\end{tabular}
\caption{The small-step semantics for the language. \label{fig:lang-semantics}}
\end{center}
}
\end{figure*}

\subsection{Semantics}
\comment{
We now give  a small-step semantics for our  language.
We first specify the domains used by the semantics:
\[
\renewcommand{\arraystretch}{1}
\begin{array}{@{}r@{\ }l@{\ \ }c@{\ \ }l@{\hspace{0.5em}}l}
\heap: & \mathit{Heap} & =&\mathit{Loc} \rightarrow Data & \mbox{-- Heap}\\
d: & \mathit{Data} &=&\mathit{Val} + \mathit{Clo} & \mbox{-- Values \& Closures} \\
v:   & \mathit{Val} &=& \mathbb{N} + \{\NIL\} + \mathit{Data \times Data}& \mbox{-- Values}\\
c:   & \mathit{Clo} &=& \mathit{(App \times Env)}& \mbox{-- Closures}\\
\rho: & \mathit{Env} &=&\mathit{Var} \rightarrow \mathit{Loc} &
\mbox{-- Environment} \\
\end{array}
\]

Here  $\mathit{Loc}$ is  a countable  set  of locations  in the  heap.
These  either contain  a value in  \emph{WHNF} (a  number, the
empty  list  $\NIL$,  or  a  \CONS\  cell  with  possibly  unevaluated
constituents) or  a {\em closure}.   A closure  is a pair  $\langle s,
\rho\rangle$ in  which $s$ is  an unevaluated application,  and $\rho$
maps free variables of $s$ to their respective locations. Since all data objects are boxed, we model an environment as a mapping
from the set  of variables of the program  $\mathit{Var}$ to locations
in the  heap.

The  semantics  of   expressions  (and  applications\footnote{In  most
  contexts, we  shall use the  term 'expression' and the  notation $e$
  for both expressions and applications.}) are given by transitions of
the form $\rho, \stk, \heap,  e \rightarrow \rho', \stk', \heap', e'$.
Here \stk\ is a stack of continuation frames.  Each continuation frame
is of the form $(\rho, \ell,  e)$, signifying that the location $\ell$
has  to  be  updated  with  the  value  of  the  currently  evaluating
expression, and $e$ is to be evaluated next in the environment $\rho$.
Note that the  state of the transition system consists  of the current
evaluation  context,  given  by  $\rho$ and  $e$,  and  the  suspended
evaluation contexts (represented as  continuation frames) in the stack
$S$.  The heap $H$ is common to all contexts.

The  initial  state  of  the  transition  system  described  above  is
$([\;]_\rho,\,   ([\;]_\rho,   \,   \ans,\,   (\print~\ans)):[\;]_{S},
[\;]_{H},\, (\mainpgm))$,  in which  $[\;]_\rho$ and $[\;]_H$  are the
initial environment and  the initial heap. The  initial stack consists
of  a single  continuation frame  in  which \ans\  is a  distinguished
variable  that  would   eventually  be  updated  with   the  value  of
(\mainpgm).  In  addition, \print\ is  a function modeling  a printing
mechanism---a   standard   run-time   support  assumption   for   lazy
languages---that prints  the value of  (\mainpgm).  Notice the  use of
the operator $:$ for adding elements to the top of the stack.

The   notation  $[\myvec{x}   \mapsto  \myvec{\ell}]$   represents  an
environment  that  maps  variables  $x_i$ to  locations  $\ell_i$  and
$\heap[\ell := d]$  indicates the updating of a heap  \heap\ at $\ell$
with  $d$.   $\rho \oplus  \rho'$  represents  the environment  $\rho$
shadowed  by  $\rho'$  and  $\lfloor \rho  \rfloor_X$  represents  the
environment  restricted  to  the  variables in  $X$.  Finally  $FV(s)$
represents the  free variables  in the  application $s$  and $addr(c)$
gives the address of the closure $c$ in the heap.

The small-step semantics  is shown in Figure~\ref{fig:lang-semantics}.
As a  sample, consider the  three rules for  $(\CAR~x)$.  If $x$  is a
closure, it is evaluated to WHNF,  say $(d_1, d_2)$.  This is given by
the  rule  {\sc  car-clo}.  If  $d_1$  is  not  in  WHNF, it  is  also
evaluated  ({\sc car-1-clo}). The address ($addr (d_1)$) to be updated with the evaluated
value is  recorded in  a continuation frame.  This is
required for lazy evaluation,
%of an application involving a selector such  as $\CAR$, 
else, in the presence of sharing,   $d_1$  may  be  evaluated  more  than  once~\cite{Jones87}.    Only after this  is the
actual selection done ({\sc car-select}).

}
%==============================================================
\renewcommand{\pp}[2]{\ensuremath{#1\!\!:\!#2}} % prog point
%==============================================================
\begin{figure*}[t]
\renewcommand{\arraystretch}{1}
\begin{eqnarray*}
\mathit{ref\/}(\kappa,\sigma,\Lfonly)
          &=& \{\,\} \mbox{, for $\kappa$ a constant, including
$\NIL$}\\
\mathit{ref\/}((\CONS~x~y),\sigma,\Lfonly)
          &=& \{x.\alpha \mid \acar\alpha \in \sigma\} \cup \{y.\alpha
\mid \acdr\alpha \in \sigma\} \\
\mathit{ref\/}((\CAR~x),\sigma,\Lfonly)
          &=&    \begin{array}{l l}
                    \{x.\epsilon\} \cup \{x.\acar\alpha \mid \alpha \in
\sigma\}, & \mbox{if}~\sigma \ne \emptyset\\
                    \emptyset  & \mbox{otherwise}
                 \end{array} \\
\mathit{ref\/}((\CDR~x),\sigma,\Lfonly)
          &=&    \begin{array}{l l}
                    \{x.\epsilon\} \cup \{x.\acdr\alpha \mid \alpha \in
\sigma\}, & \mbox{if}~\sigma \ne \emptyset\\
                    \emptyset  & \mbox{otherwise}
                 \end{array} \\
\mathit{ref\/}((\PRIM~x~y),\sigma,\Lfonly)
          &=&    \begin{array}{l l}
                    \{x.\epsilon, y.\epsilon\},  & \mbox{if}~\sigma \ne
\emptyset\\
                    \emptyset  & \mbox{otherwise}
                 \end{array} \\
\mathit{ref\/}((\NULLQ~x),\sigma,\Lfonly)
          &=&    \begin{array}{l l}
                    \{x.\epsilon\},  & \mbox{if}~\sigma \ne \emptyset\\
                    \emptyset  & \mbox{otherwise}
                 \end{array} \\
\mathit{ref\/}((f~\myvec{x}),\sigma,\Lfonly)
%          &=& \bigcup_{i=1}^n y_i.\Lf{f}{i}{\sigma}
          &=&  \begin{array}{@{}l}  % to discourage \displaystyle
               \bigcup_{i=1}^n x_i.\Lf{f}{i}{\sigma}
               \end{array}
%          &=& \bigcup \{y_i.\Lf{f}{i}{\sigma} \mid i=1,\ldots, n\}
\\[1ex]
\mathcal{L}((\SRETURN~\psi:x),\sigma,\Lfonly) &=& \lbrack \mathit{\epath}
\mapsto
x.\sigma\rbrack, \mbox{ where $\mathit{\epath}$ is a
  new e-path terminating with $\psi$}\\
\mathcal{L}((\SIF~\psi:x~e_1~e_2),\sigma,\Lfonly) &=&
\begin{array}{l l}
                    \mathcal{L}(e_1,\sigma,\Lfonly) \uplus
        \mathcal{L}(e_2,\sigma,\Lfonly) \uplus
        \lbrack\lbrace \mathit{\epath} \mapsto
\{x.\epsilon\}\rbrace\rbrack,  & \mbox{if}~\sigma \ne \emptyset\\
        \emptyset_{\cal{M}}  & \mbox{otherwise}
                 \end{array} \\
&&\mbox{where $\mathit{\epath}$ is a new e-path
  terminating with $\psi$}\\
\mathcal{L}(\LET~x \leftarrow~s~\IN~e),\sigma,\Lfonly) &=&
        \lambda\mathit{\epath}.\; \Lfonly_f(\Lv(x)) \cup \Lv~~
\mbox{where}~\mathcal{M} =
\mathcal{L}(e,\sigma,\Lfonly),~\Lv =
\mathcal{M}(\epath),~\mbox{and}~f~\mbox{is the new function}\\
&& \makebox[0mm]{\hspace*{7cm}
 $(\DEFINE~(f\,x_1\,x_2\, \ldots \, x_n)~(\SIF\,*~x~s[(f\,x_1\,
           x_2\, \ldots\, x_n)/x]))$,} \\
 \makebox[0mm]{\hspace*{8cm} where
     $x_1\, \ldots\, x_n$ are the variables in $s$}
\end{eqnarray*}
\begin{minipage}{0.85\textwidth}
\infrule[live-define]
        {\mathcal{L}(e_f,\sigma,\Lfonly) =
           \bigcup_{i=1}^n z_i.\Lf{f}{i}{\sigma}
              \mbox{ for each $f$ and $\sigma$}
        }
        { \mathit{df_1} \ldots \mathit{df_k} \len \Lfonly
\\ \makebox[0mm]{where
     $(\DEFINE\ (f\ z_1\ \ldots\ z_n)\ \ e_f)$ is a member of $\mathit{df_1}
\ldots \mathit{df_k}$}}
\end{minipage}
  \caption{Liveness equations and judgement rule}\label{fig:live-judge}
\end{figure*}

\section{Liveness}
\label{sec:liveness}

A variable is {\em live} if there  is a possibility of its value being
used in  future computations and  dead if  it is definitely  not used.
Heap-allocated data needs a richer model of liveness which talks about
liveness of  references.  Using  $\acar$, $\acdr$ to  represent access
using  $\CAR$  and  $\CDR$  fields,  the  liveness  of  the  structure
reachable from a  variable can be represented by a  set of {\em access
  paths}     i.e.     prefix-closed     set     of    strings     from
$\{\acar,\acdr\}^\ast$.  As an example, if $x$ is a list with liveness
$\{\epsilon, \acdr,  \acdr\acar, \acdr\acdr,  \acdr\acdr\acar\}$, then
future computations can only refer up  to the second and third members
of $x$.  A  {\em liveness environment} is a mapping  from variables to
subsets of $\{\acar,\acdr\}^\ast$,  but often expressed as  a set, for
example by writing $\{x.\epsilon, x.\acdr, x.\acdr\acdr, y.\epsilon\}$
instead       of       $[x       \mapsto\{\epsilon,\acdr,\acdr\acdr\},
  y\mapsto\{\epsilon\}, z\mapsto\{\}]$.  In  this notation, $x \mapsto
\{\epsilon\}$ represents access using $x$  itself and $x \mapsto \{\}$
indicates that $x$  is dead.  We associate  liveness environments with
program points.  The  liveness at $\pi:e$ is the  liveness just before
executing $e$.

A  notion  related  to  liveness  is  {\em  demand}.   While  liveness
represents the future use of variables, a demand represents the future
use of the value of an expression.  The demand on an expression $e$ is
again a  set of  access paths---that subset  of $\{\acar,\acdr\}^\ast$
which the context of $e$ may explore of $e$'s result.  To see the need
for  demands,  consider   the  let  expression  $\pi:(\LET~x\leftarrow
(\CDR~y)~\IN~  \pi':\SRETURN~x)$.  Assume  that  the  context of  this
expression produces  the liveness $[x \mapsto  \lbrace \epsilon, \acar
  \rbrace]$ at $\pi'$.  Due to the \LET\ definition which binds $x$ to
$(\CDR~y)$, the  liveness of $x$ at  $\pi'$ now becomes the  demand on
$(\CDR~y)$.    This,  in   turn,  generates   the  liveness   $\lbrace
y.\epsilon, y.\acdr, y.\acdr\acar \rbrace$ at $\pi$. These are the set
of $y$-rooted  accesses required  to explore $\lbrace  \epsilon, \acar
\rbrace$  paths  of  the  result of  $(\CDR~y)$.  In  strong  liveness
analysis       (dual      of       classical      faint       variable
analysis~\cite{horwitz.faint}),  $y$ and  $z$  are live  at the  entry
$\pi: x:=y+z$, if  and only if $x$  is live at exit of  $\pi$.  In our
terminology, the  liveness $\lbrace  \epsilon\rbrace $  of $x$  at the
exit from $n$ becomes the demand on $y+z$, and this, in turn generates
the liveness $\lbrace y.\epsilon, z.\epsilon  \rbrace$ at the entry of
$n$.

We follow the notations appearing in \cite{asati14lgc}. In particular,
We use $\sigma$  to range over demands, $\alpha$ to  range over access
paths and  $\Lv$ to  range over  liveness environments.   The notation
$\sigma_1\sigma_2$  denotes  the  set $\lbrace  \alpha_1\alpha_2  \mid
\alpha_1 \in \sigma_1, \alpha_2  \in \sigma_2\rbrace$.  Often we shall
abuse notation to  juxtapose an edge label and a  set of access paths:
$\acar\sigma$ is a shorthand for $\lbrace\acar\rbrace\sigma$.

\subsection{Liveness Analysis for lazy languages}
\label{sec:liveness-analysis}

Figure~\ref{fig:live-judge}  describes  our  analysis  which  has  two
parts. The function $\mathit{ref}$,  takes  an application $s$ and
a demand $\sigma$, and returns  the incremental liveness generated for the
free  variables   of  $s$  due  to  the   application.   The  function
$\mathcal{L}$  uses   $\mathit{ref}$  to  propagate   liveness  across
expressions.

 Since  in a  lazy language,  an  expression is  not evaluated  unless
 required. The null demand ($\emptyset$) does not generate liveness in
 any  of  the  rules  defining  $\mathit{ref}$  or  $\mathcal{L}$.   A
 non-null  demand of  $\sigma$ on  (\CDR~$x$), is  transformed to  the
 liveness $\{x.\epsilon, x.\acdr\sigma\}$.  In  an opposite sense, the
 demand  of $\acdr\sigma$  on  (\CONS~$y$~$z$) is  transformed to  the
 demand  $\sigma$  on $z$.   Since  \CONS\  does not  dereference  its
 arguments, there  is no $\epsilon$ demand  on $y$ and $z$.  The rules
 for (\PRIM~x~y) and (\NULLQ~x) are similar. Constants do not generate
 any liveness.

For the  case of a  function call we  use a third  parameter $\Lfonly$
that  represents  the  summaries  of all  functions  in  the  program.
$\Lfonly_{\mathit  f}$  (the component  of  $\Lfonly$  for a  specific
function $f$), expresses  how the demand $\sigma$ on a  call to $f$ is
transformed into  the liveness of  its parameters at the  beginning of
the   call.    $\Lfonly$  is   determined   by   the  judgement 
$\mathit{Prog} \len \Lfonly$ using inference rule ({\sc live-define}).
This  rule  describes the  fixed-point  property  to be  satisfied  by
$\Lfonly$, namely, the demand transformation assumed for each function
in  the  program should  be  the  same  as the  demand  transformation
calculated    from    its    body.      As    we    shall    see    in
Section~\ref{sec:grammar-formulation},  we  convert  the rule  into  a
grammar and  the the language generated  by this grammar is  the least
solution satisfying  the rule. We  prefer the least solution  since it
ensures the safe collection of the greatest amount of garbage.

We next  describe the function $\mathcal{L}$  that propagates liveness
across expressions.  Note that,  unlike eager languages, the syntactic
occurrence  of an expression  in a  program denotes  the point  of
creation of its closure, not necessarily its evaluation.
To see the  consequences of this, consider
Figure~\ref{fig:mot-example2} which shows an analysis  of the body of
a function $\length$ with a demand $\sigma$.  Clearly, this is also the
demand on the return value $z$ giving the liveness at the program point
$\pi_8$ as $\{\pz.\sigma\}$.  The liveness at $\pi_8$ is propagated
backwards as in traditional liveness analysis.

Since   \pz\  is   $(1  +   \py)$,  the   liveness  of   $\sigma$  for
\pz\ translates into a demand  of $\sigma$ on $(1+\py)$, which in turn
generates a  liveness of $\{\py.\epsilon\}$ at  $\pi_7$.  We now make
the following observations:
\begin{enumerate}
\item While the liveness of \py\  should be killed at $\pi_6$ since it
  is  beyond the scope  of the  definition of  \py, we  do not  do so.
  However, this  does not cause any  imprecision in the way  of the GC
  marking more cells as live,  since, during execution, \py\ is yet to
  come  into  existence at  $\pi_6$,  and  its  liveness will  not  be
  considered by the garbage collector at this point.
\item In an  eager language, the heap location referred  by \py\ would
  not have been live at $\pi_8$,  a program point beyond its last use.
  In  a  lazy  language,   however,  the  definition  $\pz  \leftarrow
  (1+\py)$, does not  result in an evaluation of  $(1+\py)$; instead a
  closure  is  created.  The  evaluation  actually  takes place  while
  evaluating $(\SRETURN~\pz)$.   We therefore  record the  liveness of
  $\py$ at  the program point  $\pi_8$, indeed at every  program point
  from $\pi_8$ up to the beginning of the program $\pi_1$ (since we do
  not kill liveness).
\end{enumerate}

We thus  regard the  body of  the function as  an expression  tree and
consider paths from  the root of this tree to  the nodes corresponding
to the program points which trigger evaluation\footnote{Program points
  which      trigger      evaluation       are      labeled      with
  $\psi$.}---\SIF\ conditions and \SRETURN\ expressions.  We call such
paths {\em e-paths}.  For example, for the function \length, there are
three  such paths  $\epath_1 =  \langle \pi_1,  \pi_2, \psi_1\rangle$,
$\epath_2  = \langle\pi_1,  \pi_2,  \pi_3,  \pi_4, \psi_2\rangle$  and
$\epath_3  =  \langle  \pi_1,   \pi_2,  \pi_5,  \pi_6,  \pi_7,  \pi_8,
\psi_3\rangle$.

Given  a function  $\mathit{f}$ and  a demand  $\sigma$,  the liveness
analysis  computes  two   maps:  $\mathcal{L}(e_f,  \sigma,  \Lfonly)$
computes a  map that takes an  e-path in ${\mathit f}$  and returns the
liveness environment  arising out of  all uses of variables  along the
e-path.  A second map $\cal{P}_{\mathit  f}$ maps any program point in
$\mathit f$  to the set  of e-paths sharing  the point.  If  a program
point happens to fall on more  than one such e-path, then the liveness
environment of  the program  point is the  variable-wise union  of the
liveness  environments  of the  individual  e-paths.   In the  example
program,  since ${\mathcal P}_{\length}(\pi_1)  = \lbrace  \epath_1, \epath_2,
\epath_3\rbrace$, the  liveness at $\pi_1$  is ${\cal M}(\epath_1)  \cup {\cal
  M}(\epath_2) \cup  {\cal M}(\epath_3)$,  where ${\cal M}  = \mathcal{L}(e_f,
\sigma, \Lfonly)$.  In the  sequel we shall only define $\mathcal{L}$.
The formulation  of $\cal{P}$ is easy  and we do not  elaborate it any
further.

Consider the  $\mathcal{L}$-rules for {\LET}, {\SIF},  and {\SRETURN}.
The expression  $(\SRETURN~\psi:x)$ with  a demand $\sigma$  forces an
evaluation  of  $x$ and  hence  a  new map  $[\mathit{\epath}  \mapsto
  \{x.\sigma\}]  $ needs  to be  created,  where $\epath$  is a  fresh
evaluation path  terminating in  $\psi$.  The  map for  the expression
$(\SIF~\psi:x~e_1~e_2)$ is a point-wise-union of the maps of $e_1$ and
$e_2$. In addition, since the condition also triggers an evaluation of
$x$, the map $[\mathit{\epath} \mapsto \{x.\epsilon\}]$ is created and
added  to the  union.   Also notice  a  consequence of  laziness: the
entire  expression including  the condition  is not  evaluated if  the
demand on it is $\emptyset$.  This  results in the empty map, which is
denoted by $\emptyset_{\cal{M}}$.

 The liveness  of {\LET} has  an elegant formulation.   Recollect that
 our language permits lazy {\LET}s whose definitions can be recursive,
 i.e.  in the expression $\LET~x \leftarrow~s~\IN~e$, $x$ can occur in
 $s$.   We  model  the  definition  $x  \leftarrow~s$  as  a  possibly
 recursive function (named $\mathit{f}$ in the rule) which captures an
 arbitrary number of ``unrollings'' of the definition\footnote{The $*$
   in  the definition  of ${\mathit  f}$ represents  non-deterministic
   choice.  During liveness analysis it  can be assumed to generate no
   liveness.}.  We  pass to  the demand  transformer $\Lfonly_{\mathit
   f}$ of this function the demand  arising out of the liveness of $x$
 at the beginning of $e$. This  gives the liveness of the variables of
 $s$.

Section~\ref{sec:computing}   shows   how  the   demand   transformers
$\Lfonly$  for  a  program  (representing  a  fully  context-sensitive
analysis) can be safely approximated  by a {\em procedure summary} for
each function.   The summary is  in the  form of a  demand transformer
that maps a demand on a call to the function to demands on each of its
arguments.

\subsection{Correctness of liveness analysis}

We shall now  outline a proof of correctness of  the liveness analysis
presented   in    Section~\ref{sec:liveness-analysis}.    This   proof
technique  is  similar to  the  technique  used in  \cite{asati14lgc}.

The  idea behind  the  proof is  as follows:  We  modify the  standard
semantics to a semantics called  {\em minefield semantics} that models
a liveness based  garbage collection at each  transition.  Before each
transition,  we start  from the  root-set and  insert a  special value
$\bot$ at each  location in the heap that contains  a reference and is
reachable but not live. The proof is based on a context-sensitive form
of    liveness   of    which   our    calculated   liveness    is   an
over-approximation. The state of the transition system is augmented to
carry enough information to calculate  the liveness environment at the
current  evaluation  context  and  each of  the  suspended  evaluation
contexts on  the stack.   To simulate  garbage collection, before each
transition,  we start  from the  root-set and  insert a  special value
$\bot$ at each  location in the heap that contains  a reference and is
reachable but  not live by  our analysis.  Any attempt  to dereference
such locations during the transition will result in entering a special
state denoted \bang.  The main proof is to show that no program enters
the \bang\ state in the minefield semantics.

\newcommand{\cred}[1]{\psframebox[linestyle=none,
    fillcolor=lightgray,fillstyle=solid,framesep=0.5pt]{#1}}

\comment{
\begin{enumerate}
\item Do you need to carry the function body?
\item Minefield doesn't insert bottom inside closures. Separate rule to
be inserted.
\item Examine Proposition 4.1, See that different cases are disjoint
  and together cover all possible cases
\item Suppose the statement that we want to prove is this: Every step
  (step = GC + $\rightsquigarrow$) goes without a bang. How to divide
  a (n+1)-step transition into 1 and n. What is the exact invariant
  required before each step. 
\end{enumerate}

}

\section{Minefield semantics for the language}
\begin{figure*}[ht!]
\begin{center}\footnotesize
\begin{tabular}{|c|c|c|}
\hline
Premise & Transition & Rule name \\ 
\hline
\hline
          &\makecell{ $\rho, (\rho', x, e, \cred{\sigma', e_f'})\!:\!S,
  H, \kappa, \cred{\sigma, e_f}$ \\ $\rightsquigarrow \rho', S, H[\rho'(x) :=
    \kappa], e, \cred{\sigma', e_f'}$ }   &  \sc{const}
\\
\hline
          & \makecell{$\rho, (\rho', z, e, \cred{\sigma',
    e_f'})\!:\!S, H, (\CONS~x~y), \cred{\sigma, e_f}$ \\ $\rightsquigarrow
  \rho', S, H[\rho'(z) := (\rho(x),\rho(y))], e, \cred{\sigma', e_f'}$}     &  \sc{cons} \\
\hline
$H(\rho(x)) \mbox{ is } (l_1, l_2)$ & \makecell{$\rho, (\rho', z, e,
  \cred{\sigma', e_f'} )\!:\!S, H, (\CAR~x), \cred{\sigma, e_f}$ \\ $
  \rightsquigarrow \rho', S, H[\rho'(z) := H(l_1)], e, \cred{\sigma', e_f'}$}      &
\sc{car-whnf} \\
\hline
$H(\rho(x)) \mbox{ is } \langle s, \rho'\rangle$ &\makecell{ $\rho, S,
  H, (\CAR~x), \cred{\sigma, e_f}$ \\ $\rightsquigarrow \rho', (\rho, x,
  (\CAR~x), \cred{\sigma, e_f})\!:\!S, H, s, \cred{(\clazy \cup \acar)\sigma , \_ }$}      &
\sc{car-clo}
\\
\hline
$H(\rho(x)), H(\rho(y)) \in \mathbb{N}$
 & \makecell{$\rho, (\rho', z, e, \cred{\sigma', e_f'})\!:\!S, H,
  (+~x~y), \cred{\sigma, e_f}$ \\  $\rightsquigarrow \rho', S, H[\rho'(z)
    := H(\rho'(x)) + H(\rho'(y))], e, \cred{\sigma', e_f'}$}      &
\sc{prim-whnf} \\
\hline
$H(\rho(x)) \mbox{ is } \langle s, \rho'\rangle$ &\makecell{$\rho, S,
  H, (+~x~y), \cred{\sigma, e_f}$ \\ $\rightsquigarrow \rho', (\rho, x,
  (+~x~y), \cred{\sigma, e_f})\!:\!S, H, s, \cred{\clazy\sigma, \_}$}      &
\sc{prim-1-clo} \\
\hline
$H(\rho(y)) \mbox{ is } \langle s, \rho'\rangle $ & \makecell{$\rho,
  S, H, (+~x~y), \cred{\sigma, e_f}$ \\ $\rightsquigarrow \rho', (\rho, y,
  (+~x~y), \cred{\sigma, e_f})\!:\!S, H, s, \cred{\clazy\sigma, \_ }$}      &
\sc{prim-2-clo} \\
\hline
{$\mathit{g}~\mbox{defined as}$
$~(\DEFINE~(g~\myvec{y})~e_{\mathit{g}})$}  & \makecell{$\rho, S, H,
  (g~\myvec{x}), \cred{\sigma, e_f}$ \\ $\rightsquigarrow [\myvec{y} \mapsto
    \rho(\myvec{x})], S, H, e_{\mathit{g}}, \cred{\sigma, e_g}$}      &
\sc{funcall} \\
\hline
$\ell$ is a new location& \makecell{$\rho, S, H, (\LET~x\leftarrow
  s~\IN~e), \cred{\sigma, e_f}$ \\ $ \rightsquigarrow \rho\oplus[x \mapsto \ell], S, H[\ell \mapsto \langle s, \lfloor\rho\rfloor_{FV(s)}  \oplus [x \mapsto
  \ell]\rangle], e, \cred{\sigma, e_f}$} &
\sc{let} \\
\hline
$H(\rho(x)) \ne 0$ & \makecell{$\rho, S, H, (\pi:\SIF~\psi:x~e_1~e_2),
  \cred{\sigma, e_f}$ \\ $\rightsquigarrow \rho, S, H,  e_1, \cred{\sigma, e_f}$} & \sc{if-true} \\
\hline
$H(\rho(x)) = 0$ & \makecell{$\rho, S, H, (\pi:\SIF~\psi:x~e_1~e_2),
  \cred{\sigma, e_f}$ \\   $\rightsquigarrow
\rho, S, H,  e_2, \cred{\sigma, e_f}$} & \sc{if-false} \\
\hline
$H(\rho(x)) = \langle s, \rho' \rangle $ & \makecell{$\rho, S, H,
  (\pi:\SIF~\psi:x~e_1~e_2), \cred{\sigma, e_f}$ \\ $\rightsquigarrow
\rho', (\rho, x, (\SIF~x~e_1~e_2),  \cred{\sigma, e_f})\!:\!S, H, s,
\cred{\clazy\sigma, e_f}$}
&
\sc{if-clo} \\
\hline
{$H(\rho(x))~\mbox{is}$ $\mbox{whnf with value}~v$}& \makecell{$\rho,
  (\rho', z, e, \cred{\sigma', e_f'})\!:\!S, H, (\SRETURN~x), \cred{\sigma,
  e_f}$ \\ $\rightsquigarrow \rho', S, H[\rho'(z) := v], e,
  \cred{\sigma', e_f'}$} &
\sc{return-whnf}\\
\hline
$H(\rho(x)) = \langle s, \rho' \rangle $ & \makecell{$\rho, S, H,
  (\psi:\SRETURN~x), \cred{\sigma, e_f}$ \\ $
  \rightsquigarrow$
$\rho',~ (\rho, x, (\SRETURN~x), \cred{\sigma, e_f})\!:\!S, H,  s,
  \cred{\sigma, e_f}$} &
\sc{return-clo} \\
\hline
\end{tabular}
\caption{Minefield semantics.\label{fig:minefield-semantics}}
\end{center}
\vskip -5mm
\end{figure*}

To set up the minefield semantics, we follow these steps:
\begin{enumerate}
\item We enrich  the abstract machine state $\rho,$ $S,$  $ H,$ $e$ to
  $\rho,$ $S,$ $H,$ $e,$ $\sigma,$ $e_f$, where $\sigma$ is the demand
  on the expression $e$, and $e_f$ is the body of the function. As the
  section on liveness  analysis shows, the entire  function body $e_f$
  and $\sigma$  are needed  to calculate  the liveness  environment at
  different program  points. The  demand sigma  is ``dynamic''  in the
  sense  that it  arises from  the actual  function calls  made during
  execution.   The  0-CFA  demand  is an  over-approximation  of  this
  demand.  The information  in a suspended evaluation  context is also
  similarly augmented with  the demand on it and the  function body of
  which the suspended evaluation is a  part.  Thus a stacked entry now
  takes the  form $(\rho, x, e,  \sigma, e_f)$. A modification  of the
  small step semantics which carries the extra information is shown in
  Figure~\ref{fig:minefield-semantics}.
\item Given  the current  evaluation context $\rho,  S, H,  e, \sigma,
  e_f$ and $\mathcal{LF}$, we can construct a liveness environment as follows:
  \begin{enumerate}
  \item If $e$ is an expression, then:  
  $$ \Lv = \bigcup_{\psi \in \mathcal{P}(e)} \mathcal{M}(\psi),
    \text{where }
  \mathcal{M} = \mathcal{L}(e_f, \sigma, \Lfonly)$$
\item If $e$ is an application, then:  
  $$ \Lv = ref(e, \sigma, \Lfonly)$$
  \end{enumerate}
We can define in the same way the liveness environment for each of the
suspended  evaluation contexts in  $S$  giving  a  stack of  liveness
environments that we shall denote $\mathsf{SL}$.
\item $GC$ models a liveness based garbage collection:
$GC(\rho,$ $ S,$ $ H,$ $ \mathcal{L},$ $ \mathsf{SL}) = (\rho', S', H')$ 
  \begin{enumerate}
  \item For each $x \in domain(\rho)$,  $\rho'(x)=\bot$ iff $x.\epsilon
    \notin \Lv$.
\item For each stack entry $(\rho,\_,\_,\_,\_)$ in $S$ with $\Lv'$ as
  the corresponding liveness environment in $\mathsf{SL}$,  and for each $x
  \in domain(\rho)$,  $\rho'(x)=\bot$
  iff $x.\epsilon \notin \Lv'$.
\item For each location $\ell$, $H'(\ell) = \bot$, iff  there is no
  $x$ in either the current environment or one of the stacked
  environments $\rho$ such that for some $\alpha$, $H[x.\alpha] =
  \ell$  and $x.\alpha \in \Lv$, .
  \end{enumerate}
\end{enumerate}

Starting   from  the   initial  state   $([\;]_\rho,([\;]_\rho,  \ans,
(\print~\ans)):[\;]_{S} ,  [\;]_{H}, (\mainpgm)$, every  transition is
preceded by  a garbage collection  using $GC$.  We consider  a garbage
collection followed  by a transition as  a step. We show  by induction
that, for programs  without function calls, starting  from the initial
state that  has $e_{\main}$  as the first  expression to  be evaluated
instead  of $\main$,  any transition  of  $n$ steps  occurs without  a
\bang.

Observe  that  the  transitions  that dereference  and  therefore  can
potentially lead to \bang\ are those  which have a $H(\ldots)$ term in
the  transition  rule.   Of  these,  terms  of  the  form  $H(\ldots)$
dereference a root  variable, and the term $H(l)$  dereferences a heap
location.  In  both the  cases we show  that the  dereferencing cannot
lead to a \bang\ state in programs without function calls.

We  need the  following proposition that  connects  liveness analysis  with
minefield semantics transitions.

\begin{proposition}
\label{prop:correctness}
\begin{enumerate} 
\item   Consider  a   \LET\  definition:   \linebreak  $\pi\!\!\!:   (\LET~x
  \leftarrow~\ldots~\IN~e)$, such  that there is  a use of $x$  in $e$,
  i.e.  an occurrence of $\LET~y~\leftarrow~s~in \ldots$, where $s$ is
  $(\CAR\ x)$, $(\CDR\ x)$  or $(+\ x\ y)$, or $(\SIF\  x\ e1\ e2)$ or
  $(\SRETURN\ x)$. Then the demand on $x$ at $\pi$ is non-null.
\item    Assume    that    a   \LET\    definition    $\pi:    (\LET~z
  \leftarrow~(\CONS~x~y)~\IN~e)$  has an  occurrence  of $(\CAR~z)$  in
  $e$.  Then  the  liveness  of  $x$  at  $\pi$  includes  $\epsilon$.
  Similarly  if $e$  contains $(\CDR~z)$,  then the  liveness of $y$  at
  $\pi$ includes $\epsilon$.
\end{enumerate}
\end{proposition}

In  the second  case for  example,  the liveness  of $x$  due to  the
\CONS\ is  $\bcar\sigma$, where $\sigma$  is the demand on  $z$. Since
$\sigma$  includes  the  demand  $\acar\alpha$   due  to  its  use  in
$(\CAR~z)$, the liveness on $x$  includes $\epsilon$. We thus have the
following lemma:

\begin{lemma}\label{lemma:no-calls-correctness} 
A program without function calls cannot  enter the \bang\ state in the
minefield semantics.
\end{lemma}

\begin{proof}
For a program  executing according to {\em minefield  semantics}, if a
\bang\ occurs then some variable which is  dereferenced must be
set to  $\bot$, between its  creation and  use points.  This  can only
happen if  the liveness of  the variable being dereferenced  was null.
From  the Proposition~\ref{prop:correctness}  it  is clear  that this  cannot
happen.
\end{proof}

For   a    program   with   functions,   we    use   the   techniques
in~\cite{asati14lgc}. We  assume as  a contradiction that  the program
can enter a \bang\ state  and inline the  functions called to  obtain a
program that is without functions  but has the same execution behavior
and    demands    as   the    original    program.     We   then    use
Lemma~\ref{lemma:no-calls-correctness}  to  show  that  the  resulting
functionless program cannot enter a bang state.

\section{Computing liveness and its encoding as DFA}\label{sec:computing}
Section~\ref{sec:liveness} gave a  context-sensitive liveness analysis
and proved it  correct with reference to a  {\em minefield} semantics.
In  our  analysis, $\mathcal{L}$  and  \pmap{f}  together described  a
liveness set for each program point in  a function in terms of a given
$\sigma$  and \Lfonly.   However, we  still  have to  describe how  to
obtain demand  transformers \Lfonly\  from the rule  {\sc live-define}
and how to compute the specific demand $\sigma$ on each function.
%%
To  answer these  questions, it  is convenient  to cast  the equations
arising from  $\mathit{ref}$ and  $\mathcal{L}$ as  the rules  of a
grammar.   To do  so, we  need  to modify  the rules  themselves to  a
different form.

\subsection{Modifying liveness rules}

The      $\mathit{ref}$     rule      for     \CONS,      shown     in
Figure~\ref{fig:live-judge},  requires   us  to  remove   the  leading
\acar\ and \acdr\  from the access paths in  $\sigma$.  Similarly, the
rules for  \CAR, \CDR, \PRIM, \NULLQ,  and \SIF\ require  us to return
$\emptyset$, if  $\sigma$ itself is $\emptyset$ and
$\lbrace\epsilon\rbrace$ otherwise.  To  realize these rules
$\sigma$ needs to be known. This creates difficulties since we want to
solve the equations arising from liveness symbolically.

The  solution  is  to  also   treat  the  operations  mentioned  above
symbolically.  We  introduce three new symbols:  \bcar, \bcdr, \clazy.
These symbols are defined as a relation $\hookrightarrow$ between sets
of access paths:
\begin{align*}
  &\bcar\sigma \hookrightarrow \sigma' \mbox{ where } \sigma' = \{\alpha \mid \acar\alpha \in \sigma\}\\
  &\bcdr\sigma \hookrightarrow \sigma' \mbox{ where } \sigma' = \{\alpha \mid \acdr\alpha \in \sigma\}
\end{align*}
Thus \bcar\ selects those entries in $\sigma$ that have leading \acar, and removes the leading \acar\ from them.
The symbol \clazy\ reduces the set of strings following it to a set containing only $\epsilon$. It filters out, however, the empty set of strings.
\begin{align*}
  \clazy\sigma \hookrightarrow & \left\{ 
  \begin{array}{ll}
    \emptyset&\mbox{if}~\sigma = \emptyset\\
    \{\epsilon\} & \mbox{otherwise}
  \end{array}\right.
\end{align*}
We can  now rewrite the \CONS\  and the \CAR\  rules of $\mathit{ref}$
as:
\begin{align*}
&\mathit{ref\/}((\CONS~x~y),\sigma,\Lfonly)
= x.\bcar\sigma \cup y.\bcdr\sigma  \label{eqn:mod-cons},~
\mbox{and} \\
&\mathit{ref\/}((\CAR~x),\sigma,\Lfonly)
          =   x.\clazy\sigma \cup x.0\sigma
\end{align*}
and the \Lfunonly\ rule
for \SIF\ as:
\begin{align*}
\mathcal{L}((\SIF~x~e_1~e_2),\sigma,\Lfonly) =
                    &\mathcal{L}(e_1,\sigma,\Lfonly)~\uplus
        \mathcal{L}(e_2,\sigma,\Lfonly)~\uplus\\
        &[\,\mathit{\epath} \mapsto  x.\clazy \sigma]
\end{align*}
The rules for  \CDR, \PRIM\ and \NULLQ\ are also  modified in a manner
similar to \CAR.

When  there  are  multiple  occurrences of  \bcar,  \bcdr\  and  \clazy,
$\hookrightarrow$  is  applied  from  right to  left.   The  reflexive
transitive   closure  of   $\hookrightarrow$   will   be  denoted   as
$\stackrel{*}{\hookrightarrow}$.   The  following proposition  relates
the original and the modified liveness rules.
\begin{proposition}
Assume that a liveness computation  based on the original set of rules
gives the liveness  of the variable $x$ at a  program point $\pi_i$ as
$\sigma$   (symbolically,   $\Lanv{i}{x}=   \sigma$).   Further,   let
$\Lanv{i}{x}=  \sigma'$ when the  modified rules  are used  instead of
\Lfunonly.  Then $\sigma' \stackrel{*}{\hookrightarrow} \sigma$.
\end{proposition}

For  an explanation  of why  the  proposition holds  for the  modified
\CONS\  rule,   we  refer   the  reader  to   \cite{asati14lgc}.   The
proposition also holds for other modified rules for similar reasons.

\newcommand{\emm}[2]{\ensuremath{\mathcal{#1}_{#2}}}
\subsection{Generating Liveness Equations}

\begin{figure*}[t!]
\renewcommand{\arraystretch}{1}
  \begin{tabular}{cc}
    \begin{minipage}{.40\textwidth}
        \small
        \renewcommand{\arraystretch}{1}{
          \begin{uprogram}
            \UNL{1} (\DEFINE\ (\length~\xl)
            \UNL{2}  $\pi_1\!\!:\, $(\LET\ \px\ $\leftarrow $\
(\NULLQ~\xl) \IN
            \UNL{3} \hspace*{.05cm} $\pi_2\!\!:\,$(\SIF\
$\psi_1\!\!:\,$ \px
            \UNL{4} \hspace*{.27cm} $\pi_3\!\!:\,
            $(\LET\ \pv\ $\leftarrow 0$ \IN
            \UNL{5} \hspace*{.32cm} $\pi_4\!\!:\,
(\SRETURN~\psi_2:\pv)$
            \UNL{4} \hspace*{.29cm}    $\pi_5\!\!:\, $(\LET~\pu\
$\leftarrow$  (\CDR~\xl)  \IN
            \UNL{5} \hspace*{.34cm}   $\pi_6\!\!:\, $(\LET~\py\
$\leftarrow$  (\length~\pu)  \IN
            \UNL{6} \hspace*{.34cm} $\pi_7\!\!:\,
            $(\LET~\pz\ $\leftarrow$ (1~+~\py) \IN
            \UNL{7} \hspace*{.34cm} $\pi_8\!\!:\,
(\SRETURN~\psi_3:\pz)$)))))))
        \end{uprogram}}
        \renewcommand{\arraystretch}{1}{
	  \begin{uprogram}
	  \UNL{1} $(\DEFINE\ (\main)$
           \UNL{2} \!\!$\pi_9\!\!:\, (\LET\  \pa\  \leftarrow$
           (
           \scalebox{0.8}{\psframebox[framearc=.5,
	fillcolor=lightgray,fillstyle=solid,framesep=2pt]{%
             \begin{tabular}{@{}c@{}}
               {a BIG closure}
             \end{tabular}}}
           ) \IN  
          \UNL{3} \!\!$\pi_{10}\!\!:\, (\LET\  \pb\  \leftarrow ($+$\ \pa\ \acdr)$
\IN
	  \UNL{4}   \hspace*{.05cm}$\pi_{11}\!\!:\,      $ (\LET\ \pc\
$\leftarrow  (\CONS\ \pb\ \NIL)$ \IN
          \UNL{5}   \hspace*{.15cm}    $\pi_{12}\!\!:\,
          $(\LET\ \pw\  $\leftarrow  (\length\ \pc)$ \IN
          \UNL{6}  \hspace*{.25cm}  $\pi_{13}\!\!:\,
(\SRETURN~\psi_4:\pw)))))$
\end{uprogram}}
        %%}
    \end{minipage}

    &

    \begin{minipage}{.51\textwidth}

      \small
\begin{eqnarray*}
        \mathcal{L}(e_{\length},\sigma,\Lfonly) &=&
              [\;\epath_1 \mapsto \lbrace \px.\clazy \sigma,
\xl.\clazy\sigma\rbrace ,~\epath_2 \mapsto \lbrace 
\pv.\sigma  \rbrace,\\
 &&          \;\;  \epath_3 \mapsto \lbrace \pz.\sigma,
                \py.\clazy\sigma,
\pu.\Lf{\length}{\mbox{1}}{\clazy\sigma},\\
&&            \;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;
\xl.\{ \acdr\Lf{\length}{\mbox{1}}{\clazy\sigma} \cup
%%\;\;\;\;\;\;\; \;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\\
\clazy\Lf{\length}{1}{\clazy\sigma}\}\}
             \rbrack \\
            \cal{P_{\mbox{\tiny \length}}} &=&
                [\;\pi_1 \mapsto \lbrace \epath_1, \epath_2, \epath_3 \rbrace,  \
                 \pi_2 \mapsto \lbrace \epath_1, \epath_2, \epath_3
                  \rbrace, \\
&&\;\;  \pi_3 \mapsto \lbrace \epath_2
                  \rbrace ,  \pi_4 \mapsto \lbrace \epath_2
                  \rbrace, \pi_5 \mapsto \lbrace  \epath_3 \rbrace, \\
 &&  \;\;                        \pi_6 \mapsto \lbrace \epath_3 \rbrace,
\pi_7 \mapsto
                   \lbrace \epath_3\rbrace, \pi_8 \mapsto \lbrace \epath_3
\rbrace~\rbrack\\
              \mathcal{L}(e_{\main},\sigma,\Lfonly) &=&
                [\;
                  \epath_4 \mapsto \lbrace \pw.\sigma,
                                       \pc.\Lf{\length}{\mbox{1}}{\sigma},\\
&& \;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;
\pb.\bcar\Lf{\length}{\mbox{1}}{\sigma},
\pa.\clazy\bcar\Lf{\length}{\mbox{1}}{\sigma} \rbrace
                               ] \\
               \cal{P_{\mbox{\tiny \main}}} &=&
                  \lbrack~ \pi_9 \mapsto \lbrace \epath_4 \rbrace,  \
                    \pi_{10} \mapsto \lbrace \epath_4 \rbrace,  \
                   \pi_{11} \mapsto \lbrace \epath_4 \rbrace, \\
                   && \;\; \pi_{12} \mapsto \lbrace \epath_4 \rbrace,
\pi_{13} \mapsto \lbrace \epath_4 \rbrace~\rbrack
\end{eqnarray*}
    \end{minipage} \\
(a)&
(b)
    %% \end{ tabular}}
  \end{tabular}
%\end{picture}
\kern -3ex
%\vspace{0.5in}
\caption{(a) Example program and (b) its liveness
maps.}\label{fig:mot-example2}
\end{figure*}
%================================================================

Given  a  function  $\mathit{f}$,  we now  describe  how  to  generate
equations  for the  demand  transformation \Lfonly$_\mathit{f}$.   The
program in Figure~\ref{fig:mot-example2} serves  as a running example.
Starting   with    a   symbolic   demand   $\sigma$,    we   determine
\Lfun{e_{\mathit{f}}}{\sigma}{\Lfonly}.   In  particular, we  consider
$\Lanv{1}{x_i}$,  the liveness  of  the $i^{\mbox{\footnotesize  th}}$
parameter   $x_i$  at   the  program   point  at   the  beginning   of
$e_{\mathit{f}}$   (assumed  to   be  $\pi_1$).   By  the   rule  {\sc
  live-define},      this     should      be      the     same      as
$\Lf{f}{i}{\sigma}$. Applying this to \length, we have:
\begin{eqnarray*}
&& \Lanv{1}{\xl} = \clazy\sigma \cup \acdr\Lf{\length}{1}{\clazy\sigma}
  \cup \clazy\Lf{\length}{1}{\clazy\sigma}
\end{eqnarray*}
and  the only equation defining \Lfone{\length} is:
  \begin{eqnarray*}
   && \Lf{\length}{1}{\sigma}
    =  \clazy\sigma \cup \acdr\Lf{\length}{1}{\clazy\sigma} \cup
\clazy\Lf{\length}{1}{\clazy\sigma}
\end{eqnarray*}

In general, the equations for  \Lfonly\ are recursive since, as in the
case  of  \length,  \Lan{1}\  may   have  been  defined  in  terms  of
\Lfone{\mathit{f}}.  However,  it is desirable  to have a  closed form
solution for $\Lfone{\mathit{f}}$.  As mentioned in \cite{asati14lgc},
each  of the liveness  rules modifies  a demand  by prefixing  it with
symbols  in the  alphabet  $\lbrace \acar,  \acdr,\bcar, \bcdr,  \clazy
\rbrace$, and therefore we can assume that $\Lf{f}{i}{\sigma}$ has the
closed form:
\begin{eqnarray}
\label{eq:LF:DI}
  \Lf{f}{i}{\sigma} = \Df{f}{i}\sigma
\end{eqnarray}
where \Df{f}{i} are sets of strings over the alphabet mentioned above.
Substituting   the   guessed   form   in   the   equation   describing
\Lfonly$_{\mathit f}$, and factoring  out $\sigma$, we get an equation
for  \Df{f}{i} that  is  independent of  $\sigma$.   Any solution  for
\Df{f}{i}  yields a  solution for  \Lfonly$_{\mathit f}$.   Applied to
\Lfonly$_{\mathit \length}$, we get:
  \begin{eqnarray*}
&&  \Lf{\length }{1}{\sigma} = \Df{\length}{1}\sigma,~\mbox{and}\\
&&   \Df{\length}{1} = \clazy \cup \acdr\Df{\length}{1}\clazy
       \cup \clazy\Df{\length}{1}{\clazy}
  \end{eqnarray*}

Note that this equation can also be viewed as a CFG with \{\acdr,
\clazy\} as terminal symbols and \Df{\length}{1} as the sole
non-terminal.

\subsection{Generating liveness equations \Lv\  for function bodies}
\label{sec:bodylivenessbodies}

To avoid analyzing the body of a function for each call, we calculated
the liveness  for the arguments and  the variables in  a function with
respect to a symbolic demand  $\sigma$.  To get the actual liveness we
calculate an over-approximation of the  actual demands made by all the
calls and calculate the liveness at each GC
point inside the  function based on this approximation.
The  0-CFA-style  {\em
  summary demand}  is calculated by taking  a union of  the demands at
every call site of a function.

Consider a function $g$ containing a  call to $f$ at a site $\pi$, say
$\pi\!\!:\!(\LET~x   \leftarrow   (f\,y_1\,\ldots\,y_n)~\IN~\pi_i:e)$.
Let the  demand on $g$  be $\sigma_g$ and,  based on this  demand, the
liveness of  $x$ at $\pi_i$ be  $\Lanv{i}{x}$.  By the  $\LET$ rule of
Figure~\ref{fig:live-judge},    the   call   at    $\pi$   contributes
$\Lanv{i}{x}$ to the demand $\sigma_f$. Let us denote the contribution
of a call  site $\pi$ in a  function $g$ to the overall  demand on the
function $f$  as $\deltacall{f}{\pi}{g}$. Assuming that  there are $k$
call sites to function $f$, $\pi^1$ (in function $g^1$) \ldots $\pi^k$
(in   function  $g^k$),  the   over-approximation  of   $\sigma_f$  is
$\deltacall{f}{\pi^1}{g^1}          \cup          \cdots          \cup
\deltacall{f}{\pi^k}{g^k}$.  The distinguished function \mainpgm\ is a
special case.   We assume  it is called  through a  printing mechanism
with   demand   $\sigma_\mainpgm   =  \{\acar,\acdr\}^\ast$ (denoted
$\sigma_{\!all}$) if \mainpgm\ returns a structure and $\epsilon$ if
it returns a base value.

For  the  running  example,   $\length$  has  calls  from  $\main$  at
$\pi_{12}$ and a recursive call at $\pi_6$.
So $\sigma_{\length} =
     \deltacall{\length}{\pi_{12}}{\main}  \cup
\deltacall{\length}{\pi_6}{\length}$.
Filling in  the values gives:
\begin{eqnarray*}
\sigma_{\length}    &=&
 \{\epsilon\}  ~\cup~{\clazy\sigma_{\mathit{\length}}}
\end{eqnarray*}
As examples, the liveness of \Lanv{1}{\pl} and \Lanv{9}{\pa} in terms
of  $\sigma_{\length}$ are:
\begin{align*}
\Lanv{1}{\pl} &= (\clazy \cup \acdr\Df{\length}{1}\clazy
  \cup \clazy\Df{\length}{1}{\clazy})\sigma_{\mathit{\length}}
 \\
\Lanv{9}{\pa} &= \clazy \bcar \Df{\length}{1}\{\epsilon\}
\end{align*}

In summary, the equations generated during liveness analysis are:
\begin{enumerate}
\item For each function $\mathit{f}$, equations defining \Df{f}{i} for
  use by \Lfonly$_{\mathit f}$.
\item For each function $\mathit{f}$, an equation defining the summary
  demand $\sigma_{\mathit f}$ on $e_f$.
\item  For  each  function   $\mathit{f}$  (including  $\mainpgm$  for
  $e_\mainpgm$)  an equation  defining liveness  at each  GC point  of
  $e_{\mathit f}$.
\end{enumerate}
\subsection{Solving liveness equations---the grammar
interpretation}\label{sec:grammar-formulation}
The  equations above can now be re-interpreted as a
context-free grammar (CFG)  on the alphabet $\lbrace\acar, \acdr,
\bcar, \bcdr, \clazy\rbrace$.  Let  \var{$X$} denote the non-terminal
for
a variable  $X$ occurring on  the LHS of the  equations generated
from the analysis.  We can  think of the resulting productions as
being associated with several  grammars, one for each non-terminal
\var{\Lanv{i}{x}} regarded as a start symbol.  As an example, the
grammar    for   \var{\Lanv{1}{\xl}}   comprises    the   following
productions:
\begin{eqnarray*}
  \var{\Lanv{1}{\xl}}  &\rightarrow& 
  \clazy\var{\sigma_{\length}} \mid \acdr \var{\Df {\length}{1}}{\clazy\var{\sigma_{\length}}}  \\
  & & \mid
  \clazy\var{\Df{\length}{1}}{\clazy\var{\sigma_{\length}}} \\
  \var{\Df{\length}{1}} &\rightarrow& \clazy \mid
  \acdr\var{\Df{\length}{1}}\clazy
       \mid \clazy\var{\Df{\length}{1}}{\clazy}\\
\langle {\sigma_{\length}} \rangle
&\rightarrow&
\epsilon  \mid \clazy\langle{\sigma_{\length}}\rangle \\
\var{\Lanv{9}{\pa}} &\rightarrow& \clazy \bcar \var{\Df{\length}{1}}
\end{eqnarray*}
Other equations can be converted similarly.  The language generated by
\var{\Lanv{i}{x}},  denoted  $\mathscr{L}(\var{\Lanv{i}{x}})$, is  the
desired solution  of \Lanv{i}{x}.  However,  note that the  strings in
the  language are  over alphabet  $\lbrace \acar,  \acdr,\bcar, \bcdr,
\clazy \rbrace$, but  we want {\em forward} access paths  only, i. e.,
access paths over the alphabet $\lbrace \acar, \acdr \rbrace$. In
other words,  the decision  problem that we  are interested  in during
garbage collection at a program point $\pi_i$ is:
\begin{quote}
Let  $x.\alpha$ be  a forward  access  path consisting  only of  edges
\acar\    and   \acdr\    (but    not   \bcar, \bcdr\ or \clazy).     Let
$\mathscr{L}(\var{\Lanv{i}{x}}) \stackrel{*}{\hookrightarrow} \sigma$,
where $\sigma$ consists  of forward access paths only. Then  does $\alpha \in
\sigma$?
\end{quote}

We now model the above problem  as one of deciding the membership of a
context-free  grammar  augmented  with  a fixed  set  of  unrestricted
productions.

\begin{definition}\label{def:specialgrammar}
Consider the  grammar $(N,T, p_1\cup p_2,S)$ in which $N$ is a set of
non-terminals, $T = \{\acar, \acdr, \bcar, \bcdr, \clazy, \$\}$, $p_1$
is a set of context-free productions that contains the distinguished
production $S \rightarrow \alpha\$$, $\alpha$ is a string of grammar
symbols that does not contain $S$, and $p_2$ is the fixed set of
unrestricted productions $\bcar\acar \rightarrow \epsilon$,
$\bcdr\acdr \rightarrow \epsilon$, $\clazy\acar \rightarrow \clazy$,
$\clazy\acdr \rightarrow \clazy$, and $\clazy\$ \rightarrow
\epsilon$.
\end{definition}

From Sections \ref{sec:liveness}  and \ref{sec:computing}, it is clear
that the results of liveness analysis of any program can be modeled by
the kind  of grammar described above. The  following proposition shows
that the converse also holds.
\begin{proposition}
Given    a     grammar    $G$    of    the     form    described    in
Definition~\ref{def:specialgrammar},  it is  possible  to construct  a
program $p$  with program points  $\pi_i$ and variables $x$  such that
the liveness  analysis of $p$  is same as  $G$ except for a  change in
non-terminal names which now have the form \var{\Lanv{i}{x}}.
\end{proposition}
We now show that the decision  problem that is required to be answered
at       garbage      collection       time       is      undecidable.
\newcommand{\state}{\ensuremath{\mathsf{S}}}
\newcommand{\nont}[2]{\ensuremath{\mathsf{S}_{#1}^{#2}}}  
\begin{lemma}
Consider    a    grammar    $G$    of   the    kind    described    in
Definition~\ref{def:specialgrammar} and a forward access path $\alpha$
consisting  of symbols \acar\  and \acdr\  only. The  decision problem
$\alpha \in \mathscr{L}(G)$ is undecidable.
\end{lemma}

\begin{proof}
Given a  Turing machine  and an input  $w\in (1+0)^*$, we  construct a
grammar $G$  such that  the machine will  halt on  $w$ if and  only if
$\epsilon \in  \mathscr{L}(G)$. The grammar includes the  fixed set of
unrestricted productions  in Definition~\ref{def:specialgrammar}.

We  shall   denote  the  configuration   of  the  Turing   machine  as
$w_l(\state,c)w_r$, where $w_l$ is the string to the left of the head,
$w_r$ is the string to the right, $c$ is the symbol under the head and
\state\ is the current state  of the machine.  For each combination of
state  and   symbol  $(\state,c)$,   the  grammar  will   contain  the
non-terminal \nont{}{c}. We shall synchronize each move of the machine
to a  derivation step  using a context  free production,  followed, if
possible, by  a derivation step  using either $\bcar\acar    \rightarrow
\epsilon$ or $\bcdr\acdr    \rightarrow
\epsilon$.
After each synchronization, we shall establish the following invariant
relation between the machine configuration and the  sentential form:

\begin{quote}
  If the configuration of the machine is $w_l(\state,c)w_r$, then the
  sentential form will be $\overline{w}_l\nont{}{c}\,w_r$, where
  $\overline{w}_l$ is the same as $w_l$ but with each symbol $d$ in $w_l$
  replaced by $\overline{d}$.
\end{quote}

Assume that  the TM starts  in a  state $S_\mathit{init}$ with  a tape
$cw$ and  the head positioned on  the symbol $c$. Then  the sentential
form    corresponding     to    the    initial     configuration    is
$\nont{\mathit{init}}{c}w$ (we  can assume that there  is a production
$\nont{}{} \rightarrow  \nont{\mathit{init}}{c}w$, where  \nont{}{} is
the start symbol of the  grammar). Further correspondences between the
Turing machine moves and the grammar productions are as follows:

\begin{enumerate}
\item For each transition  $(S_i, c) \rightarrow (S_j,c',L)$, there are
  two  productions $\nont{i}{c}  \rightarrow  \acar \nont{j}{\acar}c'$
  and $\nont{i}{c} \rightarrow \acdr \nont{j}{\acdr}c'$.
\item For each  transition $(S_i, c) \rightarrow (S_j,c',R)$, there
  are two productions
  $\nont{i}{c} \rightarrow c \nont{j}{\acar}\bcar$ and $\nont{i}{c}
  \rightarrow c \nont{j}{\acdr}\bcdr$. 
\end{enumerate}
The idea behind  the productions is explained with  an example: Assume
that        the        current        sentential        form        is
\bcar\bcdr\nont{i}{\acar}\acar\acar.  Assume that  the  machine has  a
transition  $(S_i,\acar) \rightarrow (S_j,\acdr,  L)$. Since  the next
corresponding step in the derivation  has to be done without any prior
knowledge of whether the symbol to the left of the tape is a \acar\ or
a  \acdr, two  productions are  provided,  and the  invariant will  be
maintained  only   if  the  production   $\nont{i}{\acar}  \rightarrow
\acdr\nont{j}{\acdr}\acdr$  is  chosen  for   the  next  step  in  the
derivation.         This         gives        the        configuration
\bcar\bcdr\acdr\nont{i}{\acdr}\acdr\acar\acar.    Simplification  with
the    production    $\bcdr\acdr    \rightarrow    \epsilon$    yields
\bcar\nont{i}{\acdr}\acdr\acar\acar, which  exactly corresponds to the
changed configuration  of the machine.  Notice carefully that  a wrong
choice breaks the invariant and it cannot be recovered subsequently by
any choice of productions.

After  the Turing machine  has halted,  there are  further ``cleanup''
derivations  that derive  $\epsilon$ only  if the  invariant  has been
maintained  so far.  For every  symbol  $c$, we introduce  a  non-terminal
$\nont{\mathit{final}}{c}$ where $\nont{\mathit{final}}{}$ is the
final state of the Turing machine. We add
productions $\nont{\mathit{final}}{c} \rightarrow \acar
\nont{\mathit{final}}{c}$ and $\nont{\mathit{final}}{c} \rightarrow \acdr
\nont{\mathit{final}}{c}$ for cleaning up the \bcar\ and \bcdr\ symbols on
the left of the head and $\nont{\mathit{final}}{c} \rightarrow 
\nont{\mathit{final}}{c}\bcar$ and $\nont{\mathit{final}}{c} \rightarrow 
\nont{\mathit{final}}{c}\bcdr$ for cleaning up \acar\ and \acdr\ on the
right of the tape head. This completes the reduction.
\end{proof}

We circumvent the problem  of undecidability by over- approximating the
CFG by non-deterministic finite state automata (NFA). The NFAs are then
simplified using  the $\hookrightarrow$ rules.  Finally the simplified
NFAs are converted to DFAs.

%------------------------------------------------------------%

\begin{figure}[t!]
\scalebox{0.8}{
  \begin{tabular}{cc}
\psset{unit=1mm,nodesep=0mm,labelsep=0.5mm}
\begin{pspicture}(0,-5)(50,12)
  \putnode{t0}{origin}{5}{0}{\var{\Df{\length}{1}}}
  \putnode{t1}{t0}{15}{0}{\pscirclebox{\mbox{\ \ \ \ }}} \hspace{5mm}
  \putnode{t2}{t1}{15}{0}{\pscirclebox[doubleline=true]{\mbox{\ \ \ \ }}}
  \psset{arrows=->}
  \ncline{t0}{t1}
  \ncline{t1}{t2}
  \putnode{l0}{t1}{7}{2}{\clazy}
  \nccurve[angleA=45, angleB=135, ncurv=4, nodesep=-1]{t1}{t1}
  \putnode{l1}{t1}{0}{8}{\acdr, \clazy}
  \nccurve[angleA=45, angleB=135, ncurv=3, nodesep=-1]{t2}{t2}
  \putnode{l2}{t2}{0}{8}{\clazy}
\end{pspicture}
&
\psset{unit=1mm,nodesep=0mm,labelsep=0.5mm}
\begin{pspicture}(0,-5)(30,12)
  \putnode{t0}{origin}{5}{0}{\var{\Df{\length}{1}}}
  \putnode{t1}{t0}{15}{0}{\pscirclebox[doubleline=true]{\mbox{\ \ \ \ }}}
  \psset{arrows=->}
  \ncline{t0}{t1}
  \nccurve[angleA=45, angleB=135, ncurv=3, nodesep=-1]{t1}{t1}
  \putnode{l1}{t1}{0}{8}{\acdr}
\end{pspicture}
\\
(a) & (b) \\
\psset{unit=1mm,nodesep=0mm,labelsep=0.5mm}
\begin{pspicture}(0,-5)(50,12) %\psframe(0,-5)(40,12)
  \putnode{t0}{origin}{5}{0}{\var{\Lanv{9}{\pa}}}
  \putnode{t1}{t0}{9}{0}{\pscirclebox{\mbox{\ \ \ \ }}} \hspace{5mm}
  \putnode{t2}{t1}{12}{0}{\pscirclebox{\mbox{\ \ \ \ }}}
%  \putnode{t3}{t2}{12}{0}{\pscirclebox[doubleline=true]{\mbox{\ \ \ \ }}}
  \psset{arrows=->}
  \ncline{t0}{t1}
  \ncline{t1}{t2}
  \putnode{l0}{t1}{5}{2}{\clazy}
%  \ncline{t2}{t3}
%  \nccurve[angleA=55, angleB=125, ncurv=3, nodesep=-.8]{t3}{t3}
%  \putnode{l2}{tD1}{0}{10}{\acdr}
 \putnode{tD1}{t2}{12}{0}{\pscirclebox{\mbox{\ \ \ \ }}} \hspace{5mm}
  \putnode{tD2}{tD1}{12}{0}{\pscirclebox[doubleline=true]{\mbox{\ \ \ \ }}}
  \psset{arrows=->}
  \ncline{t2}{tD1}
  \putnode{l1}{t2}{5}{2}{\bcar}
  \ncline{tD1}{tD2}
  \putnode{lD0}{tD1}{5}{2}{\clazy}
  \nccurve[angleA=45, angleB=135, ncurv=4, nodesep=-1]{tD1}{tD1}
  \putnode{lD1}{tD1}{0}{8}{\acdr, \clazy}
  \nccurve[angleA=45, angleB=135, ncurv=3, nodesep=-1]{tD2}{tD2}
  \putnode{lD2}{tD2}{0}{8}{\clazy}
\end{pspicture}
&
\psset{unit=1mm,nodesep=0mm,labelsep=0.5mm}
\begin{pspicture}(0,-5)(20,12) %\psframe(0,-5)(40,12)
  \putnode{t0}{origin}{5}{0}{\var{\Lanv{9}{\pa}}}
  \putnode{t1}{t0}{9}{0}{\pscirclebox{\mbox{\ \ \ \ }}} \hspace{5mm}
  \psset{arrows=->}
  \ncline{t0}{t1}
\end{pspicture}
\\
(c) & (d)  \end{tabular}}
\caption{(a) The  grammar rules for \var{\Df{\length}{1}}
  converted into an automaton, and (b) its DFA. The same for \var{\Lanv{9}{\pa}}
  are in (c) and (d).}\label{fig:example-automata}
\figrule
\end{figure}

\subsubsection{From CFGs to GC-ready DFA}
\label{sec:NFA-approx}
 We use  the algorithm by Mohri  and Nederhof~\cite{mohri00regular} to
 approximate a CFG by a {\em strongly regular\/} grammar.  For
 example, the grammar
 fragment  for  the  non-terminal  $\var{\Df{\length}{1}}$  after  the
 Mohri-Nederhof transformation is:
 \begin{eqnarray*}
   \var{\Df{\length}{1}} &\rightarrow& \clazy\var{\Df{\length}{1'}} \mid
   \acdr\var{\Df{\length}{1}}
   \mid \clazy\var{\Df{\length}{1}}\\
   \var{\Df{\length}{1'}} &\rightarrow& \clazy\var{\Df{\length}{1'}}
   \mid \epsilon
 \end{eqnarray*}

The strongly regular grammar is converted  into a set of NFAs, one for
each $\var{\Lanv{i}{x}}$.  The $\hookrightarrow$ simplification is now
done on the NFAs by  repeatedly introducing $\epsilon$ edges to bypass
pairs  of consecutive  edges labeled  \bcar\acar\ or  \bcdr\acdr\ and
constructing the $\epsilon$-closure. Since the $\epsilon$-closure step
may  create  more  adjacent  \bcar\acar\ and  \bcdr\acdr\  edges,  the
simplification is continued till a fixed point is reached, after which
the edges labeled \bcar \ and  \bcdr\ are deleted.  The details of the
algorithm, its correctness and termination proofs are given by Karkare
et.  al.~\cite{karkare07liveness,asati14lgc}.  The resulting automaton
has  edges labeled  with  \acar,  \acdr\ and  \clazy\  only. In  this
automaton, for every edge labeled \clazy, we check if the source node
of the edge has a valid path to  a final final state.  If yes, we mark
the mark  the source node as  final. Finally, we remove  all the edges
labeled  \clazy\  and  convert  the automaton  into  a  deterministic
automaton.    This   effectively  implements   the   $\hookrightarrow$
simplification rules for  \bcar, \bcdr, and \clazy\  to obtain forward
paths. While  checking for liveness  during GC, a  forward access-path
is valid only if it can reach a final state.
Figure~\ref{fig:example-automata}(a)  shows the  NFA that  is obtained
from the grammar for \var{\Df{\length}{1}}.  The final DFA is shown in
Figure~\ref{fig:example-automata}(b).  This expectedly says that given
a  demand $\epsilon$  on  \length,  the liveness  of  its argument  is
$\acdr^{*}$  (the  spine  of   the  list  is  traversed).   Similarly,
Figure~\ref{fig:example-automata}(c)     shows     the     NFA     for
\var{\Lanv{9}{\pa}}.          The         DFA         shown         in
Figure~\ref{fig:example-automata}(d)  does  not   accept  any  forward
paths.   This  reflects  the  lazy  nature  of  our  language.   Since
\length\ does not  evaluate the member elements of  the argument list,
the  closure for  \pa\ is  never evaluated  and is  reclaimed whenever
liveness based GC is triggered beyond $\pi_9$.

\section{The Garbage Collection Scheme}
\label{sec:GC-scheme}
Our experimental setup consists of  an interpreter for our language, a
liveness  analyzer, and  a single  generation copying  collector.  The
garbage collector  can be  easily configured to  work on the  basis of
reachability (RGC  mode) or use  the liveness DFA (LGC  mode).  While
exploring the  activation record of a  function, LGC uses  the DFA at
the latest program  point traversed in the function's  body.  Thus, if
LGC is invoked  at the program point $\pi$, it  would use the liveness
DFA  at  $\pi$ itself  for  exploring the  root  set  of the  current
activation record. For  any other activation record in  the stack, say
for a function ${\mathit f}$  calling ${\mathit g}$ in the call-chain,
the evaluation  point in ${\mathit f}$  which resulted in  the call to
${\mathit g}$ would be used.

We shall call  a unit allocatable memory as a {\em  cell}.  A cell can
hold  a  basic  value   ($\mathit{bas}$),  the  constructor  \CONS\  $(\mathit{cons~
arg_1~arg_2}$) or   a closure.  The closure, in  turn, can be  one of
($\mathit{unop~arg_1}$),  ($\mathit{ binop~arg_1~arg_2}$)  and  function  application
($\mathit{app~arg_1~arg_2})$.  Here each $\mathit{ arg_i}$ is a reference to another heap
cell or terminated through the function $\mathit{makenull}$.

While exploring the heap, the GC  may encounter closures. In RGC mode,
the closure  will be explored  as usual  and copied by  traversing the
references corresponding to the arguments. Since our liveness analysis
predicts the  liveness of  data in its  evaluated form  only, closures
present a challenge for LGC. LGC will  need to map the liveness of the
free variables of the closure to  their evaluation point.  This is not
straight  forward as  variables may  escape their  creation scope  and
need to  be handled in an  approximate way.

\subsection{A liveness-based garbage collection scheme}
\label{sec:live-clo}
  \SetStartEndCondition{ }{}{}%
  \SetKwProg{Fn}{procedure}{\string:}{}
  \SetKwFunction{Range}{range}%%
  \SetKw{KwTo}{in}\SetKwFor{For}{for}{\string:}{}%
  \SetKwIF{If}{ElseIf}{Else}{if}{:}{elif}{else:}{}%
  \SetKwFor{While}{while}{:}{fintq}%
  \AlgoDontDisplayBlockMarkers\SetAlgoNoEnd\SetAlgoNoLine%
  \SetKwFunction{Lgc}{$lgc$}%
  \SetKwFunction{Copy}{$lgcCopy$}%
  \SetKwFunction{LCopy}{$lgcCopy$}%
  \SetKwFunction{RCopy}{$rgcCopy$}%

\begin{algorithm}[t]
  \Fn{\Lgc{}}
     {
       \For {each reference $\mathit{ref}$ in root set}
            {\Copy($\mathit{ref}$)\;}
            $\mathit{copyReferencesOnPrintStack()}$\;  
     }

     \Fn{\Copy{$\mathit{ref}$}}
        {
          \If{$\mathit{ref}$ is found live using dfa}
             {
               $\mathit{newRef} = $  $\mathit{dupHeapCell(ref})$\;
               \If{cellType($\mathit{ref}$) is $cons$}
                  {
                    {
                      $\mathit{newCar}  = $  \Copy($\mathit{ref\!.car})$\;
                      $\mathit{newRef.car}   =  newCar$\;
                    }
                    {
                      $\mathit{newCdr}$ =   \Copy($\mathit{ref.cdr})$\;
                      $\mathit{newRef.cdr}   = \mathit{newCdr}$\;
                    }    
                  }
               \If {cellType($\mathit{ref}$) is closure}
                   {
                     $\mathit{newRef.arg1}$ =  \Copy$\mathit{(cell.arg1)}$\;
                     $\mathit{newRef.arg2}$ =  \Copy$\mathit{(cell.arg2)}$\;
                   }
             }
        }

        \caption{LGC with closure collection based on liveness.\label{algo:lgc-a}}
\end{algorithm}
Algorithm~\ref{algo:lgc-a}   describes    a   liveness-based   garbage
collection scheme.  Starting with the root set, each live reference is
explored  using   $\mathit{  lgcCopy}$.    Copying  \CONS\   cells  is
simple---it  just involves  copying  the cell  itself and  recursively
copying the  \CAR\ and the  \CDR\ fields.   On the other  hand copying
closures  requires  determining  the  liveness  of  its  arguments  at
run-time.

Let   us  consider   determining  the   liveness  of   a  closure   at
runtime. Consider a closure $(app~f~y)$  has a reference from the root
set  variable  $x$.  If $x$  is  live,  then  we  know that  the  cell
$(app~f~y)$ has  to be copied.  To copy  the heap structure  rooted at
$y$, we need the liveness of  $y$ at the program point which triggered
the evaluation of  $(app~f~y)$. However, this is only  possible if the
activation record of the function  which created this closure is still
on the activation stack. This does  not always happen since a function
could pack a closure in a \CONS\ cell and return the \CONS\ cell.

An alternative could be to  carry with each closure, extra information
about the program point which triggered its evaluation.  This gives us
precise liveness  information of the  free variables of a  closure but
involves the overhead of updating the program point information during
execution.  To avoid  this overhead we go for a  safe approximation by
storing the {\em creation point of  the closure} and using it to check
liveness.  This  is safe since the  liveness at the creation  point of
the closure dominates  all possible evaluation points  of the closure,
but it  might be imprecise  as the actual  liveness might be  way less
than what is found at the creation point.

Note that  when a reachability-based  collector visits a cell  for the
first time,  it is  marked.  Explorations starting  from the  cell are
curtailed if the cell is visited subsequently due to sharing.  Since a
liveness-based collector may reach the  same cell multiple times, each
time with a different automaton state, this curtailment cannot be done
in the case  of LGC. This is an inherent  drawback of a liveness-based
collector.  This  can be mitigated  by maintaining a list  of liveness
automata states  that a  particular heap cell  was traversed  with and
avoid repeated traversals if we reach the cell in one of these states.
However this also involves storage and run-time penalty.

To  avoid  performance  penalty,  we experimented  with  a  mixed-mode
garbage collection scheme  which uses liveness for  evaluated data and
reachability for closures.

\subsection{Liveness-based collector with reachability-based closure
  collection}
\begin{algorithm}[t]
  \Fn{\Lgc{}}
     {
       \For {each reference $\mathit{ref}$ in root set}
            {\LCopy($\mathit{ref}$)\;}
            $\mathit{copyReferencesOnPrintStack()}$\;  
     }
     \Fn{\LCopy{$\mathit{ref}$}}
        {
          \eIf{cellType$(\mathit{ref})$ is $cons$}
              {
                \eIf {$\mathit{ref}$ is found live using dfa and has not been copied using
                  reachability}
                     {
                       $\mathit{newRef}$ =  $\mathit{dupHeapcell(ref})$\; 
                       $\mathit{ newCar}  =$  \LCopy$\mathit{(ref.car)}$\;
                       $\mathit{ newRef.car}   = \mathit{ newCar}$\;
                       $\mathit{ newCdr}  =$  \LCopy$\mathit{(ref.cdr)}$\;
                       $\mathit{ newRef.cdr}   = \mathit{newCdr}$\;
                     }
                     { $\mathit{makenull(ref)}$\;}}
              {
                \If {cellType$\mathit{(ref)}$ is closure and 
                  has not been copied}
                    { 
                      $\mathit{newRef}$  = \RCopy$(\mathit{ref})$\;
                    }
              }
        }
        \caption{LGC with reachability-based closure collection.\label{algo:lgc-b}}
\end{algorithm}

\begin{table}[ht!]
\caption{Time taken by LGC for different strategies.  The programs are
  modified (smaller) instances of original benchmarks.}
\label{tab:compare-live-clo-reach-clo}
\centering
\begin{tabular}
{| l | r | r | r |}
\hline
Program        & \#GC & \multicolumn{2}{c|}{Closure collected using} \\ \cline{3-4}
               &      & Reachability & Liveness                      \\ \hline
\verb@nqueens@ & 183  & 0.009 sec    & 0.228 sec                     \\\hline
\verb@lambda@  & 11   & 0.001 sec    & 4.992 sec                     \\ \hline
\verb@lcss@    & 1    & 0.000 sec    & 34.668 sec                    \\ \hline
\end{tabular}
\end{table}

Algorithm~\ref{algo:lgc-b} describes a garbage collection scheme where
the liveness automata is consulted  only for \CONS\ cells; and assumes
everything under  a closure to be  live and copies everything  that is
reachable.

A consequence of  copying everything under a closure is  that we might
reach  a non-live  cell that  has been  reclaimed earlier  by liveness
based GC.   This can  happen if  a \CONS\  cell having  partially live
sub-structure  is copied  using liveness  during a  collection and  it
subsequently  becomes part  of a  live closure.   During a  subsequent
collection, the collector may try to copy everything under the closure
including the dead  sub-structure of the \CONS\ cell.   To avoid this,
we ensure  that any non-live  references are not carried  across GC's.
We do this by scanning the live  part of the buffer after every GC and
nullifying any  reference which is  not live. This is  correct because
any reference which  is dead during a collection is  guaranteed not be
dereferenced any further by the program.

Another advantage of  this mixed garbage collection scheme  is that it
avoids the problem of multiple  explorations from the same cell.  This
is done by maintaining a flag in  each cell to indicate whether it was
copied using reachability.

\subsection{Garbage collection for references on print stack}
In lazy  languages the evaluation of  closures is driven  by the print
function~\cite{Jones87}.  Printing of  atomic values is
simple  and  does  not  trigger  a garbage  collection.   In  case  of
\CONS\ cells, the \CAR\ is first  printed followed by \CDR.  In a lazy
language, both \CAR\ and \CDR\ could be closures requiring evaluation,
and this  may trigger  a garbage collection.   When this  happens.  we
have to consider  any references that might be on  the print stack and
copy them. We extend the liveness analysis to the print function and
use its result during garbage collection.

\section{Experimental Evaluation}
\label{sec:experiments}
To   test  the   utility  of   our  method   we  have   implemented  a
garbage-collector for a simple scheme-like first order language having
non-strict semantics. It consists of an interpreter, liveness analyzer
and  a  garbage-collector  that  can optionally  use  reachability  or
liveness.  Our benchmark consists  of programs from nofib~\cite{nofib}
suite and some  external programs manually converted to  ANF.  All our
benchmarks were executed on a  machine having 8 core Intel(R) Core(TM)
i7-4770 3.40GHz  CPU ,  8192 KB  L2 cache,  16 GB  RAM running  64 bit
Ubuntu 14.04.

The process of liveness-based garbage collection involves going through 
each activation record on the stack and exploring each variable in the 
current activation record (root set). The liveness of each variable is 
determined using the program point and the variable name. The liveness 
of all variables are stored as DFA. These DFA are then encoded as a 
table. A variable is live only if an entry is found in this table.
All cells that are live are copied to the live semi-space. In case of 
\CONS\ cells the \CAR\ and \CDR\ pointers are chased and if they are 
live the copied \CONS\ cell will get the updated addresses of its 
\CAR\ and \CDR\ fields.

\subsection{Copying closure using liveness vs copying closures using reachability}\label{sec:strategies}

To compare the  effect of using reachability to  copy closures instead
of liveness  we take some  representative programs from  our benchmark
suite    and    execute    them    with    both.     As    shown    in
Table~\ref{tab:compare-live-clo-reach-clo},  the   number  of  garbage
collections in both cases are same but copying closures using liveness
takes a lot more time compared to reachability based copying.

The number  of garbage  collections is the same  since we  approximate the
liveness of  a closure  to the  liveness at  its creation  point.  The
liveness  we get  is usually  grossly over  approximated.  Using  this
liveness may not  give too much of a benefit  compared to reachability
as nearly everything would be live at its point of creation.

As  mentioned   in  Section~\ref{sec:live-clo},  using   liveness  for
closures  may  entail  multiple  traversals over  the  same  structure
consuming more  time.  The  advantage of avoiding  multiple traversals
can be clearly seen from the lcss example where a single GC takes less
than a millisecond for reachability  based closure collection where as
the  liveness  based collection  is  of  the  order of  seconds.  Lazy
languages tend  to have a  lot of sharing  and hence using  a liveness
based collection for closures is impractical.

Therefore we believe  that using reachability for closures  is a sweet
spot   in   terms   of   performance  for   liveness   based   garbage
collectors.  For the  remainder  of  the section  all  LGC results  we
discuss uses  a mixed-mode collector--liveness for  evaluated data and
reachability for closures.

\begin{table*}[ht!]
\caption{Statistics for liveness analysis and garbage collection}
\label{tab:exp-results}
\centering
 \begin{tabular}{|l||c@{\ }|@{\ }c@{\ }|@{\ }c@{\ }|@{\ }c@{\ }|@{\ }c@{\ }|@{\ }c@{\ }|@{\ }c@{\ }|@{\ }c@{\ }|@{\ }c|} \hline
   \multicolumn{1}{|r||}{Program} & \verb@lambda@ & \verb@nperm@ & \verb@treejoin@ & \verb@lcss@ & \verb@sudoku@ & \verb@fibheap@ & \verb@nqueens@ & \verb@knightstour@ & \verb@gc_bench@ \\ \hline\hline
   \#CFG Nonterminals            & 1078          & 794          & 709             & 839         & 1758          & 780            & 529            & 659                & 417             \\ \hline
   \#CFG Rules                   & 1823          & 1038         & 1610            & 1752        & 2509          & 1358           & 748            & 909                & 486             \\  \hline 
   \#DFA States                  & 3312          & 2055         & 2332            & 2340        & 5672          & 2254           & 1229           & 1742               & 729             \\\hline
   \#DFA Transitions             & 5792          & 3632         & 4235            & 4261        & 10857         & 4076           & 2078           & 3101               & 1124            \\\hline
   DFA Generation Time (sec)      & 14.7          & 1.1          & 2690.5          & 13.3        & 798.5         & 49.8           & 0.5            & 5.1                & 0.1             \\\hline
 \end{tabular}\\
 (a) Data for Liveness Analysis \\[1pt]
\include{autotable}
\vskip -5mm
\end{table*}

\subsection{Results}

The statistics related to compile  time liveness analysis are shown in
Table~\ref{tab:exp-results}(a). We observe that the analysis requires
a reasonable time except for  \verb@treejoin@ and  \verb@sudoku@. 
The bottleneck in our analysis is the NFA to DFA conversion
(worst-case exponential behaviour). To avoid this compile-time
overhead, we have added a caching mechanism in our simulator that can
reuse the DFA from earlier analysis if the program has not changed..

Table~\ref{tab:exp-results}(b)   compares   the   garbage   collection
statistics  for RGC  and  LGC.  We  report the  number  of GC  events,
average  number of  cells reclaimed  per GC,  average number  of cells
touched  per GC,  total  number of  dragged cells  and  total time  to
perform  all collections.   It is  no  surprise that  number of  cells
reclaimed per  GC is  higher (and  consequently the  number of  GCs is
lower)  for LGC  based collection.   LGC  also reduces  the number  of
dragged cells significantly. However, the cost  for LGC is that per GC
execution  time is  more  for  LGC, which  results  in higher  overall
execution time  for garbage  collection, even  with reduced  number of
collections.   However, LGC  run time  is still  competitive (slowdown
within 5X of RGC for worst case), and is better for 3 benchmarks (best
case  2X speedup).   Note  that  \verb@gc_bench@~\cite{gc_bench} is  a
synthetic benchmark  that allocates complete binary  trees of various
sizes.  These binary trees are never used by the program. As a result,
the benchmark highly favours LGC.  We have included this benchmark for
completeness  only  and   the  numbers  for  it   are  not  considered
representative of practical programs.

\newcommand{\hgt}{2.9cm}
\begin{figure*}[p]
\renewcommand{\arraystretch}{.1}
\begin{tabular}{@{}c@{}@{}c@{}}
   \hskip -4mm{\epsfig{file=lambda.eps, height=\hgt}}
   & {\epsfig{file=lambda_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=nperm.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=nperm_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=treejoin.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=treejoin_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=lcss.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=lcss_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=sudoku.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=sudoku_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=fibheap.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=fibheap_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=nqueens.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=nqueens_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=knightstour.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=knightstour_win.eps, height=\hgt}}
\\ (a) & (b)
\end{tabular}%\vskip -10mm
 \caption{Memory usage.  Column (a) shows complete  usage, column
   (b) enlarges  a part  of the  usage.  The blue  and the  red curves
   indicate the number of cons cells  in the active semi-space for RGC
   and LGC  respectively.  The  black curve  represents the  number of
   reachable cells and  the lightblue curve represents the  number of cells
   that are  actually live (of  which liveness analysis does  a static
   approximation).   x-axis   is  the  time  measured   in  number  of
   cons-cells allocated (scaled down by  factor $10^5$). y-axis is the
   number of cons-cells (scaled down by  $10^3$).}
\label{fig:memory-usage} \figrule
\end{figure*}

Memory   usage   graphs   for   select   benchmarks    are   shown
Figure~\ref{fig:memory-usage}, Column  (a).  As the number  of garbage
collections tend to  be very large, for each benchmark  we also show a
window which  is representative  of the  behavior for  that particular
benchmark (Column (b)).  In all the programs we can see that the curve
corresponding to  LGC (red  line) regularly dips  below the  RGC curve
(blue line), the dip is specially visible for the benchmarks where LGC
outperforms  RGC substantially.   The  graphs also  include curve  for
reachable cells (black,  obtained approximately by forcing  RGC to run
at a very high frequency) and the live cells for particular run (lightblue,
obtained by post processing the heap access at the end of a program).

One area  of concern is the  huge gap between the  actual liveness and
the liveness  perceived by our  collector.  In  case of LGC  for eager
languages~\cite{asati14lgc} the gap was  very narrow and almost touched
the actual  liveness curve.  In  case of  lazy languages, this  gap is
largely   due  to   approximate  reachability   based  collection   of
closures. To implement liveness based  collection of closures, we need
to  record liveness  data  for  each closure  and  update  it at  each
evaluation  point.   These  space  and  run  time  overheads  make  it
infeasible  to  incorporate  liveness  based  GC  for  closures  in  a
practical garbage collector.

\section{Related Work} 
\label{sec:relatedwork}
Although augmenting  garbage collection with liveness  information has
been  studied earlier,  it was  mainly  in the  context of  imperative
languages~\cite{Albert, Hirzel,  khedker07heap}.  Due to  the presence
of global variables and mutation, practical utility of simple liveness
based techniques are found to be inefficient.

In the space  of functional languages, improving  space efficiency has
been  mainly  studied as  a  compile  time activity.   Either  through
rewriting                methods                such                as
deforestation~\cite{wadler88deforest,gill93ashort,chitil99deforest},
sharing  analysis   based  reallocation~\cite{jones89compile},  region
based analysis~\cite{tofte98region},  or through insertion  of compile
time              nullifying              statements              such
as~\cite{inoue88analysis,lee05static,Hamilton}.    The  compile   time
marking approaches all rely on an efficient and precise alias analysis
and in the  absence of it cannot provide  notable improvement. Another
important approach  advocated by  Hofmann~\cite{HofmannJ03} is  to use
linear  typing to  analyze  first-order programs  for  heap usage  and
annotated  them. The  system  then uses  these  annotations to  re-use
allocated memory  cells instead  of requesting  for newer  cells. This
requires the  user to write programs  in a specific way  and hence may
not be practical

Simplifiers~\cite{ONeill}, described as lightweight daemons, 
improve the efficiency of the program in general and the garbage
collector in particular. Most of the simplifications mentioned in this
work are subsumed by a liveness  based collector. We feel this work is
orthogonal and can augment our  liveness based collector.  The closest
to our  approach is the  liveness based garbage  collector implemented
in~\cite{asati14lgc}. We  extend their work to  handle lazy evaluation
and closures.

%% A similar approach has been suggested in~\cite{ONeill} where they suggest augmenting garbage collection with extra information using {\it simplifiers}.

%% Previous  attempts  to increase  the  space  efficiency of  functional
%% programs  by  additional  reclamation  of memory  fall  in  two  broad
%% categories. In the first, the program itself is instrumented to manage
%% reclamation and reallocation without the aid of the garbage collector.
%% Such      attempts      include:      sharing      analysis      based
%% reallocation~\cite{jones89compile},                      deforestation
%% techniques~\cite{wadler88deforest,gill93ashort,chitil99deforest},
%% methods  based  on   linear  logic~\cite{hofmann00linear}  and  region
%% analysis~\cite{tofte98region}.   Closer  to  our approach,  there  are
%% methods   that  enable   the   garbage  collector   to  collect   more
%% garbage~\cite{inoue88analysis,lee05static}  by  explicitly  nullifying
%% pointers  that are  not  live.  However,  the  nullification, done  at
%% compile  time,  requires sharing  (alias)  analysis.   Our method,  in
%% contrast, does not require sharing  because of the availability of the
%% heap itself  at runtime.  To  the best of  our knowledge, this  is the
%% first attempt  at liveness-based  marking of  the heap  during garbage
%% collection.

%% \begin{figure}[t]
%% \centerline{\epsfig{file=lambda.eps, height=4cm, width=7cm}}
%%  \caption{Memory  usage and garbage collection pattern of  {\tt
%% lambda}.}
%% \label{fig:memory-usage-lambda} \figrule
%% \end{figure}
\vskip -5mm
\section{Conclusions and Future Work}
\label{sec:conclusion}
%% We  have  defined  a  notion  of liveness  on  structured  data;  this
%% generalizes classical liveness and strong liveness.  We started with a
%% general fully context-sensitive analysis  which we proved correct with
%% respect to  a minefield semantics  (this models the effect  of garbage
%% collection between every evaluation step).

%% To  avoid scalability  issues (and  to  avoid performing  part of  the
%% liveness computation at run time)  we defined an 0-CFA version of this
%% liveness analysis  in which  demands for function  $f$ at  all calling
%% contexts are conflated into  a single demand $\sigma_f$.  This enabled
%% us to treat the liveness equations symbolically obtaining context-free
%% grammars for liveness at each GC point (calls to user functions and to
%% $\CONS$).  These were then converted to DFA for run-time consultation
%% by  the garbage collector.  Experiments confirm  the precision  of the
%% analysis.

%% To obtain performance figures we compared a reachability-based garbage
%% collector with a liveness-based  collector.  This showed a decrease in
%% the  number  of  GCs,   more  garbage  collected  per  invocation.   A
%% significant benefit of LGC is  that programs can run in smaller memory
%% when compared to  RGC. This is potentially useful  in situations where
%% memory  is  limited---as with  embedded  systems.  For  a majority  of
%% programs, the garbage collection times were reduced.

%% One issue we  highlighted was that while fewer  nodes were marked (and
%% more garbage collected), sometimes  cons cells could be visited and
%% traversed  multiple times  with different  sets of  liveness  paths to
%% explore.  Further work includes  improvements to the classical copying
%% collector to reduce the cost of this.
We extended the liveness based garbage collection to lazy
  languages and shown its benefit  for garbage collection in practical
  programs. We  defined a context  sensitive liveness analysis  that uses
  context independent summaries of functions obtained using a symbolic
  demand. We showed  that obtaining precise solution  to these equations
  is undecidable in  our formulation, and hence  we safely approximate
  the result using DFA. These DFA are consulted by the garbage collector
  to improve  collection.

  Liveness   of   closures   presented  further   challenges   to   our
  implementation.   We compared  different approximate  strategies for
  handling   liveness   of   closures,   and  found   that   a   mixed
  strategy---reachability  based  collection   of  closure  arguments,
  liveness  based  collection  of   everything  else---works  best  in
  practice as it  avoids runtime and space  overheads.  With mixed-mode
  garbage collection scheme,  we were able to collect  more cells than
  reachability based collectors at a reasonable overhead.

  Although we  provide an  implementation for  a liveness  based garbage
  collection scheme for lazy language, we  do not provide a formal proof
  of its correctness.  A formal  proof of correctness would describe how
  closures  are  handled  during  garbage  collection  and  liveness  is
  propagated inside closures.

  Another interesting exercise is to narrow the gap between the actual
  liveness and  the perceived liveness  of our garbage  collector. Our
  experiments (not  reported here) show  that a significant  number of
  dead cells  get trapped inside closures.   Using strictness analysis
  to eagerly evaluate closures might  release more of these dead cells
  to be  garbage collected.   Since our  garbage collector  can handle
  closures  (suspended evaluations),  another  major  challenge is  to
  extend it to support higher order programs.

  Orthogonally,  we plan  to improve  the efficiency  of the  liveness
  based garbage collector using heuristics  such as limiting the depth
  of  DFA,   merging  nearly-equivalent   states  and   using  better
  representation  and algorithms  for automata  manipulation. We  also
  need  to   investigate  the  interaction  of   liveness  with  other
  collection schemes, such as incremental and generational collection.
  In  summary, we  need to  investigate  ways to  make liveness  based
  garbage collection attractive for practical collectors.

%\balance
\bibliography{fun_hra}{}
\bibliographystyle{abbrv}
\end{document}