### 2016-01-28 paper/paper.tex
towards 10 page limit

\documentclass[9pt]{sigplanconf}
\usepackage{etex}
\usepackage{float}
\usepackage{mathrsfs}
%%\usepackage{lineno}
\usepackage{booktabs}
\usepackage{stmaryrd}
\usepackage{boxedminipage}
\usepackage[T1]{fontenc}
\usepackage{epsfig}
\usepackage{multirow}
\usepackage{url}
\usepackage[normalem]{ulem}
\usepackage{schemepgm}
\usepackage{graphicx}
\usepackage{graphpap}
\usepackage{tabularx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{pstricks}
\usepackage{pst-text}
\usepackage{pst-node}
\usepackage{pst-tree}
\usepackage{pst-rel-points}
\usepackage{bcprules}
\usepackage{subfigure}
\usepackage[boxed]{algorithm2e}
\usepackage{makecell}
\usepackage{balance}

\input{fh_defs}
\def\drawplusplus#1#2#3{\hbox to 0pt{\hbox to #1{\hfill\vrule height #3 depth
      0pt width #2\hfill\vrule height #3 depth 0pt width #2\hfill
      }}\vbox to #3{\vfill\hrule height #2 depth 0pt width
      #1 \vfill}}
      %Poor man's typography
\def\concat{\mathrel{\drawplusplus {12pt}{0.4pt}{5pt}}}
      %It would be better to specify these in font-relative measures, but it
      %probably doesn't scale anyway.
%\newcommand{\mycomment}[1]{} 
\definecolor{Myblue}{rgb}{.2,0,1}
\definecolor{Myred}{rgb}{1,.2,.2}

\newcommand{\comment}[1]{{\color{Myblue}{#1}}}
\newcommand{\warning}[1]{{\color{Myred}{#1}}}
%\newcommand{\blankout}[1]{}
\newcommand{\oxlb}{\ensuremath{[\!|}}
\newcommand{\oxrb}{\ensuremath{|\!]}}
\newcommand{\cred}[1]{\psframebox[linestyle=none, fillcolor=lightgray,fillstyle=solid,framesep=0.5pt]{#1}}
%\newcommand{\deltacall}[3]{\delta_{#1}({#2},{#3})}
\newcommand{\scmin} {\mbox{\sf\em in}}
\newcommand{\scmnull}{\mbox{\sf\em null?}}
\newcommand{\scmpair}{\mbox{\sf\em pair?}}
\newcommand{\scmprim}{\ensuremath{\mathsf{+}}}
\newcommand{\sh}[1]{{\colorbox{gray!20}{\framebox{$#1$}}}}
\def\myvec{\mathaccent"017E } 
\newcommand{\stk}{{S}}      
\newcommand{\ID}{\mbox{$\mathbf{ id}$}}
\newcommand{\bang}{\mbox{\sc bang}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{lemma}[theorem]{Lemma}
\begin{document}
%% \linenumbers 
% --- Author Metadata here ---
%\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
%\CopyrightYear{2007} % Allows default copyright year (20XX) to be
%over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data
%(0-89791-88-6/97/05) to be over-ridden - IF NEED BE. 
% --- End of Author Metadata ---

\title{Liveness-Based Garbage Collection for Lazy Languages}

% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.
%\cmt{{
%
%\numberofauthors{3} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors
%section.
%
%\author{
% 1st. author
\authorinfo {Prasanna Kumar. K} 
       {IIT Bombay, Mumbai 400076, India}
       {prasannak@cse.iitb.ac.in}
% 2nd. author
\authorinfo {Amitabha Sanyal} 
       {IIT Bombay, Mumbai 400076, India}
       {as@cse.iitb.ac.in}
% 3rd. author
\authorinfo {Amey Karkare}
       {IIT Kanpur, Kanpur 208016, India}
       {karkare@cse.iitk.ac.in}
% 3rd. author
%% \alignauthor Amey Karkare\\
%%        \affaddr{IIT Kanpur}\\
%%        \affaddr{Kanpur 208016, India}
%%        \email{karkare@cse.iitk.ac.in}
%}
%}}
%\authorinfo{Double Blind Review}{}{}
\maketitle

\begin{abstract} 
We consider  the problem of reducing  the memory required to  run lazy
first-order functional programs. Our approach is to analyze
programs  for liveness  of  heap-allocated data.   The  result of  the
analysis is  used to preserve  only live data---a subset  of reachable
data---during garbage  collection.  The result  is an increase  in the
garbage reclaimed and  a  reduction in  the peak memory requirement 
of programs.   While this  technique has already  been shown  to yield
benefits for  eager first-order  languages, the  lack of  a statically
determinable execution  order and  the presence  of closures  pose new
challenges  for lazy  languages.  These  require changes  both in  the
liveness analysis itself and in the design of the garbage collector.

%% We use the  result of the analysis to annotate  each potential garbage
%% collection  point  in   the  program  with  a   set  of  deterministic
%% finite-state automata (DFA) describing the liveness at that point.
To  show the  effectiveness of  our method,  we implemented  a copying
collector that uses  the results of the liveness  analysis to preserve
live  objects,  both evaluated  (i.e.,  in  WHNF) and  closures.   Our
experiments confirm  that for  programs running with  a liveness-based
garbage  collector, there  is a  significant decrease  in peak  memory
requirements.   In addition,  a  sizable reduction  in  the number  of
collections  ensures that  in spite  of using  a more  complex garbage
collector, the execution  times of programs running  with liveness and
reachability-based collectors remain comparable.
\end{abstract}

\category{D.3.4}{Programming Languages}{Processors}[Memory Management
  (Garbage Collection), Optimizations]
\category{F.3.2}{Logic and Meanings Of Programs}{Semantics of
  Programming Languages}[Program Analysis]

\terms{Algorithms, Languages, Theory}

\keywords{Heap Analysis, Liveness Analysis, Memory Management, Garbage
  Collection, Lazy Languages}

\section{Introduction}
\label{sec:intro}

Functional  programs  make  extensive  use  of  dynamically  allocated
memory.  The allocation is either explicit (i. e., using constructors)
or  implicit   (creating  closures).   Programs  in   lazy  functional
languages put additional  demands on memory, as  they require closures
to be carried from the point of creation to the point of evaluation.

While  the runtime  system  of most  functional  languages includes  a
garbage collector to efficiently  reclaim memory, empirical studies on
Scheme~\cite{karkare06effectiveness}   and,   more   importantly,   on
Haskell~\cite{rojemo96lag} programs have shown that garbage collectors
leave uncollected a large number  of memory objects that are reachable
but not live (here {\em live} means the object can potentially be used
by the program at a  later stage).  This results in unnecessary memory
retention.

In this paper,  we propose the use of liveness  analysis of heap cells
for garbage collection in a lazy first-order functional language.  The
central notion in our analysis  is a generalization of liveness called
{\em  demand}---the  pattern  of  future  uses  of  the  value  of  an
expression.   The  analysis  has  two parts.   We  first  calculate  a
context-sensitive  summary   of  each   function  as  a   {\em  demand
  transformer} that transforms a {\em  symbolic} demand on its body to
demands  on its  arguments.   This  summary is  used  to step  through
function calls  during analysis.   The concrete  demand on  a function
body  is  obtained through  a  conservative  approximation similar  to
0-CFA~\cite{Shivers:1988} that  combines the demands on  all the calls
to  the function.   The result  of the  analysis is  an annotation  of
certain program points with  deterministic finite-state automata (DFA)
capturing the  liveness of  variables at  these points.   Depending on
where garbage collection is triggered, the collector consults a set of
automata to restrict reachability during  marking.  This results in an
increase in  the garbage reclaimed  and consequently in  fewer garbage
collections.

While this idea has been shown  to be effective for a first-order {\em
  eager}  language~\cite{asati14lgc}, a  straightforward extension  of
the technique is not possible for lazy languages, where heap-allocated
objects may  include closures (runtime representations  of unevaluated
expressions).   Firstly, since  data  is made  live  by evaluation  of
closures, and  in lazy languages the  place in the program  where this
evaluation  takes  place  cannot be  statically  determined,  laziness
complicates  liveness analysis  itself.  Moreover,  for liveness-based
garbage collection to  be effective, we need to extend  it to closures
apart from  evaluated data.  Since a  closure can escape the  scope in
which it was created, it has  to carry the liveness information of its
free variables. As  execution progresses and possible  future uses are
eliminated, we  update the  liveness information in  a closure  with a
more precise version.  Because of these reasons, the garbage collector
also  becomes significantly  more  complicated  than a  liveness-based
collector for a eager language.

Experiments    with   a    single    generation   copying    collector
(Section~\ref{sec:experiments})   confirm  the   expected  performance
benefits.  Liveness-based collection results in an increase in garbage
reclaimed.  As  a consequence, there is  a reduction in the  number of
collections from 1.6X  to 23X and a decrease  in the minimum
memory requirement from  1X to 1389X.  As  an added benefit,
there is also a reduction in the overall execution time in 5 out of 12
benchmark programs.

%------------------------------------------------------------%
\subsection{Motivating example}
\label{sec:motiv}
\input{fig_motiv}   Figure~\ref{fig:mot-example}   shows  an   example
program and the state of the heap at the program point $\pi$\footnote{We write
$\pi\!:\!e$ to  associate a  label $\pi$ with  the program  point just
before  the  expression  $e$. The label is not part of the syntax.},  i.e. just  before  the  evaluation  of
$(\length\ \pz)$.  The heap is represented  by a graph in which a node
either  represents  atomic  values   ($\NIL$,  integers  etc.),  or  a
\CONS\  cell  containing  $\CAR$  and  $\CDR$  fields,  or  a  closure
(represented  by  shaded  clouds).   Edges   in  the  graph  are  {\em
  references} and represent values of variables or fields.  The figure
shows the  lists \px\  and \pz\  partially evaluated  due to  the
\SIF\ condition 
(\NULLQ~(\CAR~\pz)).
The edges  shown by thick  arrows are those  which are live  at $\pi$.
%% Only those  cell which have a future use  should be preserved
%% during garbage collection, all other
%% cells can be  reclaimed.
Thus if a garbage collection takes  place at $\pi$ with the heap shown
in Figure~\ref{fig:mot-example}(b),  a liveness-based  collector (LGC)
will preserve only  the cells referenced by $\pz$, and  the live cells
constituting the  closure referenced by $(\CDR~\pz)$.   In contrast, a
reachability-based collector (RGC) will preserve all cells.

In  this work  we show  that  static analysis  of heap  data can  help
garbage  collectors   in  reclaiming   more  garbage.    The  specific
contributions of this paper are
\begin{myitemize}
\item  We  formulate  a  liveness  analysis for  a  lazy  first-order
  functional language and  prove its correctness. The 
  proof involves specifying liveness through a non-standard semantics  and then proving the analysis 
  correct with respect to the specification.   
\item The analysis results  in a set  of context-free  grammars along
  with  a fixed  set  of non-context-free  productions.  The  decision
  whether   to  copy  a  cell   during  garbage  collection
  translates  to a  membership problem  for such  grammars.  In  this
  paper, we  have shown the  undecidability of this problem and 
  circumvent the  problem by over-approximating the  grammars
  with DFAs.
\item We  have implemented a garbage  collector that uses the  DFAs to
  retain live  cells.  Our  experiments reveal  interesting space-time
  trade-offs  in  the  engineering  of  the  collector---for  example,
  updating liveness  information carried in closures  during execution
  results  in   more  garbage  being  collected.    Empirical  results
  show the effectiveness of liveness-based garbage collection.
\end{myitemize}
As far  as we know, this  is the first work  on liveness-based garbage
collection      for      lazy       languages.       All      previous
attempts~\cite{shaham01heap,   ran.shaham-sas03,   shaham02estimating,
  asati14lgc, karkare06effectiveness} have  involved either imperative
or eager functional languages.

\subsection{Organization of the paper}

Section~\ref{sec:defs}  introduces  the   syntax  of  the  programming
language considered  and gives a small-step  operational semantics for
it.  The liveness  analysis for this language and  its soundness proof
is          presented          in          Section~\ref{sec:liveness}.
Section~\ref{sec:computing} describes  the formulation of  liveness as
grammars. We also give a proof  of undecidability of such grammars and
show     how     they     can     be     approximated     by     DFAs.
Section~\ref{sec:GC-scheme}   discusses   details   of   the   garbage
collector,  in  particular  the  use  of  liveness  DFAs  for  garbage
collection.     We     report    our    experimental     results    in
Section~\ref{sec:experiments}    along    with   some    observations.
Section~\ref{sec:relatedwork}  discusses  previous   work  related  to
garbage  collection  and   liveness  and  Section~\ref{sec:conclusion}
discusses possible extensions and concludes the paper.
\section{The target language---syntax and semantics}
\label{sec:defs}
Figure~\ref{fig:lang-syntax} describes the syntax  of our language. It
is a first order language with lazy semantics. Programs are restricted
to        be        in        Administrative        Normal        Form
(ANF)~\cite{chakravarty03perspective}, where  all actual  parameters to
functions  are  variables.  While  this  restriction  does not  affect
expressibility,  this form  has  the benefit  of  making explicit  the
creation of closures through the $\LET$ construct.   We
  assume that  $\LET$s  in  our  language  are  non-recursive;  in  the
  expression $\LET\,\, x  \leftarrow s\,\, \IN\,\, e$,  $x$ should not
  occur in  $s$. The restriction of  \LET\ to a single  definition is
for ease  of exposition---generalization to multiple  definitions does
not add conceptual difficulties.  We further restrict each variable in
a program  to be  distinct, so that  no scope  shadowing occurs; this
simplifies reasoning.

We denote the body of a function ${\mathit  f}$   as $e_{\mathit f}$.
 We assume that each program has a distinguished function
\mainpgm,    defined as    $(\DEFINE\      ({\tt
  \mainpgm})\  e_\mainpgm)$,  and  the program begins execution
with the call to \mainpgm.

\begin{figure}[t]\footnotesize
\renewcommand{\arraystretch}{1}
\begin{eqnarray*}
   p \in \mathit{Prog} & \!\!\!::=\!\!\! & d_1 \ldots d_n \,\,\,\, e_\mainpgm
   \hspace{5em} \,\,\,\,\,\,\,\,\,\; \mbox{\em --- program}\\
    \mathit{df} \in Fdef & \!\!\!::=\!\!\! & (\DEFINE\,\, (f\,\, x_1 \,\, \ldots
\,\,x_n)\,\,
    e)
    \hspace{0.2em} \ \ \ \ \ \ \ \ \  \mbox{\em --- function def} \\
e \in \mathit{Expr} & \!\!\!::=\!\!\! &
\left\{\begin{array}{@{}ll@{\hspace{1em}}l}
       (\SIF\,\, x\,\, e_1\,\, e_2) && \!\!\!\mbox{\em --- conditional} \\
       (\LET\,\, x \leftarrow s\,\, \IN\,\, e) &&\!\!\! \mbox{\em --- let
binding} \\
       (\SRETURN\,\, x) && \!\!\!\mbox{\em --- return from function}
    \end{array}\right. \\
s \in \mathit{App} & \!\!\!::=\!\!\!  &
\left\{\begin{array}{@{}l@{\hspace{1.2em}}l}
       k & \mbox{\em --- constant (numeric or $\NIL$)}\\
       (\CONS\,\, x_1\,\, x_2) & \mbox{\em --- constructor} \\
       (\CAR\,\, x) & \\
       (\CDR\,\, x) & \mbox{\em --- selectors} \\
       (\NULLQ\,\, x) & \\
       (\PRIM\,\, x_1\,\, x_2) & \mbox{\em ---  tester/generic arithmetic} \\
%%       (\ID\,\, x) & \mbox{\em ---  identity function (for inlining)} \\
       (f\,\, x_1\,\,\ldots\,\, x_n) & \mbox{\em --- function application}
    \end{array}\right.
\end{eqnarray*}
\vspace*{-6pt}
  \caption{The syntax of our language}\label{fig:lang-syntax}
%\figrule
\normalsize
\vspace*{-6pt}
\end{figure}

\begin{figure*}[t!]
%\comment{

\begin{center}\footnotesize
\renewcommand{\arraystretch}{1.5}
\begin{tabular}{|c|c|c|}
\hline
Premise & Transition & Rule name \\
\hline
\hline 
          & $\rho,\, (\rho', \ell, e)\!:\!S,\, \heap,\, \kappa
  \rightsquigarrow \rho',\, S,\, \heap[\ell := \kappa],\, e$    &  {\sc const}
\\
\hline
          & {$\rho, \,(\rho', \ell, e)\!:\!S,\, \heap,\, (\CONS~x~y)
\rightsquigarrow
$  $\rho',\, S,\, \heap[\ell := (\rho(x), \rho(y))],\, e$}     &  {\sc cons} \\
\hline

$\heap(\rho(x)) \mbox{ is } (v, d)$ & $\rho,\, (\rho', \ell, e)\!:\!S,\, \heap,\,
(\CAR~x)  \rightsquigarrow \rho',\, S,\, \heap[\ell := v],\, e$      &
{\sc car-select} \\
\hline

$\heap(\rho(x)) \mbox{ is } (\langle s, \rho'\rangle, d)$ & $\rho,\, S,\,  \heap,\,
(\CAR~x)  \rightsquigarrow \rho', \,(\rho, addr(\langle s, \rho'\rangle), (\CAR~x))\!:\!S,\, \heap,\, s$      &
{\sc car-1-clo} \\
\hline

$\heap(\rho(x)) \mbox{ is } \langle s, \rho'\rangle$ & $\rho,\, S,\, \heap,\, (\CAR~x)
\rightsquigarrow
\rho',\, (\rho, \rho(x), (\CAR~x))\!:\!S, \,\heap,\, s$      &
{\sc car-clo}
\\
\hline

$\heap(\rho(x)), \heap(\rho(y)) \in \mathbb{N}$
 & {$\rho,\, (\rho', \ell, e)\!:\!S,\, \heap,\, (+~x~y)  \rightsquigarrow$
$\rho', \,S,\, \heap[\ell := \heap(\rho'(x)) + \heap(\rho'(y))],\, e$}      &
{\sc prim-add} \\
\hline

$\heap(\rho(x)) \mbox{ is } \langle s, \rho'\rangle$ & $\rho,\, S,\, \heap,\, (+~x~y)
\rightsquigarrow
\rho',\, (\rho, \rho(x), (+~x~y))\!:\!S,\, \heap,\, s$      &
{\sc prim-1-clo} \\
\hline
$\heap(\rho(y)) \mbox{ is } \langle s, \rho'\rangle $ & $\rho,\, S,\, \heap,\, (+~x~y)
\rightsquigarrow
\rho', (\rho,\, \rho(y),\, (+~x~y))\!:\!S,\, \heap,\, s$      &
{\sc prim-2-clo} \\
\hline
{$\mathit{f}~\mbox{defined as}$
$~(\DEFINE~(f~\myvec{y})~e_{\mathit{f}})$}  & $\rho,\, S,\, \heap,\,
(f~\myvec{x})  \rightsquigarrow
[\myvec{y} \mapsto \rho(\myvec{x})],\, S,\, \heap,\, e_{\mathit{f}}$      &
{\sc funcall} \\
\hline
$\ell$ is a new location& {$\rho,\, S,\, \heap,\, (\LET~x\leftarrow s~\IN~e)
  \rightsquigarrow$
$\rho\oplus[x \mapsto \ell],\, S,\, \heap[\ell := \langle s,
    \lfloor\rho\rfloor_{FV(s)} %% \oplus [x \mapsto  \ell] -- NO LAZY LETS
    \rangle],\, e$} &
{\sc let} \\
\hline
$\heap(\rho(x)) \ne 0$ & $\rho,\, S,\, \heap,\, (\SIF~x~e_1~e_2)   \rightsquigarrow
\rho,\, S,\, \heap,\,  e_1$ & {\sc if-true} \\
\hline
$\heap(\rho(x)) = 0$ & $\rho,\, S,\, \heap,\, (\SIF~x~e_1~e_2)   \rightsquigarrow
\rho,\, S, \,\heap, \, e_2$ & {\sc if-false} \\
\hline
$\heap(\rho(x)) = \langle s, \rho' \rangle $ & {$\rho,\, S,\, \heap,\,
  (\SIF~x~e_1~e_2)   \rightsquigarrow
\rho',\, (\rho, \rho(x), (\SIF~x~e_1~e_2))\!:\!S,\, \heap, \, s$}
&
{\sc if-clo} \\
\hline
{$\heap(\rho(x))~\mbox{is in WHNF with value}~v$}& $\rho,\, (\rho', \ell,
e)\!:\!S,\, \heap,\,
(\SRETURN~x)  \rightsquigarrow \rho',\, S,\, \heap[\ell := v],\, e$ &
{\sc return-whnf}\\
\hline
$\heap(\rho(x)) = \langle s, \rho' \rangle $ & {$\rho,\, S,\, \heap,\, (\SRETURN~x)
  \rightsquigarrow$
$\rho',\, (\rho, \rho(x), (\SRETURN~x))\!:\!S,\, \heap,\,  s$} &
{\sc return-clo} \\
\hline
\end{tabular}
\caption{A small-step semantics for the language. \label{fig:lang-semantics}}
\end{center}
%}\comment
\vspace*{-6pt}
\end{figure*}

\subsection{Semantics}\label{sec:semantics}
%\comment{
We now give  a small-step semantics for our  language.
We first specify the domains used by the semantics:
\[
\renewcommand{\arraystretch}{1}
\begin{array}{@{}r@{\ }l@{\ \ }c@{\ \ }l@{\hspace{0.5em}}l}
\heap: & \mathit{Heap} & =&\mathit{Loc} \rightarrow (Data + \{empty\}) & \mbox{-- Heap}\\
d: & \mathit{Data} &=&\mathit{Val} + \mathit{Clo} & \mbox{-- Values \& Closures} \\
v:   & \mathit{Val} &=& \mathbb{N} + \{\NIL\} + \mathit{Data \times Data}& \mbox{-- Values}\\
c:   & \mathit{Clo} &=& \mathit{(App \times Env)}& \mbox{-- Closures}\\
\rho: & \mathit{Env} &=&\mathit{Var} \rightarrow \mathit{Loc} &
\mbox{-- Environment} \\
\end{array}
\]

Here  $\mathit{Loc}$ is  a countable  set  of locations  in the  heap.
A non-empty location either contains  a value in  \emph{WHNF} (a  number, the
empty  list  $\NIL$ or  a  \CONS\  cell  with  possibly  unevaluated
constituents) or  a {\em closure}.   A closure  is a pair  $\langle s,
\rho\rangle$ in  which $s$ is  an unevaluated application,  and $\rho$
maps free variables of $s$ to their respective locations. Since all data objects are boxed, we model an environment as a mapping
from the set  of variables of the program  $\mathit{Var}$ to locations
in the  heap.

The  semantics  of   expressions  (and  applications\footnote{In  most
 contexts, we  shall use the  term 'expression' and the  notation $e$
to stand   for both expressions and applications.}) 
are given by transitions of the form $\rho, \stk, \heap, e \rightsquigarrow
\rho',  \stk', \heap',  e'$.  Here  $\stk$ is  a stack  of continuation
frames.  Each  continuation frame  is a triple $(\rho,  \ell, e)$,
signifying that the  location $\ell$ has to be updated  with the value
of the  currently evaluating  expression and $e$  is to  be evaluated
next  in the  environment $\rho$.  
The  initial  state  of  the  transition  system  is:
\[([\;]_\rho,\,     (\rho_\mathit{init},     \,    \ell_{\ans}     ,\,
(\print~\ans)):[\;]_{S},   [\;]_{\heap},\,   (\mainpgm))\]  in   which
$[\;]_\rho$, $[\;]_{\mbox  \footnotesize {H}}$ and $[\;]_{S}$  are are
the empty environment, heap and  stack respectively. The initial stack
consists  of  a  single  continuation   frame  in  which  \ans\  is  a
distinguished variable that will eventually  be updated with the value
of  (\mainpgm),  and  $\rho_\mathit{init}$  maps \ans\  to  a  location
$\ell_{\ans}$.  In addition, \print\ is a function modeling a printing
mechanism---a   standard   run-time   support  assumption   for   lazy
languages~\cite{Jones87}---that prints  the value of  (\mainpgm).  The
operator~$:$ pushes elements on top of the stack.

The   notation  $[\myvec{x}   \mapsto  \myvec{\ell}]$   represents  an
environment  that  maps  variables  $x_i$ to  locations  $\ell_i$  and
$\heap[\ell := d]$  indicates updation of  \heap\ at $\ell$
with  $d$.   $\rho \oplus  \rho'$  represents  the environment  $\rho$
shadowed  by  $\rho'$  and  $\lfloor \rho  \rfloor_X$  represents  the
environment  restricted  to  the  variables in  $X$.  Finally  $FV(s)$
represents the  free variables  in the  application $s$  and $addr(c)$
gives the address of the closure $c$ in the heap.

The small-step semantics  is shown in Figure~\ref{fig:lang-semantics}.
Unlike   an   eager  language,   evaluation   of   a  let   expression
$(\LET~x\leftarrow  s~\IN~e)$ does  not  result in  the evaluation  of
$s$. Instead,  as the {\sc let}  rule shows, a closure  is created and
bound to  $x$. The  program points which  trigger evaluation  of these
closures   are  an $\SIF$   condition ({\sc  IF-clo})   and
 a \SRETURN\ ({\sc  return-clo}).  We  call such  points \emph{evaluation
  points   $(\epath)$}  and   label  them   with  $\psi$   instead  of
$\pi$. As an  example of closure  evaluation, we  explain the
three rules for  $(\CAR~x)$.  If $x$ is a closure,  it is evaluated to
WHNF, say $(d_1, d_2)$.  This is  given by the rule {\sc car-clo}.  If
$d_1$ is  not in WHNF,  it is  also evaluated ({\sc  car-1-clo}).  The
address  to be  updated  with the  evaluated value  is
recorded  in  a  continuation  frame.    This  is  required  for the
evaluation  to be lazy, else $d_1$ may  be evaluated more than once
due to  sharing~\cite{Jones87}.  Only after this  is the actual selection  done ({\sc
  car-select}).

\section{Liveness}
\label{sec:liveness}

A variable is {\em live} if there  is a possibility of its value being
used in  future computations and  dead if  it is definitely  not used.
Heap-allocated data  needs a richer model  than classical liveness---a
model which talks about liveness of references.

%% The model use the notion of  {\em access paths}, i.e.
%% a  prefix-closed set of strings  over $\{\acar,\acdr\}^\ast$, in
%% which 
%% $\acar$,  $\acdr$ represent  access  using $\CAR$  and $\CDR$  fields.

An {\em access path}
is a  prefix-closed set of strings  over $\{\acar,\acdr\}^\ast$, where
$\acar$,  $\acdr$ represent  access  using $\CAR$  and $\CDR$  fields respectively.
Given an initial location $\ell$ (usually a reference corresponding to
a variable)  and a  heap \heap, semantically  an access  path $\alpha$
represents     a      reference,     denoted     $\heap\llbracket\ell,
\alpha\rrbracket$,  in the  heap  that is  obtained  by starting  with
$\ell$ and chasing the \CAR\ or  \CDR\ fields in the heap as specified
by  the  access   path.   $\heap\llbracket\ell,  \alpha\rrbracket$  is
defined only if  the path followed in the  heap is \emph{closure-free}
(does not cross closures), else it is undefined.

Access paths are used to  represent liveness. As an example, a list
$x$ with 
liveness  $\{\epsilon,  \acar,  \acdr,  \acdr\acar,
\acdr\acdr,  \acdr\acdr\acar\}$ means that future computations    only
refer up  to the  second and  third members of  $x$.  A  {\em liveness
  environment} is a mapping from  variables to access paths, but often
expressed as  a set,  for example  by writing  $\{x.\epsilon, x.\acdr,
x.\acdr\acdr,        y.\epsilon\}$        instead        of        $[x
  \mapsto\{\epsilon,\acdr,\acdr\acdr\},          y\mapsto\{\epsilon\},
  z\mapsto\{\}]$.    In  this   notation,  $y   \mapsto  \{\epsilon\}$
represents access using $y$ itself and $z \mapsto \{\}$ indicates 
$z$ is dead.  In lazy  languages, liveness environments are associated
with regions of programs instead of program points.

A notion  that generalizes liveness  is {\em demand}.   While liveness
gives the patterns of future uses of a variable, demand represents the
future uses of the value of an expression.  The demand on an expression
$e$   is   again   a   set    of   access   paths---that   subset   of
$\{\acar,\acdr\}^\ast$ which the  context of $e$ may  explore of $e$'s
result.   To see  the need  for demands,  consider the expression
$(\LET~x\leftarrow  (\CDR~y)~\IN~   (\SRETURN~x))$.   Assume   that  the
context of this expression places  the demand $\lbrace \epsilon, \acar
\rbrace$. Since the  value of the expression is the  value of $x$, the
demand translates to the liveness  $[x \mapsto \lbrace \epsilon, \acar
  \rbrace]$.   Due  to  the  \LET\   definition  which  binds  $x$  to
$(\CDR~y)$, the liveness of $x$  now becomes the demand on $(\CDR~y)$.
This, in  turn, generates  the liveness $\lbrace  y.\epsilon, y.\acdr,
y.\acdr\acar  \rbrace$.   These are  the  set  of $y$-rooted  accesses
required to  explore $\lbrace  \epsilon, \acar  \rbrace$ paths  of the
result of $(\CDR~y)$.
%% As an  analogy with classical  (strong) liveness analysis,  $y$ and
%% $z$ are live at the entry $\pi: x:=y+z$, if and only if $x$ is live
%% at  exit  of $\pi$.   In  our  terminology, the  liveness  $\lbrace
%% \epsilon\rbrace $ of $x$ at the  exit from $\pi$ becomes the demand
%% on  $y+z$,  and  this,  in turn  generates  the  liveness  $\lbrace
%% y.\epsilon, z.\epsilon \rbrace$ at the entry of $\pi$.

We use $\sigma$  to range over demands, $\alpha$ to  range over access
paths and $\Lv$  to range over liveness environments.  The liveness of
an  individual variable  $y$ in  \Lv\  is $\Lv(y)$,  but more  commonly
written as  $\Lv_y$.  The notation $\sigma_1\sigma_2$  denotes the set
$\lbrace  \alpha_1\alpha_2 \mid  \alpha_1 \in  \sigma_1, \alpha_2  \in
\sigma_2\rbrace$.  Often we shall abuse  notation to juxtapose an edge
label and  a set  of access  paths; $\acar\sigma$  is a  shorthand for
$\lbrace\acar\rbrace\sigma$.

\input{example-program}
%%\label{fig:mot-example2}

%==============================================================
\renewcommand{\pp}[2]{\ensuremath{#1\!\!:\!#2}} % prog point
%==============================================================

\subsection{Liveness analysis for lazy languages}
\label{sec:liveness-analysis}

Consider the  program in Figure~\ref{fig:mot-example2-a}.   As mentioned
earlier, a lazy evaluation of the $\LET$ expression at $\pi_1$ creates
a closure  for $(\length~\px)$  instead of  evaluating it.   Since the
closure may escape the scope in which it is created, it carries a copy
of $\px$ within itself.  We treat the  copy of $\px$ in the closure as
being separate from the $\px$ introduced  by the $\LET$, and call it a
\emph{closure  variable}.  For  liveness calculations,  such variables
are distinguished  from variables  introduced by  \LET s  and function
arguments  (called \emph{stack  variables},  since they  reside in  the
activation  stack).   We notationally distinguish  a closure  variable
from  its corresponding  stack variable  by subscripting  it with  the
label of the program   point  where   the   closure  was   created\footnote{Multiple
  occurrences  of the  same  variable in  an  application are  further
  distinguished by their positions in the application.}.

  Since a  closure is evaluated  only at evaluation points,  a closure
  variable is attributed  with the same liveness in  the entire region
  of  the  program from  the  point  of  creation  of the  closure  to
  reachable evaluation points.  This is  also true of stack variables,
  because, as we shall see, stack variables derive their liveness from
  closure variables. Thus, there are two major differences between our
  formulation of  liveness of  lazy languages  with liveness  of eager
  languages~\cite{asati14lgc}:   (i)  the   introduction  of   closure
  variables in the  liveness calculations, and (ii)  a single liveness
  value for each  variable that is applicable from  its creation point
  to evaluation points.

\input{fig-liveness}

  Closure  variables get  their  liveness values  through  a chain  of
  dependences beginning at  a variable at an evaluation  point.  As an
  example, consider the following program:

\input{example-to-illustrate-liveness}

The variable  \pz\ returned  at $\psi_3$ depends  on \py\  through the
expression $(+~\py~1)$.  $\py$ in turn depends on the closure variable
$\px_{\mathit{\pi}_1}$  through $(\length~\px_{\mathit{\pi}_1})$.   We
denote this  chain of  dependences as  $\lbrack \psi_3\!\!\!:\!\!\!\pz
\leftarrow            (+~\py~1),\,           \py            \leftarrow
(\length~\px_{\mathit{\pi}_1})\rbrack$.  Indeed,  the chains of closures  in the
heap are a  runtime representation of these  dependences.  Since $\pz$
is evaluated  at $\psi_3$  due to  the expression  $\SRETURN~\pz$, the
demand made by the calling context(s)  of $f$ places a demand on $\pz$
which  will  impart a  liveness  to  $\px_{\pi_1}$.  Other  dependence
chains  which result  in  a liveness  for  $\px_{\pi_1}$ are  $\lbrack
\psi_1\!\!\!:\!\!\py   \leftarrow  (\length~\px_{\pi_1})\rbrack$   and
$\lbrack   \psi_2\!\!\!:\!\!\!\pw   \leftarrow  (/~\pu~\py),\,\,   \py
\leftarrow   (\length~\px_{\pi_1})\rbrack$.   The   liveness  analysis
described in this section declares the liveness of $\px_{\pi_1}$ to be
a union of  the liveness arising from these dependence  chains.  To be
safe, a garbage collection during  evaluation of $\py$ at $\psi_1$ has
to use  this liveness to  copy the heap starting from  $\px_{\pi_1}$.  However,
notice that if a garbage collection takes place while evaluating $\pz$
at $\psi_3$, it can safely consider only the liveness arising from the
dependence chain\linebreak  $\lbrack \psi_3\!\!\!:\!\!\!\pz \leftarrow
(+~\py~1),\, \py \leftarrow (\length~\px_{\pi_1})\rbrack$.  The garbage
collection  scheme  described  in Section~\ref{sec:GC-scheme}  uses  a
generalization of this observation  to dynamically select a evaluation
point specific liveness in order to collect more garbage.

Figure~\ref{fig:live-judge}  describes  our  analysis  which  has  two
parts. The  function $\mathit{ref}$,  takes an  application $s$  and a
demand $\sigma$ and returns the incremental liveness generated for the
free variables of $s$ due to  the application.  This will be consulted
during garbage collection  while exploring the heap  starting from the
closure  variables.  The function $\mathcal{L}$  uses  $\mathit{ref}$ to  propagate
liveness across expressions.

In  a   lazy  language,   an  expression   is  not   evaluated  unless
required. Therefore  the null  demand ($\emptyset$) does  not generate
liveness in any of the rules defining $\mathit{ref}$ or $\mathcal{L}$.
A non-null  demand of  $\sigma$ on (\CDR~$x$),  is transformed  to the
liveness $\{x.\epsilon,  x.\acdr\sigma\}$.  In an opposite  sense, the
demand  of  $\acdr\sigma$ on  (\CONS~$y$~$z$)  is  transformed to  the
demand  $\sigma$  on  $z$.   Since \CONS\  does  not  dereference  its
arguments, there  is no $\epsilon$ demand  on $y$ and $z$.   The rules
for (\PRIM~x~y) and (\NULLQ~x) are  similar. Constants do not generate
any liveness.

In case of a  function call, we  use the third  parameter $\Lfonly$
that  represents  the  summaries  of all  functions  in  the  program.
$\Lfonly_{\mathit  f}$  (the summary for a  specific
function $f$) expresses  how the demand $\sigma$ on a  call to $f$ is
transformed into  the liveness of  its parameters at the  beginning of
the  call.  $\Lfonly$  is determined  by the  judgement $\mathit{Prog}
\len  \Lfonly$ using  inference rule  ({\sc live-define}).   This rule
describes  the  fixed-point property  to  be  satisfied by  $\Lfonly$,
namely, the  demand transformation  assumed for  each function  in the
program should  be the  same as  the demand  transformation calculated
from      its      body.       As       we      shall      see      in
Section~\ref{sec:grammar-formulation},  we  convert  the rule  into  a
grammar and  the language generated  by this grammar is  the least
solution satisfying  the rule. We  prefer the least solution  since it
ensures the safe collection of the greatest amount of garbage.

We next  describe the function $\mathcal{L}$  that propagates liveness
across  expressions.   Consider  the $\mathcal{L}$-rules  for  {\LET},
{\SIF}, and {\SRETURN}.  Since the value of $(\SRETURN~x)$ is the
value  of  $x$,  a  demand $\sigma$  on  $(\SRETURN~x)$  gives  a
liveness   of  $\{x.\sigma\}$.    The  liveness   of  the   expression
$(\SIF~x~e_1~e_2)$  is  a union  of  the  liveness of  $e_1$  and
$e_2$. In  addition, since  the condition $x$  is also  evaluated, the
liveness $\{x.\epsilon\}$ is created and  added to the union.  
To  understand  the  liveness rule  for  $\LET~x  \leftarrow~s~\IN~e$,
observe that the value  of $\LET$ is the value of  its body $e$.  Thus
the  liveness environment  $\Lv$ of  $e$ is  calculated for  the given
demand $\sigma$. Since the stack variable $x$ is copied to each of the
closure variable  $x_{\pi}$, the liveness of  $x$ is the union  of the
liveness of the  closure variables.  This liveness,  say $\sigma'$, is
also  the  demand  on  $s$, thus the  liveness  environment
$\mathit{ref}(s,   \sigma',   \Lfonly)$   is  added   to   $\Lv   \cup
\{x.\sigma\}$. Finally, the stack variables corresponding to the free
variables of $s$ are updated and added 
  to  give the  overall liveness environment  for $(\LET~x
\leftarrow~s~\IN~e)$.

As noted earlier, $x.\alpha \in  \Lv_y$ specifies the liveness  of the
reference $\heap\llbracket\rho(x), \alpha\rrbracket$  only if $\alpha$ corresponds to  a closure-free path
in \heap\   starting from $\rho(x)$.  If this  path is intercepted  by a
closure,  say  $(\CAR~{x_{\pi}})$, then  the  liveness  of further  heap
contents starting from $\mathit{x_{\pi}}$ is given by $\Lv_\mathit{x_{\pi}}$.

%} %comment

\subsection{Soundness of analysis}

We shall now present a proof of soundness of the analysis presented in
Section~\ref{sec:liveness-analysis}.   Here are  the ideas  behind the
proof.
\begin{enumerate}
\item       We   augment     the        standard       semantics
  (Figure~\ref{fig:lang-semantics}) to model  a dynamic liveness analysis 
that records liveness values at each reference created during execution.
Any attempt to dereference a location with a null-liveness  
  during  a  transition will  result  in entering  a
  special state denoted \bang. We call the semantics after
  augmentation,  \emph{minefield semantics}.
\item \label{inline1} Assuming that a  program enters the \bang\ state,
  we  show  how to  construct,  through  inline expansion,  a  program
  without function calls which has the same minefield behavior.
\item  The final  step   shows  that  no program  without
  function calls can enter a \bang\ state. As a consequence no program
  (with or without function calls) can enter the \bang\ state. 
\end{enumerate}

\begin{figure*}[t!]
\begin{center}\footnotesize
\renewcommand{\arraystretch}{1.5}
\begin{tabular}{|@{}c@{\ }|@{\ }c@{\ }|@{\ }c@{\ }|}
\hline
Premise & Transition & Rule name \\ 
\hline

\hline
\makecell[t]{$\cred{GC(\rho_1, S, \heap_1, (\CAR~x), \sigma) = (\rho, S, \heap)}$, \\
  $\rho(x) \mbox{ is } \bot$} & $\rho, S,
  \heap, (\CAR~x), \cred{\sigma} \rightsquigarrow \bang$ &
{\sc car-bang} 
\\
\hline
\makecell[t]{$\cred{GC(\rho_1, S, \heap_1, (\CAR~x), \sigma) = (\rho,
    S, \heap)}$, \\$\heap(\rho(x)) \mbox{ is } (\langle s, \rho'\rangle, d)$} & \makecell[t]{$\rho,\, S,\,  \heap,\,
(\CAR~x), \cred{\sigma} \rightsquigarrow$ \\ $ \rho', \,(\rho, addr(\langle
s, \rho'\rangle), (\CAR~x),\cred{\sigma} )\!:\!S,\, \heap,\, s, \, \cred{\sigma}$ }     &
{\sc car-1-clo} 
\\
\hline
\makecell[t]{$\cred{GC(\rho_1, S, \heap_1, (\CAR~x), \sigma) = (\rho, S, \heap)}$,\\ $\heap(\rho(x))
\mbox{ is } \langle s, \rho'\rangle$} & $\rho, S,
  \heap, (\CAR~x), \cred{\sigma} \rightsquigarrow   \rho', (\rho, \rho(x),
  (\CAR~x), \cred{\sigma})\!:\!S, \heap, s, 
\cred{\renewcommand{\arraystretch}{1}\left\{
\begin{array}{lr}
  {\emptyset}&\mbox{if } \sigma = \emptyset \\       
  {\{\epsilon\} \cup \acar\sigma }&\mbox{otherwise}
\end{array}\right.
}$         
&
{\sc car-clo}\\
\hline
\end{tabular}
\end{center}
\caption{Minefield semantics for {\sc car-bang} and
  {\sc car-clo} \label{fig:minefield-semantics-for-some}}
\end{figure*}

To set up the minefield semantics, we follow these steps:
\begin{enumerate}

\item  We  extend  the  abstract  machine  state  $\rho,S,\heap,e$  to
  $\rho,S,\heap,e,\sigma$.   We call  such a  state a  \emph{minefield
  state}.  Here $\sigma$  is the ``dynamic'' demand  on the expression
  $e$, i.e., it arises from the actual sequence of function calls that
  led  to the  evaluation  of $e$. The demand for the initial state is  $\sigma_{all}$, and each expression $e$ before a $\rightsquigarrow$ step transforms the demand in accordance with the liveness rules of Section~\ref{sec:liveness-analysis} and passes it to the state after the step.  The  information in  continuation
  frames  on the  stack $S$  are also  similarly augmented  with their
  demands.  Thus a  stacked entry now takes the form  $(\rho, \ell, e,
  \sigma)$.
\item  $GC(\rho,  S,  \heap,  e, \sigma)$ models  a  liveness  based  garbage
  collection that returns $(\rho', S', \heap')$, where the changes in
  $\rho, S$ and  $\heap$ are due
  to certain references being replaced by $\bot$.  This simulates the
  act of  not retaining  (effectively garbage collecting)  the objects
  pointed to by these references  during an actual garbage collection.
  This is done in the following way:
  \begin{enumerate} 
  \item  \label{gc-1}Using the  current state  $\rho, \stk,  \heap, e,
    \sigma$  and the  demand  transformers $\Lfonly$,  we construct  a
    liveness environment  \Lv\ for  the expression $e$.   We similarly
    define the liveness environments for the expressions $e$ in each of
    the  continuation frames $(\rho,  \ell, e,  \sigma)$ in
    $S$, giving a stack of  liveness environments that we shall denote
    $\mathsf{SL}$.
  \item \emph{Garbage collection of objects pointed by root sets:} For
    the environment $\rho$ in the current state and for each $x \in
    domain(\rho)$, we  set $\rho(x)$ to $\bot$  iff $x.\epsilon \notin
    \Lv$.  Similarly,  for each  stack entry $(\rho,\ell,e,\sigma)$
    in $S$ 
    with  $\Lv'$   as  the   corresponding  liveness   environment  in
    $\mathsf{SL}$, and for each $x \in domain(\rho)$, we set $\rho(x)$
    to $\bot$ iff $x.\epsilon \notin \Lv'$.
  \item \emph{Garbage  collection of  objects in  the heap:}  For each
    location $\ell$,  we set $\heap(\ell)$  to $\bot$ iff for each
    environment $\rho$ in either  the  current state or in a
    continuation frame, for each $x \in domain(\rho)$ and each forward
    access
    path $\alpha$, it is not the case that $x.\alpha \in \Lv$ and $\heap\llbracket\rho(x), \alpha
    \rrbracket = \ell$. Here 
%% $x.\alpha$  is  live  and
%%     corresponds  to   $\ell$ in \heap.  More  formally,  there   are  no  $x\in
%%     domain(\rho)$ and  $\alpha$, such that  $\heap\llbracket\rho(x), \alpha
%%     \rrbracket = \ell$  and $x.\alpha \in \Lv$, where 
\Lv\ is the liveness environment  corresponding to $\rho$ mentioned in
step~\ref{gc-1}.
  \end{enumerate}
\end{enumerate}
Starting from the initial state,
%%  $([\;]_\rho,([\;]_\rho,
%% \ans, (\print~\ans)):[\;]_{S}  , [\;]_{\heap}, (\mainpgm)$,
every transition  is preceded  by a  garbage collection
using $GC$.  The details of the transition for the {\sc
  car-clo}  rule   is  shown   in
Figure~\ref{fig:minefield-semantics-for-some}.  If  this  or  an
earlier  call to  $GC$  results  in $\rho(x)$  being
bound  to  $\bot$,  then  the  $\rightsquigarrow$  step
enters the  \bang\ state ({\sc  car-bang}). Otherwise
the transition is similar to the earlier {\sc car-clo}
rule.

Consider a trace  of a minefield execution of a  program $p$, possibly
ending in  a \bang\ state. We  can construct a call-tree  based on the
trace in  which each node represents  a function that was  called (but
did not necessarily  return because of a \bang).  Assume  that each of
the nodes of  the tree is also annotated with  the program point where
the corresponding call  was invoked.  This tree can be  used to inline
function calls in a hierarchical fashion.  The details of the inlining
can be found in~\cite{asati14lgc}.
For  a
call-less program,  the initial state of  the minefield
semantics  is  assumed  to  be  $([\;]_\rho,([\;]_\rho,
\ans,      (\print~\ans)):[\;]_{S}     ,      [\;]_{\heap},
e_{\mainpgm}$.

\subsection{Soundness result}
We first need an auxiliary  result about minefield semantics. Consider
a trace of  a minefield execution.  For every  minefield state $(\rho,
S,   \heap,  e(s),   \sigma)$   that   appears  on   the   LHS  of   a
$\rightsquigarrow$ step, the demand $\sigma$ on the expression $e$ (or
application $s$) is  non-null.  This can be proved by  an induction on
the number  of steps leading  to the  minefield state.  The  base step
holds   because  the   demand   $\sigma_{\mathit{all}}$   on  $(\mainpgm)$   is
non-null. For the inductive step we  observe that for each step of the
minefield semantics, if the demand $\sigma$  on the LHS of a minefield
step  is non-null,  the  demand  on the  RHS  is  a transformation  of
$\sigma$ (for example $(\clazy\cup\acar)\sigma$) which is also non-null.

Note that our proofs will be for a single round of minefield execution
i.e.  the evaluation of $(\mainpgm)$ (or $e_{\mainpgm}$, for call-less
programs)  to   its  WHNF  driven  by   the  printing  mechanism~(Section
\ref{sec:semantics}).  With  minor variations,  the proof will  also be
applicable for subsequent evaluations initiated by $\print$.
%% We show  next that the  minefield execution of a  call-less program
%% cannot go \bang.

\begin{lemma}
\label{lemma:call-less-cannot-go-bang}
Consider the minefield execution of  a program without function calls.
Such a program cannot enter the \bang\ state.
\end{lemma}
\begin{proof}
Call a $GC(\ldots)$  followed by a $\rightsquigarrow$  transition as a
step. Consider a state $(\rho, S,  \heap, e, \sigma)$ in the minefield
execution of a  program.  We show by induction on  the number of steps
$n$ leading  to this  state that  the next  transition cannot  enter a
\bang\ state.  When $n$ is 0,  the state is $([\;]_\rho,\,    (\rho_\mathit{init},   \, \ell_{\ans}  ,\,
(\print~\ans)):[\;]_{S}, 
[\;]_{\heap},\, (\mainpgm))$.   Since  the
            call to $GC$  in this state does  nothing, we just
            have to show that the $\rightsquigarrow$ transition cannot
            enter  a \bang\  state.  Since  our programs  are in  ANF,
            $e_{\mainpgm}$ can  only be  a $\LET$ expression.   A {\sc
              let}  step  does  not involve  dereferencing,  and  thus
            cannot result in a \bang.

For the inductive step, we shall show that none of the minefield steps
that involves dereferencing  results in a \bang.  These  are the steps
which   have   a   $\heap(\rho(...))$   in   the   premise.    Now   a
$\rightsquigarrow$ step can go \bang\ because it dereferences a $\bot$
inserted  by  the  immediately  preceding garbage  collection  or  the
garbage collection of an earlier minefield step.  However the
  demand $\sigma'$ on basis of  which an earlier collection would have
  inserted a $\bot$, would have  included the current demand $\sigma$.
  Thus, if  the earlier garbage  collection had inserted a  $\bot$, so
  would the current garbage collection.  Thus it is enough to consider
  the $\rightsquigarrow$  along with the  current GC and show  that it
  does not lead to a \bang.

We consider the rules {\sc car-clo} and {\sc car-bang} only.  
The rest  of the rules involve similar reasoning. For the {\sc
  car-clo}  rule in  the state  $\rho, S,  \heap, (\CAR~x),
\sigma$, we  know that $\sigma$ is  non-null. Therefore
the  liveness  of  $x$  includes  $\epsilon$,  and  the
dereferencing $\heap(\rho(x))$ will go without \bang.

For the  {\sc car-bang}  rule, observe that  there are
two dereferences.  First $x$  is dereferenced to  get a
cons  cell  and then  the  head  of  the cons  cell  is
dereferenced  to  obtain  a  closure.   If  the  demand
$\sigma$ on  $(\CAR~x)$ is non-null, then  the liveness
of    $x$   will    include    both   $\epsilon$    and
$\acar\epsilon$,  and  a  garbage collection with  this  liveness  will
neither bind $x$ to a  $\bot$, nor insert $\bot$ at the
first   component  of   the   cons   cell.  Thus   both
dereferences    can take place without 
$\rightsquigarrow$ entering the   \bang\ state.
\end{proof}

Now we are ready to prove the main soundness result.

\begin{theorem}
The  minefield  execution of  no  program  can enter  a
\bang\ state.
\end{theorem}

\begin{proof}
Assume to  the contrary that  a program $P$  enters the
\bang\  state.  We can  transform  $P$  to a  call-less
program $P'$ such that  the minefield executions of $P$
and $P'$  are identical  except for change  of variable
names.             However,                 by
Lemma~\ref{lemma:call-less-cannot-go-bang} we know that
$P'$,   a   call-less   program,   cannot   enter   the
\bang\  state.  Therefore   $P$  also  cannot  enter  the
\bang\ state.
\end{proof}

%% Section~\ref{sec:computing}   shows   how  the   demand   transformers
%% $\Lfonly$  for  a  program  (representing  a  fully  context-sensitive
%% analysis) can be safely approximated  by a {\em procedure summary} for
%% each function.   The summary is  in the  form of a  demand transformer
%% that maps a demand on a call to the function to demands on each of its
%% arguments.

\section{Computing liveness and its encoding as DFA}\label{sec:computing}
Section~\ref{sec:liveness}  gave   a  fully context-sensitive
liveness analysis which 
%$\mathcal{L}$ and $ref$ together
described the  liveness set for  each variable  in a function  body in
terms of  a symbolic demand  $\sigma$ and \Lfonly.  However,  we are yet
to describe  (i) how to obtain demand  transformers \Lfonly\ from
the  rule {\sc  live-define}, and  (ii)  how to  compute the  concrete
demand $\sigma$  on each function.   To do so,  we first need to  modify the
liveness rules to a different form.

\subsection{Modifying liveness rules}

The   $\mathit{ref}$   rule   for   \CONS,   shown   in
Figure~\ref{fig:live-judge}, requires us  to remove the
leading  \acar\ and  \acdr\  from the  access paths  in
$\sigma$.  Similarly, the rules  for \CAR, \CDR, \PRIM,
\NULLQ, and \SIF\ require  us to return $\emptyset$, if
$\sigma$      itself       is      $\emptyset$      and
$\lbrace\epsilon\rbrace$  otherwise.  To  realize these
rules  $\sigma$   needs  to  be  known.   This  creates
difficulties  since  we  want to  solve  the  equations
arising from liveness symbolically.

The solution is to  also treat the operations mentioned
above  symbolically.  We  introduce three  new symbols:
\bcar, \bcdr,  \clazy.  These symbols are  defined as a
relation  $\hookrightarrow$  between   sets  of  access
paths:
\begin{align*}
  &\bcar\sigma \hookrightarrow \sigma' \mbox{ where } \sigma' = \{\alpha \mid \acar\alpha \in \sigma\}\\
  &\bcdr\sigma \hookrightarrow \sigma' \mbox{ where } \sigma' = \{\alpha \mid \acdr\alpha \in \sigma\}
\end{align*}
Thus \bcar\ selects those entries in $\sigma$ that have leading \acar, and removes the leading \acar\ from them.
The symbol \clazy\ reduces the set of strings following it to a set containing only $\epsilon$. It filters out, however, the empty set of strings.
\begin{align*}
  \clazy\sigma \hookrightarrow & \left\{ 
  \begin{array}{ll}
    \emptyset&\mbox{if}~\sigma = \emptyset\\
    \{\epsilon\} & \mbox{otherwise}
  \end{array}\right.
\end{align*}
We can  now rewrite the \CONS\  and the \CAR\  rules of $\mathit{ref}$
as:
\begin{align*}
&\mathit{ref\/}((\CONS~x~y),\sigma,\Lfonly)
= x.\bcar\sigma \cup y.\bcdr\sigma  \label{eqn:mod-cons},~
\mbox{and} \\
&\mathit{ref\/}((\CAR~x),\sigma,\Lfonly)
          =   x.\clazy\sigma \cup x.0\sigma
\end{align*}
and the \Lfunonly\ rule
for \SIF\ as:
\begin{align*}
\mathcal{L}((\SIF~x~e_1~e_2),\sigma,\Lfonly) =
                    &\mathcal{L}(e_1,\sigma,\Lfonly)~\cup
        \mathcal{L}(e_2,\sigma,\Lfonly)~\cup
          \{x.\clazy \sigma\}
\end{align*}
The rules for  \CDR, \PRIM\ and \NULLQ\ are also
modified similarly.

When   there  are   multiple   occurrences  of   \bcar,
\bcdr\  and \clazy,  $\hookrightarrow$ is  applied from
right  to left.   The reflexive  transitive closure  of
$\hookrightarrow$      will      be     denoted      as
$\stackrel{*}{\hookrightarrow}$.      The     following
proposition  relates  the  original  and  the  modified
liveness rules. 
\begin{proposition}
Assume  that  a  liveness   computation  based  on  the
original  set  of  rules  gives  the  liveness  of  the
variable  $x$   as $\sigma$
(symbolically,  $\Lanv{x}{}=  \sigma$).  Further,  let
$\Lanv{x}{}= \sigma'$ when the modified rules are used
instead      of      \Lfunonly.      Then      $\sigma'
\stackrel{*}{\hookrightarrow} \sigma$.
\end{proposition}

An  explanation of  why the  proposition holds  for the
modified  \CONS\ rule  is  given in  \cite{asati14lgc}.
The proposition also holds for other modified rules for
similar reasons.

\subsection{Generating equations for $\Lfonly_{\mathit f}$}
Given a  function $\mathit{f}$, we now  describe how to
generate  equations   for  the   demand  transformation
\Lfonly$_\mathit{f}$.        The       program       in
Figure~\ref{fig:mot-example2-a}   serves  as   a  running
example.  Starting with a  symbolic demand $\sigma$, we
determine  \Lfun{e_{\mathit{f}}}{\sigma}{\Lfonly}.   In
particular, we  consider $\Lanv{x_i}{}$,  the liveness
of the $i^{\mbox{\footnotesize th}}$ parameter $x_i$.   By   the   rule   {\sc
  live-define},   this   should    be   the   same   as
$\Lf{f}{i}{\sigma}$. Applying this to \length, we have:
$$
 \Lf{\length}{1}{\sigma} = \Lanv{\xl}{} = \clazy\sigma \cup \acdr\Lf{\length}{1}{\clazy\sigma}
  \cup \clazy\Lf{\length}{1}{\clazy\sigma}
$$

In general, the equations for \Lfonly\ are recursive as in the case of
\Lfonly$_\mathit{f}$. A closed  form solution for $\Lfone{\mathit{f}}$
can be derived by observing that each of the liveness rules modifies a
demand  only by  prefixing it  with symbols  in the  alphabet $\lbrace
\acar, \acdr,\bcar,  \bcdr, \clazy  \rbrace$. Therefore we  can assume
that $\Lf{f}{i}{\sigma}$ has the closed form:
\begin{eqnarray}
\label{eq:LF:DI}
  \Lf{f}{i}{\sigma} = \Df{f}{i}\sigma
\end{eqnarray}
where \Df{f}{i}  are sets of strings  over the alphabet
mentioned above.  Substituting the  guessed form in the
equation   describing    \Lfonly$_{\mathit   f}$,   and
factoring  out   $\sigma$,  we  get  an   equation  for
\Df{f}{i}  that   is  independent  of   $\sigma$.   Any
solution   for   \Df{f}{i}   yields  a   solution   for
\Lfonly$_{\mathit  f}$.   Applied to  \Lfonly$_{\mathit
  \length}$, we get:
  \begin{eqnarray*}
&&  \Lf{\length }{1}{\sigma} = \Df{\length}{1}\sigma,~\mbox{and}\\
&&   \Df{\length}{1} = \clazy \cup \acdr\Df{\length}{1}\clazy
       \cup \clazy\Df{\length}{1}{\clazy}
  \end{eqnarray*}

Note that this equation can also be viewed as a CFG with \{\acdr,
\clazy\} as terminal symbols and \Df{\length}{1} as the sole
non-terminal.

\subsection{Generating liveness equations \Lv\  for function bodies}
\label{sec:bodylivenessbodies}

To  avoid analyzing  the body  of a  function for  each
call, we calculated the  liveness for the arguments and
the variables in a function  with respect to a symbolic
demand  $\sigma$.   To  get   the  actual  liveness  we
calculate an  over-approximation of the  actual demands
made by  all the  calls and  calculate the  liveness at
each  GC  point  inside  the  function  based  on  this
approximation.  The 0-CFA-style {\em summary demand} is
calculated by  taking a union  of the demands  at every
call site of a function.

%% Consider a function  $g$ containing a call to  $f$ at a
%% site    $\pi$,   say    $\pi\!\!:\!(\LET~x   \leftarrow
%% (f\,y_1\,\ldots\,y_n)~\IN~\pi_i:e)$.  Let the demand on
%% $g$  be  $\sigma_g$  and,  based on  this  demand,  the
%% liveness of  $x$ at  $\pi_i$ be $\Lanv{i}{x}$.   By the
%% $\LET$ rule of Figure~\ref{fig:live-judge}, the call at
%% $\pi$   contributes   $\Lanv{i}{x}$   to   the   demand
%% $\sigma_f$. Let  us denote  the contribution of  a call
%% site $\pi$ in  a function $g$ to the  overall demand on
%% the function  $f$ as  $\deltacall{f}{\pi}{g}$. Assuming
%% that there are $k$ call  sites to function $f$, $\pi^1$
%% (in function $g^1$) \ldots $\pi^k$ (in function $g^k$),
%% the     over-approximation     of     $\sigma_f$     is
%% $\deltacall{f}{\pi^1}{g^1}     \cup     \cdots     \cup
%% \deltacall{f}{\pi^k}{g^k}$.  The distinguished function
%% \mainpgm\ is  a special case.   We assume it  is called
%% through    a    printing    mechanism    with    demand
%% $\sigma_\mainpgm   =   \{\acar,\acdr\}^\ast$   (denoted
%% $\sigma_{\!all}$) if \mainpgm\  returns a structure and
%% $\epsilon$ if it returns a base value.

For  the  running  example (Figure~\ref{fig:mot-example2-a}),  $\length$  has  calls  from  $\main$  at
$\pi_{12}$ and a recursive call at $\pi_6$. The demands on these calls
are $\epsilon$ and $\clazy\sigma_{\mathit{\length}}$. Thus:
\begin{eqnarray*}
\sigma_{\length}    &=&
 \{\epsilon\}  ~\cup~{\clazy\sigma_{\mathit{\length}}}
\end{eqnarray*}
As examples, the closure variables $\xl_{\mbox{$\pi$}_1}$ and $\xl_{\mbox{$\pi$}_5}$, and the
stack variables $\pl$ and $\pa$ have the following liveness in terms
of  $\sigma_{\length}$:
\begin{align*}
  \Lv_{\xl_{\mbox{$\pi$}_1}} &= \clazy\sigma_{\mathit{\length}} \\
  \Lv_{\xl_{\mbox{$\pi$}_5}} &=  (\acdr\Df{\length}{1}{\clazy} \cup
  \clazy\Df{\length}{1}{\clazy}) \sigma_{\mathit{\length}} \\
  \Lanv{\pl}{} &= (\clazy \cup \acdr\Df{\length}{1}\clazy
  \cup \clazy\Df{\length}{1}{\clazy})\sigma_{\mathit{\length}} \\
  \Lanv{\pa}{} &= \clazy \bcar \Df{\length}{1}\{\epsilon\}
\end{align*}

In  summary, the  equations  generated during  liveness
analysis are:
\begin{enumerate}
\item   For  each   function  $\mathit{f}$,   equations
  defining \Df{f}{i} for use by \Lfonly$_{\mathit f}$.
\item For each function $\mathit{f}$, an equation defining the summary
  demand $\sigma_{\mathit f}$ on $e_f$.
\item   For  each   function  $\mathit{f}$   (including
  $\mainpgm$) an  equation  defining
  liveness for each garbage in  $e_{\mathit f}$.
\end{enumerate}
\subsection{Solving liveness equations---the grammar
interpretation}\label{sec:grammar-formulation}      The
equations  above   can  now  be  re-interpreted   as  a
context-free    grammar   (CFG)    on   the    alphabet
$\lbrace\acar,  \acdr,  \bcar,  \bcdr,  \clazy\rbrace$.
Let \var{$X$}  denote the  non-terminal for  a variable
$X$  occurring on  the LHS  of the  equations generated
from  the  analysis.  We  can  think  of the  resulting
productions as being  associated with several grammars,
one for each non-terminal \var{\Lanv{x}{}} regarded as
a  start  symbol.   As  an  example,  the  grammar  for
\var{\Lanv{\xl}{}} and \var{\Lanv{\pa}{}}    comprise     the     following
productions:
\begin{eqnarray*}
  \var{\Lanv{\xl}{}}  &\rightarrow& 
  \clazy\var{\sigma_{\length}} \mid \acdr \var{\Df {\length}{1}}{\clazy\var{\sigma_{\length}}}  \\
  & & \mid
  \clazy\var{\Df{\length}{1}}{\clazy\var{\sigma_{\length}}} \\
  \var{\Df{\length}{1}} &\rightarrow& \clazy \mid
  \acdr\var{\Df{\length}{1}}\clazy
       \mid \clazy\var{\Df{\length}{1}}{\clazy}\\
\langle {\sigma_{\length}} \rangle
&\rightarrow&
\epsilon  \mid \clazy\langle{\sigma_{\length}}\rangle \\
\var{\Lanv{\pa}{}} &\rightarrow& \clazy \bcar \var{\Df{\length}{1}}
\end{eqnarray*}
Other  equations  can   be  converted  similarly.   The
language   generated   by  \var{\Lanv{x}{}},   denoted
$\mathscr{L}(\var{\Lanv{x}{}})$,   is    the   desired
solution  of  \Lanv{x}{}.    
However, recall that the decision problem that we are interested in
during garbage collection is the following: 
\begin{quote}
Let $x.\alpha$ be a forward access path (consisting only
of edges  \acar\ and \acdr\  but not \bcar,  \bcdr\ or
\clazy).       Let      $\mathscr{L}(\var{\Lanv{x}{}})
\stackrel{*}{\hookrightarrow}  \sigma$, where  $\sigma$
consists  of  forward  access  paths  only.  Then  does
$\alpha \in \sigma$?
\end{quote}

We now model  the above problem as one  of deciding the
membership of  a context-free grammar augmented  with a
fixed set of unrestricted productions.

\begin{definition}\label{def:specialgrammar}
Consider the  grammar $(N,T,  p_1\cup p_2,S)$  in which
$N$ is  a set  of non-terminals,  $T =  \{\acar, \acdr,
\bcar,  \bcdr,  \clazy,  \$\}$,   $p_1$  is  a  set  of
context-free    productions     that    contains    the
distinguished  production   $S  \rightarrow  \alpha\$$,
$\alpha$ is a  string of grammar symbols  that does not
contain $S$, and $p_2$ is the fixed set of unrestricted
productions    $\bcar\acar    \rightarrow    \epsilon$,
$\bcdr\acdr    \rightarrow   \epsilon$,    $\clazy\acar
\rightarrow \clazy$,  $\clazy\acdr \rightarrow \clazy$,
and $\clazy\$ \rightarrow \$ $.
\end{definition}

%% From Sections \ref{sec:liveness}  and \ref{sec:computing}, it is clear
%% that the results of liveness analysis of any program can be modeled by
%% the kind  of grammar described above. The  following proposition shows
%% that the converse also holds.
%% \begin{proposition}
%% Given    a     grammar    $G$    of    the     form    described    in
%% Definition~\ref{def:specialgrammar},  it is  possible  to construct  a
%% program $p$  with program points  $\pi_i$ and variables $x$  such that
%% the liveness  analysis of $p$  is same as  $G$ except for a  change in
%% non-terminal names which now have the form \var{\Lanv{i}{x}}.
%% \end{proposition}
We first show that membership problem of the class of grammars in
Definition~\ref{def:specialgrammar} is undecidable. However note that
the grammars may not be necessarily generated from liveness analysis.
\newcommand{\state}{\ensuremath{\mathsf{S}}}
\newcommand{\nont}[2]{\ensuremath{\mathsf{S}_{#1}^{#2}}}  
\begin{lemma}\label{lemma:grammar-undecidable}
Given a grammar    $G$    of   the    kind    described    in
Definition~\ref{def:specialgrammar}  and a forward access path $\alpha$
consisting  of symbols \acar\  and \acdr\  only, the membership problem
$\alpha\$ \in \mathscr{L}(G)$ is undecidable.
\end{lemma}

\begin{proof}
Given a Turing machine and  an input $w\in (1+0)^*$, we
construct a grammar $G$ such that the machine will halt
on    $w$    if    and   only    if    $\$    \in
\mathscr{L}(G)$. The grammar includes  the fixed set of
unrestricted               productions               in
Definition~\ref{def:specialgrammar}.

We shall represent a Turing machine (TM) configuration 
as $w_l(\state,c)w_r$, where $w_l$ is the string to the
left of the head, $w_r$ is the string to the right, $c$
is the symbol under the head and \state\ is the current
state of  the TM.   For each combination  of state
and symbol  $(\state,c)$, the grammar will  contain the
non-terminal \nont{}{c}. We shall synchronize each move
of the TM to  a derivation  step using  a context
free production, followed, if possible, by a derivation
step using either  $\bcar\acar \rightarrow \epsilon$ or
$\bcdr\acdr   \rightarrow    \epsilon$.    After   each
synchronization,  we  shall   establish  the  following
invariant  relation between  the TM  configuration
and the sentential form:

\begin{quote}
  If    the   configuration    of   the    TM   is
  $w_l(\state,c)w_r$, then the  sentential form will be
  $\overline{w}_l\nont{}{c}\,w_r\$ $,                where
  $\overline{w}_l$ is  the same as $w_l$  but with each
  symbol $d$ in $w_l$ replaced by $\overline{d}$.
\end{quote}

Assume that  the TM starts  in a  state $S_\mathit{init}$ with  a tape
$cw$ and  the head positioned on  the symbol $c$. Then  the sentential
form    corresponding     to    the    initial     configuration    is
$\nont{\mathit{init}}{c}w\$$ (we  can assume that there  is a production
$\nont{}{} \rightarrow  \nont{\mathit{init}}{c}w\$$, where  \nont{}{} is
the start symbol of the  grammar). Further correspondences between the
TM moves and the grammar productions are as follows:

\begin{enumerate}
\item For each transition  $(S_i, c) \rightarrow (S_j,c',L)$, there are
  two  productions $\nont{i}{c}  \rightarrow  \acar \nont{j}{\acar}c'$
  and $\nont{i}{c} \rightarrow \acdr \nont{j}{\acdr}c'$.
\item For each  transition $(S_i, c) \rightarrow (S_j,c',R)$, there
  are two productions
  $\nont{i}{c} \rightarrow \overline{c'} \nont{j}{\acar}\bcar$ and $\nont{i}{c}
  \rightarrow \overline{c'} \nont{j}{\acdr}\bcdr$. 
\end{enumerate}
The idea  behind the  productions is explained  with an
example:  Assume that  the current  sentential form  is
$\bcar\bcdr\nont{i}{\acar}\acar\acar\$$. Also  assume that  the
TM  has  a   transition  $(S_i,\acar)  \rightarrow
(S_j,\acdr, L)$.  Since the next corresponding  step in
the  derivation  has  to  be  done  without  any  prior
knowledge of whether the symbol to the left of the tape
is a \acar\  or a \acdr, two  productions are provided,
and  the  invariant  will  be maintained  only  if  the
production         $\nont{i}{\acar}         \rightarrow
\acdr\nont{j}{\acdr}\acdr$ is chosen  for the next step
in  the  derivation.    This  gives  the  configuration
$\bcar\bcdr\acdr\nont{j}{\acdr}\acdr\acar\acar\$$.
Simplification   with    the   production   $\bcdr\acdr
\rightarrow               \epsilon$              yields
$\bcar\nont{j}{\acdr}\acdr\acar\acar\$$,    which   exactly
corresponds  to   the  changed  configuration   of  the
TM.  Notice  carefully that a wrong  choice breaks
the invariant  and it cannot be  recovered subsequently
by any choice of productions.

After the  TM has  halted, there  are further  ``cleanup'' derivations
that derive  $\epsilon$ only if  the invariant has been  maintained so
far.    For   every   symbol   $c$,  we   introduce   a   non-terminal
$\nont{\mathit{final}}{c}$  where   $\nont{\mathit{final}}{}$  is  the
final state  of the  TM. We add  productions $\nont{\mathit{final}}{c}
\rightarrow         \acar        \nont{\mathit{final}}{c}$         and
$\nont{\mathit{final}}{c} \rightarrow  \acdr \nont{\mathit{final}}{c}$
for cleaning up the \bcar\ and \bcdr\  symbols on the left of the head
and                $\nont{\mathit{final}}{c}               \rightarrow
\nont{\mathit{final}}{c}\bcar$      and      $\nont{\mathit{final}}{c}
\rightarrow \nont{\mathit{final}}{c}\bcdr$ for  cleaning up \acar\ and
\acdr\ on the right of the tape head.  This completes the
reduction.\\ \mbox{}
\end{proof}

We now show that the proof can be replayed for the class of grammars
generated from liveness analysis of programs.

\begin{lemma}\label{lemma:grammar-from-analysis-undecidable}
Given   a    grammar    $G$    of   the    kind    described    in
Definition~\ref{def:specialgrammar}  that  is  generated  by  liveness
analysis  of a  program  and  a forward  access  path $\alpha$,    the
membership problem $\alpha\$ \in \mathscr{L}(G)$ is undecidable.
\end{lemma}

\begin{proof}
Given a Turing machine and an input string, we shall define a function
for   each  non-terminal   introduced  in   the  proof   of
Lemma~\ref{lemma:grammar-undecidable}.
%% Liveness  analysis  of  $f$  will  result in  not  only  contain  a
%% production  $P$ itself,  but could  also contain  extra productions
%% that necessarily contain the \clazy\ symbol.
For instance, corresponding to the production $\nont{i}{\acar} \rightarrow
\acar \nont{j}{\acar}\bcar$,
the program is:

\input{equivalent-program}

The body of the function corresponds  to the RHS of the production. If
there are  more than one  productions with the same  LHS non-terminal,
then the function for the non-terminal joins the program fragments for
the  RHS  with an  $\SIF$.   If  there  are  $n$ such  functions,  then
\mainpgm\ creates  a $n$-way  branch and  inserts a  single call  to a
distinct function in each branch. It  can be verified that the grammar
for      the     function      shown      above,     includes      the
productions shown below. $\nont{\mathit{all}}{}$   corresponds   to  the   demand
$\sigma_{\mathit{all}}$:
\begin{align*}
\nont{i}{\acar} &\rightarrow \acar\nont{j}{\acar}\bcar\nont{\mathit{all}}{} \\
\nont{i}{\acar} &\rightarrow \clazy\nont{i}{\acar}\bcar\nont{\mathit{all}}{} \\
\nont{\mathit{all}}{} &\rightarrow \epsilon \mid \acar\nont{\mathit{all}}{} \mid
\acdr\nont{\mathit{all}}{}
\end{align*}

Notice that  each production  in Lemma~\ref{lemma:grammar-undecidable}
had a single non-terminal on the RHS. There are two characteristics of
the grammar produced from liveness-analysis  that are are relevant for
replaying the  earlier proof:  (a) If $\mathsf{S}  \rightarrow \beta_1
\mathsf{S'}  \beta_2$  was a  production  in  the earlier  proof,  the
grammar  generated from  liveness  analysis of the constructed program
will  have the  production
$\mathsf{S}      \rightarrow      \beta_1     \mathsf{S'}      \beta_2
\nont{\mathit{\mathit{all}}}{}$,  and (b)  other  productions are  of the  form
$\mathsf{S}       \rightarrow       \beta_1'\mathsf{S'}       \beta_2'
\nont{\mathit{all}}{}$.

It is clear that if the TM given as an instance of the halting problem
accepts the input string, then the earlier derivation can be replayed,
every   time   replacing   $\nont{\mathit{\mathit{all}}}{}$   in   the
sentential form by $\epsilon$. However, if  the TM does not accept the
input string, then every sentential form derived from the start symbol
would    have     at    least    one    non-terminal from the grammar
of Lemma~\ref{lemma:grammar-undecidable} that is different
from $\nont{\mathit{final}}{c}$.   Thus  $\$$  would  not  be
derivable from the grammar.
\end{proof}

We circumvent  the problem  of undecidability  by over-
approximating the CFG by non-deterministic finite state
automata (NFA). The NFAs  are then simplified using the
$\hookrightarrow$ rules.   Finally the  simplified NFAs
are converted to DFAs.

%------------------------------------------------------------%

\begin{figure}[t!]
\scalebox{0.8}{
  \begin{tabular}{cc}
\psset{unit=1mm,nodesep=0mm,labelsep=0.5mm}
\begin{pspicture}(0,-5)(50,12)
  \putnode{t0}{origin}{5}{0}{\var{\Df{\length}{1}}}
  \putnode{t1}{t0}{15}{0}{\pscirclebox{\mbox{\ \ \ \ }}} \hspace{5mm}
  \putnode{t2}{t1}{15}{0}{\pscirclebox[doubleline=true]{\mbox{\ \ \ \ }}}
  \psset{arrows=->}
  \ncline{t0}{t1}
  \ncline{t1}{t2}
  \putnode{l0}{t1}{7}{2}{\clazy}
  \nccurve[angleA=45, angleB=135, ncurv=4, nodesep=-1]{t1}{t1}
  \putnode{l1}{t1}{0}{8}{\acdr, \clazy}
  \nccurve[angleA=45, angleB=135, ncurv=3, nodesep=-1]{t2}{t2}
  \putnode{l2}{t2}{0}{8}{\clazy}
\end{pspicture}
&
\psset{unit=1mm,nodesep=0mm,labelsep=0.5mm}
\begin{pspicture}(0,-5)(30,12)
  \putnode{t0}{origin}{5}{0}{\var{\Df{\length}{1}}}
  \putnode{t1}{t0}{15}{0}{\pscirclebox[doubleline=true]{\mbox{\ \ \ \ }}}
  \psset{arrows=->}
  \ncline{t0}{t1}
  \nccurve[angleA=45, angleB=135, ncurv=3, nodesep=-1]{t1}{t1}
  \putnode{l1}{t1}{0}{8}{\acdr}
\end{pspicture}
\\
(a) & (b) \\
\psset{unit=1mm,nodesep=0mm,labelsep=0.5mm}
\begin{pspicture}(0,-5)(50,12) %\psframe(0,-5)(40,12)
  \putnode{t0}{origin}{5}{0}{\var{\Lanv{\pa}{}}}
  \putnode{t1}{t0}{9}{0}{\pscirclebox{\mbox{\ \ \ \ }}} \hspace{5mm}
  \putnode{t2}{t1}{12}{0}{\pscirclebox{\mbox{\ \ \ \ }}}
%  \putnode{t3}{t2}{12}{0}{\pscirclebox[doubleline=true]{\mbox{\ \ \ \ }}}
  \psset{arrows=->}
  \ncline{t0}{t1}
  \ncline{t1}{t2}
  \putnode{l0}{t1}{5}{2}{\clazy}
%  \ncline{t2}{t3}
%  \nccurve[angleA=55, angleB=125, ncurv=3, nodesep=-.8]{t3}{t3}
%  \putnode{l2}{tD1}{0}{10}{\acdr}
 \putnode{tD1}{t2}{12}{0}{\pscirclebox{\mbox{\ \ \ \ }}} \hspace{5mm}
  \putnode{tD2}{tD1}{12}{0}{\pscirclebox[doubleline=true]{\mbox{\ \ \ \ }}}
  \psset{arrows=->}
  \ncline{t2}{tD1}
  \putnode{l1}{t2}{5}{2}{\bcar}
  \ncline{tD1}{tD2}
  \putnode{lD0}{tD1}{5}{2}{\clazy}
  \nccurve[angleA=45, angleB=135, ncurv=4, nodesep=-1]{tD1}{tD1}
  \putnode{lD1}{tD1}{0}{8}{\acdr, \clazy}
  \nccurve[angleA=45, angleB=135, ncurv=3, nodesep=-1]{tD2}{tD2}
  \putnode{lD2}{tD2}{0}{8}{\clazy}
\end{pspicture}
&
\psset{unit=1mm,nodesep=0mm,labelsep=0.5mm}
\begin{pspicture}(0,-5)(20,12) %\psframe(0,-5)(40,12)
  \putnode{t0}{origin}{5}{0}{\var{\Lanv{\pa}{}}}
  \putnode{t1}{t0}{9}{0}{\pscirclebox{\mbox{\ \ \ \ }}} \hspace{5mm}
  \psset{arrows=->}
  \ncline{t0}{t1}
\end{pspicture}
\\ 
(c) & (d)  \end{tabular}}
\caption{(a) The  grammar rules for \var{\Df{\length}{1}}
  converted into an automaton, and (b) its DFA. The same for \var{\Lanv{\pa}{}}
  are in (c) and (d).}\label{fig:example-automata}
\figrule
\end{figure}

%\subsection{A liveness-based garbage collection scheme}
\label{sec:live-clo}
  \SetStartEndCondition{ }{}{}%
  \SetKwProg{Pro}{procedure}{\string:}{}
  \SetKwProg{Fn}{function}{\string:}{}
  \SetKwFunction{Range}{range}%%
  \SetKw{KwTo}{in}\SetKwFor{For}{for}{\string:}{}%
  \SetKwIF{If}{ElseIf}{Else}{if}{:}{elif}{else:}{}%
  \SetKwFor{While}{while}{:}{fintq}%
  \AlgoDontDisplayBlockMarkers\SetAlgoNoEnd\SetAlgoNoLine%
  \SetKwFunction{Lgc}{$\mathsf{lgc}$}%
  \SetKwFunction{Copy}{$\mathsf{copy}$}%
  \SetKwFunction{LCopy}{$lgcCopy$}%
  \SetKwFunction{RCopy}{$rgcCopy$}%
\begin{algorithm}[ht!]
  \Pro{\Lgc{}}
     {
       \For {each reference $\mathsf{ref}$ in root set}
            {$\mathsf{ref}$ = \Copy($\mathsf{ref}, \mathsf{init}(\mathsf{ref.dfa})$)\;}
            ${\mathsf{copyReferencesOnPrintStack}()}$\;  
     }
     \Fn{\Copy{$\mathsf{ref}, \mathsf{state}$}}
        {
          \eIf {$\mathsf{final}(\mathsf{state}$)}
             {
               $\mathsf{newRef} = \mathsf{dupHeapCell}(\mathsf{ref})$\;
               \If{$\mathsf{ref\!.cell}$($\mathsf{ref}$)
                 is a cons cell $\mathsf{(cons~arg_1~arg_2)}$}
                  {
                    {
                      $\mathsf{newRef}\!.\mathsf{arg_1}  = \Copy(\mathsf{arg_1}, \mathsf{next}(\mathsf{state}, 0))$\;
%                      $\mathsf{newRef}\!.\mathsf{arg_1}   =  \mathsf{newCar}$\;
                    }
                    {
                      $\mathsf{newRef}\!.\mathsf{arg_2} =  \Copy(\mathsf{arg_2},
                      \mathsf{next}(\mathsf{state} , 1))$\;
%                      $\mathsf{newRef}\!.\mathsf{arg_2}   = \mathsf{newCdr}$\;
                    }    
                  }
               \If
                  {$\mathsf{ref\!.cell}$ is a
                    closure cell, generically  $\mathsf{(binop~arg_1~arg_2)}$}
              %      
                   { 
                     $\mathsf{newRef}\!.\mathsf{arg_1} = \Copy(\mathsf{arg_1}, \mathsf{init}(\mathsf{arg_1.dfa}))$\;
                     $\mathsf{newRef}\!.\mathsf{arg_2} = \Copy(\mathsf{arg_2}, \mathsf{init}(\mathsf{arg_2.dfa}))$\;
                   }
             }
             {$\mathsf{newRef = ref}$}
\KwRet $\mathsf{newRef}$\;
        }
        \caption{Liveness-based
          garbage collection.  \label{algo:lgc-a}}
\end{algorithm}

\subsubsection{From CFGs to GC-ready DFAs}
\label{sec:NFA-approx}
 We use  the algorithm by Mohri  and Nederhof~\cite{mohri00regular} to
 approximate a CFG by a {\em strongly regular\/} grammar.  For
 example, the grammar
 fragment  for  the  non-terminal  $\var{\Df{\length}{1}}$  after  the
 Mohri-Nederhof transformation is:
 \begin{eqnarray*}
   \var{\Df{\length}{1}} &\rightarrow& \clazy\var{\Df{\length}{1'}} \mid
   \acdr\var{\Df{\length}{1}}
   \mid \clazy\var{\Df{\length}{1}}\\
   \var{\Df{\length}{1'}} &\rightarrow& \clazy\var{\Df{\length}{1'}}
   \mid \epsilon
 \end{eqnarray*}

The strongly regular grammar is converted  into a set of NFAs, one for
each $\var{\Lanv{x}{}}$.  The  $\hookrightarrow$ simplification is now
done on the NFAs by  repeatedly introducing $\epsilon$ edges to bypass
pairs  of consecutive  edges  labeled \bcar\acar\  or \bcdr\acdr\  and
constructing  the $\epsilon$-closure  till a  fixed point  is reached,
after which  the edges labeled  \bcar \  and \bcdr\ are  deleted.  The
details of the  algorithm, its correctness and  termination proofs are
given  in~\cite{karkare07liveness,asati14lgc}.   The
resulting  automaton   has  edges  labeled  with   \acar,  \acdr\  and
\clazy\ only.  In this  automaton, for every  edge labeled  \clazy, we
check if the source node of the edge has a path to a final 
state.  If it does, we mark  the source node as final. Finally, we
remove all the edges labeled \clazy\  and convert the automaton into a
deterministic    automaton.    These steps   effectively   implement    the
$\hookrightarrow$   simplification  rules   for   \bcar,  \bcdr,   and
\clazy\ to  obtain forward access paths.  While checking for  liveness during
garbage  collection, a  forward access path  is valid  only if  it can
reach a  final state.  Figure~\ref{fig:example-automata}(a)  shows the
NFA that is obtained from  the grammar for \var{\Df{\length}{1}}, and   the
final  DFA  is  shown in  Figure~\ref{fig:example-automata}(b).   This
expectedly  says that for a  demand  $\sigma_{\length}$,  the
liveness of  the argument of \length\  is $\acdr^{*}$  (the spine  of the  list is
traversed).  Similarly, Figure~\ref{fig:example-automata}(c) shows the
NFA     for     \var{\Lanv{\pa}{}}.      The     DFA     shown     in
Figure~\ref{fig:example-automata}(d)  does  not   accept  any  forward
paths.   This  reflects  the  lazy  nature  of  our  language.   Since
\length\ does not  evaluate the member elements of  the argument list,
the  closure for  \pa\ is  never evaluated  and is  reclaimed whenever
liveness based garbage collection is triggered beyond $\pi_9$.

\newcommand{\hgt}{2.9cm}
\newcommand{\LGCLine}{blue}
\newcommand{\RGCLine}{red}
\newcommand{\ReachLine}{black}
\newcommand{\UseLine}{light-blue}

\begin{figure*}[t!]
\renewcommand{\arraystretch}{.1}
\begin{tabular}{@{}c@{}@{}c@{}@{}c@{}}
   \hskip -4mm{\epsfig{file=fibheap.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=sudoku.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=nperm.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=paraffins.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=lcss.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=huffman.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=knightstour.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=nqueens.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=deriv.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=treejoin.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=lambda.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=gc_bench.eps, height=\hgt}}
%\\ (a) & (b)
\end{tabular}%\vskip -10mm
 \caption{Memory usage.  
%% Column (a) shows complete  usage, column 
%%    (b) enlarges  a part  of the  usage.  
The \RGCLine\ and the \LGCLine\ curves indicate the number of cons
cells  in  the  active   semi-space  for  RGC  and  LGC
respectively.  The \ReachLine\ curve represents the number of
reachable cells and the  \UseLine\ curve represents the
number  of  cells  that  are actually  live  (of  which
liveness analysis does a static approximation).  x-axis
is the time measured  in number of cons-cells allocated
(scaled down by factor $10^5$). y-axis is the number of
cons-cells (scaled down by $10^3$).}
\label{fig:memory-usage} \figrule
\end{figure*}

\section{The garbage collection scheme}
\label{sec:GC-scheme}

Our experimental setup  consists of an interpreter
for  our  language,  a liveness  analyzer,  and  a
single generation copying  collector.  The garbage
collector can be configured to work on the
basis  of  reachability  (RGC  mode)  or  use  
liveness  DFAs  (LGC  mode).

In        the       liveness        analysis       described        in
Section~\ref{sec:liveness-analysis},   the  liveness   of  a   closure
variable is  derived from dependence  chains along all paths  from the
program point  where the closure  was created to  reachable evaluation
points.  Assume,  for the  sake of  concreteness, that  $\epath_1$ and
$\epath_2$ are two such evaluation points.  During garbage collection,
we would like to use a more precise liveness, based on the actual paths
taken  during  execution.   Therefore   we  create  separate  liveness
automata for dependences  along paths to $\epath_1$  and $\epath_2$, in
addition to  automata for dependences  along paths to  both $\epath_1$
and $\epath_2$.  The closure carries  the liveness environment for its
free  variables (as  pointers  to automata,  one  for each  variable).
Initially the liveness  environment is based on  the dependences along
both  $\epath_1$   and  $\epath_2$.   However,  after   evaluating  an
\SIF\ condition, the liveness environments are updated to one based on
either   $\epath_1$  or   $\epath_2$,  so   that  subsequent   garbage
collections are based on more precise liveness information.

Based on  the above considerations,  we restrict the  possible garbage
collection points in a function body to the following:
\begin{enumerate}
\item We  maintain static estimates  of the memory required  to create
  the closures for each function body.  On entering a function, if the
  available memory is  less than this requirement for  the function, a
  garbage collection is triggered.
\item Since the evaluation of a  \SIF\ condition may trigger a 
  collection, after evaluating the condition   the available memory
  is  checked once  again against  a  revised estimate  of the  memory
  (based on  value of the condition)  required to execute the  rest of
  the program.  A garbage collection is triggered if enough memory is unavailable.
\end{enumerate}

%% While  exploring  the
%% activation record of a  function, LGC uses the DFA
%% at  the  latest  program point  traversed  in  the
%% function's body.   Thus, if LGC is  invoked at the
%% program point $\pi$, it would use the liveness DFA
%% at $\pi$ itself for exploring  the root set of the
%% current   activation   record.   For   any   other
%% activation record in the stack, say for a function
%% ${\mathit  f}$   calling  ${\mathit  g}$   in  the
%% call-chain, the evaluation  point in ${\mathit f}$
%% which resulted in the  call to ${\mathit g}$ would
%% be used.

We shall call a  unit allocatable memory as a {\em  cell}.  A cell can
hold    a    basic    value    ($\mathsf{bas}$),    the    constructor
\CONS\ $(\mathsf{cons~  arg_1~arg_2}$) or a closure.   The closure, in
turn,    can    be    one   of    ($\mathsf{unop~arg}$),    ($\mathsf{
  binop~arg_1~arg_2}$)  and  function application  ($\mathsf{f~arg})$.
Here each $\mathsf{  arg_i}$ is a reference to another  heap cell.  In
addition, the closure also carries  a pointer to a DFA (denoted
$\mathsf{arg_i.dfa_i}$) for each $\mathsf{arg_i}$.
Algorithm~\ref{algo:lgc-a}   describes   the   garbage
collection scheme.  Starting with the root set, each cell pointed by a
live reference  (i.e. whose associated  DFA state is final)  is copied
using $\mathsf{copy}$.   Copying a \CONS\ cells  just involves copying
the  cell  itself   and  conditionally  copying  the   \CAR\  and  the
\CDR\ fields after referring to the  next states of the DFA.  If
the reference points to a closure, then, as noted earlier, the closure
carries pointers  to the liveness  DFAs of its  arguments.  These
are  used to recursively  initiate copying of  the arguments.
Note that  the copying  strategy for  $\mathsf{(unop~arg_1~arg_2)}$ or
$\mathsf{(f~arg_1)}$ are similar to $\mathsf{(binop~arg_1~arg_2)}$ and
have not been shown.

\begin{table*}[t!]
\caption{Statistics for liveness analysis and garbage collection}
\label{tab:exp-results}
\centering
\include{data-liveness-analysis}
\include{autotable}
\vskip -5mm
\end{table*}

%\subsection{Garbage collection for references on print stack}
The evaluation of  the top-level expression in a program  is driven by
a printing function~(Section
\ref{sec:semantics}).  
%% Printing of  atomic values is simple  and does
%% not  trigger a  garbage  collection.   In case  of  \CONS\ cells,  the
%% \CAR\ is  first printed followed  by \CDR.   In a lazy  language, both
%% \CAR\ and \CDR\  could be closures requiring evaluation,  and this may
%% trigger  more garbage  collections.  When  this happens.   we have  to
%% consider any  references that  might be  on the  print stack  and copy
%% them.
We extend liveness-based  garbage  collection  to
this function.
\section{Experimental evaluation}
\label{sec:experiments}
  Our benchmark consists of programs taken from nofib~\cite{nofib} and
  other repositories for  functional  programs~\cite{PLT-Scheme,
    gc_bench, huffman-sicp}.   All our  benchmarks were executed  on a
  machine having 8  core Intel(R) Core(TM) i7-4770 3.40GHz  CPU , 8192
  KB L2 cache, 16 GB RAM running 64 bit Ubuntu 14.04.

%% The  process  of  liveness-based  garbage  collection  involves  going
%% through  each  activation  record  on the  stack  and  exploring  each
%% variable in the current activation  record (root set). The liveness of
%% each variable is  determined using the program point  and the variable
%% name. The liveness  of all variables are stored as  DFA. These DFA are
%% then encoded as a table. A variable  is live only if an entry is found
%% in  this table.   All  cells that  are  live are  copied  to the  live
%% semi-space. In case  of \CONS\ cells the \CAR\ and  \CDR\ pointers are
%% chased  and if  they are  live  the copied  \CONS\ cell  will get  the
%% updated addresses of its \CAR\ and \CDR\ fields.

\subsection{Results}

The statistics  related to  liveness analysis  and DFA  generation are
shown in Table~\ref{tab:exp-results}(a). We  observe that the analysis
of  all  programs  except \verb@treejoin@  and  \verb@sudoku@  require
reasonable time.   The bottleneck in  our analysis  is the NFA  to DFA
conversion with  worst-case exponential behaviour. However,  since the
analysis has  to be  done only  once and its  results can be cached and
re-used, the time spent in analysis may be considered acceptable.

Table~\ref{tab:exp-results}(b) compares  garbage collection statistics
for RGC and  LGC.  We report the number of  garbage collection events,
average  number of  cells  reclaimed per  garbage collection,  average
number of cells  touched per garbage collection and the  total time to
perform all collections.   It is no surprise that the  number of cells
reclaimed per garbage  collection is higher and the  number of garbage
collections  lower  for  LGC.   The  cost of  LGC  is  higher  garbage
collection time, which increases the  overall execution time even with
reduced number of  collections.  However, the execution time  of LGC is
still comparable for most benchmarks (slowdown within 5X of
RGC in most cases) and better  for 3 benchmarks (2X speedup in the
best case).  Note that  \verb@gc_bench@~\cite{gc_bench} is a synthetic
benchmark that allocates complete binary
trees  of various  sizes that  are  never used  by the  program. As  a
result,  the benchmark  highly favours  LGC.  The  benchmark has  been
included  for  completeness,  and  we  do  not  consider  its  numbers
as being representative of real programs.

Memory  usage graphs  for select  benchmarks are  shown
Figure~\ref{fig:memory-usage},  Column   (a).   In some cases, the
number of  garbage collections  tend to be  very large,
hence  we  show only  a  slice of the original graph  with 
representative    behavior  (Column (b)).  In all the programs we can see
that  the   curve  corresponding  to  LGC   (\LGCLine\ line)
dips   below the RGC curve 
(\RGCLine\ line) during garbage collection.  The graphs also include the
curve    for   reachable    cells   (\ReachLine)  and live cells
(\UseLine). These were  obtained
 by forcing  RGC to  run at   very  high
frequency. The curve for live cells were obtained by recording heap
access times and post processing the data 
at the end of the program. Note that the size of an LGC cell is 1.16
times the size of a RGC cell.

As  demonstrated by  the gap  between the  \RGCLine\ and  the \UseLine\
lines, a  large number of  cells which are  unused by the  program are
still copied  during RGC. LGC does  a much better job  of closing this
gap but still falls short of the  precision achieved by LGC in case of
eager languages ~\cite{asati14lgc}.  A major source of inefficiency in
LGC is  multiple traversals  of already copied  heap cells.  Since LGC
does not mark the heap cells after the first visit, the same cells can
be  repeatedly  visited  with   different  liveness  states.  We  have
mitigated  this  problem  by   implementing  a  heuristic  which  tags
\CONS\ cells with the set of states already visited.

\section{Related work} 
\label{sec:relatedwork}
The impact of  liveness on the effectiveness of  garbage collection is
investigated in~\cite{Hirzel}.   The observation is that  liveness can
significantly  impact   garbage  collection,  but  only   when  it  is
interprocedural.  As far as memory requirement is concerned, our paper
demonstrates this observation.

There have been several attempts to use liveness analysis to improve 
garbage collection for imperative languages.~\cite{khedker07heap}
presents a liveness analysis and uses the results for inserting 
nullifying statements in Java programs. In ~\cite{ran.shaham-sas03},
an automaton modelling heap accesses is created, and
temporal properties like liveness are checked  against the model.  
However both the approaches are restricted to intraprocedural scope.

In the space of functional  languages, there are: rewriting methods such as
deforestation~\cite{wadler88deforest,gill93ashort,chitil99deforest},
sharing  analysis   based  reallocation~\cite{jones89compile},  region
based   analysis~\cite{tofte98region},   and    insertion   of
compile-time                                                nullifying
statements~\cite{inoue88analysis,lee05static,Hamilton}.            All
compile-time marking approaches rely on an efficient and precise alias
analysis and  cannot provide  significant improvement in  its absence.
An interesting approach suggested  in~\cite{HofmannJ03} is to annotate
the  heap usage  of first-order  programs through  linear types.   The
annotations   are  then   used  to   serve  memory   requests  through
re-allocation.  However, this requires the user to write programs in a
specific way.

Simplifiers~\cite{ONeill} are abstractly described as lightweight
daemons that attach themselves to program data and, when activated, improve
the efficiency of the program. Our liveness based garbage collection
as an instance of a  simplifier which is tightly coupled with garbage collectors.  The
approach that is closest to the  method described in this paper is the
liveness-based garbage collector implemented
in~\cite{karkare07liveness,asati14lgc} and address eager languages. We
extend this to handle lazy evaluation and closures.

%% A similar approach has been suggested in~\cite{ONeill} where they suggest augmenting garbage collection with extra information using {\it simplifiers}.

%% Previous attempts  to increase the  space efficiency
%% of functional programs  by additional reclamation of
%% memory fall  in two broad categories.  In the first,
%% the  program   itself  is  instrumented   to  manage
%% reclamation and reallocation without  the aid of the
%% GC.   Such attempts  include: sharing
%% analysis  based  reallocation~\cite{jones89compile},
%% deforestation
%% techniques~\cite{wadler88deforest,gill93ashort,chitil99deforest},
%% methods based on linear logic~\cite{hofmann00linear}
%% and region analysis~\cite{tofte98region}.  Closer to
%% our  approach, there  are  methods  that enable  the
%% garbage      collector      to     collect      more
%% garbage~\cite{inoue88analysis,lee05static}        by
%% explicitly  nullifying pointers  that are  not live.
%% However,  the nullification,  done at  compile time,
%% requires sharing  (alias) analysis.  Our  method, in
%% contrast, does  not require  sharing because  of the
%% availability of the heap  itself at runtime.  To the
%% best of our knowledge, this  is the first attempt at
%% liveness-based  marking of  the heap  during garbage
%% collection.

%% \begin{figure}[t]
%% \centerline{\epsfig{file=lambda.eps, height=4cm, width=7cm}}
%%  \caption{Memory  usage and garbage collection pattern of  {\tt
%% lambda}.}
%% \label{fig:memory-usage-lambda} \figrule
%% \end{figure}
\vskip -5mm
\section{Future work and conclusions}
\label{sec:conclusion}
%% We have  defined a notion of  liveness on structured
%% data; this generalizes classical liveness and strong
%% liveness.    We  started   with   a  general   fully
%% context-sensitive analysis  which we  proved correct
%% with respect  to a minefield semantics  (this models
%% the  effect  of  garbage  collection  between  every
%% evaluation step).

%% To avoid scalability issues (and to avoid performing
%% part  of the  liveness computation  at run  time) we
%% defined an  0-CFA version of this  liveness analysis
%% in  which demands  for function  $f$ at  all calling
%% contexts   are  conflated   into  a   single  demand
%% $\sigma_f$.  This  enabled us to treat  the liveness
%% equations   symbolically    obtaining   context-free
%% grammars  for liveness  at each  GC point  (calls to
%% user  functions and  to $\CONS$).   These were  then
%% converted to  DFA for  run-time consultation  by the
%% garbage   collector.     Experiments   confirm   the
%% precision of the analysis.

%% To   obtain  performance   figures  we   compared  a
%% reachability-based   garbage    collector   with   a
%% liveness-based collector.  This showed a decrease in
%% the  number  of  GCs,  more  garbage  collected  per
%% invocation.  A  significant benefit  of LGC  is that
%% programs can run in  smaller memory when compared to
%% RGC. This is potentially  useful in situations where
%% memory is limited---as with embedded systems.  For a
%% majority of  programs, the garbage  collection times
%% were reduced.

%% One issue we highlighted  was that while fewer nodes
%% were marked (and  more garbage collected), sometimes
%% cons cells  could be visited and  traversed multiple
%% times  with  different  sets of  liveness  paths  to
%% explore.  Further work  includes improvements to the
%% classical copying  collector to  reduce the  cost of
%% this.
We have  extended liveness-based  garbage collection  to lazy
  languages  and shown  its benefit  for a  set of  benchmark
  programs.  We  defined a liveness analysis  of programs manipulating
  the  heap and  proved  it  correct with  respect  to a  non-standard
  semantics that served as a specification of liveness.  The result of
  the analysis  is a  set of grammars,  whose membership  question was
  shown to be undecidable. The  grammars are therefore approximated by
  DFAs and used by the garbage collector to improve collection.

In addition to  evaluated data, our collector  also reclaims closures.
For this, we  had to modify the standard run-time  representation of a
closures  to carry  liveness  of its free  variables and  periodically
update the liveness during execution  time.  As expected, our garbage
collector  reclaims more  garbage and  reduces memory  requirement of
programs.  An  additional benefit is  that in  spite of using  a more
expensive collector, the execution times remains comparable in
most cases, and even improves for some programs.

The graphs show that even in lazy languages there are large amount of
dead cells that can be collected early. In spite of collecting
closures there still exists a gap between actual and perceived
livenesses.
Language features that could be introduced are higher-order functions
and recursive lets. We believe that functions can be used to model
recursive lets and this should be a simple extension to our liveness
analysis.
For higher-order we intend to use firstification techniques~\cite{Mitchell:2009}.

  Orthogonally, we  plan to  improve the  efficiency of
  the liveness based garbage collector using heuristics
  such   as  limiting   the  depth   of  DFA,   merging
  nearly-equivalent    states    and    using    better
  representation(for example BDDs~\cite{Bryant86})    and    algorithms   for    automata
  manipulation.  We   also  need  to   investigate  the
  interaction   of  liveness   with  other   collection
  schemes,   such  as   incremental  and   generational
  collection.  It might be interesting to use a mixed mode garbage collection
  scheme which allows the costs of LGC to be amortized over several runs
  of RGC. In summary, we  need to investigate ways
  to make liveness  based garbage collection attractive
  for practical collectors.

\clearpage
\balance
\bibliography{fun_hra}{}
\bibliographystyle{abbrv}
\clearpage

\appendix
%\section{Appendix: not to be included in the paper.}
%%%%% prasanna minefield
\begin{figure*}[b!]
\noindent{\Large \bf A. Minefield semantics(Not included in the paper)}
\begin{center}\footnotesize
\renewcommand{\arraystretch}{1.2}

\begin{tabular}{|c|c|c|}
\hline
Premise & Transition & Rule name \\ 
\hline
\hline
&\makecell{ $\rho, (\rho', \ell, e, \cred{\sigma'})\!:\!S,
  \heap, \kappa, \cred{\sigma}$  $\rightsquigarrow \rho', S, \heap[\ell :=
    \kappa], e, \cred{\sigma'}$ }   &  {\sc const}
\\
\hline
& \makecell[t]{$\rho, (\rho', \ell, e, \cred{\sigma'})\!:\!S, \heap,
  (\CONS~x~y), \cred{\sigma}$  $\rightsquigarrow$ \\ $
  \rho', S, \heap[\ell := (\rho(x),\rho(y))], e, \cred{\sigma'}$}     &  {\sc cons} \\
\hline
\makecell[t]{$\cred{GC(\rho_1, S, \heap_1, (\CAR~x), \sigma) = (\rho, S, \heap)}$,\\
  $\rho(x) \mbox{ is } \bot$} & $\rho, S,
  \heap, (\CAR~x), \cred{\sigma} \rightsquigarrow \bang$   &
{\sc car-bang} 
\\
\hline
\makecell[t]{$\cred{GC(\rho_1, S, \heap_1, (\CAR~x), \sigma) = (\rho, S,
    \heap)}$, \\ $\heap(\rho(x)) \mbox{ is } (v, d)$} & \makecell[t]{$\rho, (\rho', \ell, e,
  \cred{\sigma'} )\!:\!S, \heap, (\CAR~x), \cred{\sigma}$  $
  \rightsquigarrow $ \\ $\rho', S, \heap[\ell := v], e, \cred{\sigma'}$}      &
{\sc car-select} \\
\hline
\makecell[t]{$\cred{GC(\rho_1, S, \heap_1, (\CAR~x), \sigma) = (\rho,
    S, \heap)}$, \\ $\heap(\rho(x))
\mbox{ is } \langle s, \rho'\rangle$ }& \makecell[t]{$\rho, S,
  \heap, (\CAR~x), \cred{\sigma} \rightsquigarrow $\\ $  \rho', (\rho, \rho(x),
  (\CAR~x), \cred{\sigma})\!:\!S, \heap, s, \cred{(\clazy \cup \acar)\sigma }$ }        &
{\sc car-clo}\\
\hline
\makecell[t]{$\cred{GC(\rho_1, S, \heap_1, (\CAR~x), \sigma) = (\rho,
    S, \heap)}$, \\$\heap(\rho(x)) \mbox{ is } (\langle s, \rho'\rangle, d)$} & \makecell[t]{$\rho,\, S,\,  \heap,\,
(\CAR~x), \cred{\sigma} \rightsquigarrow$ \\ $ \rho', \,(\rho, addr(\langle
s, \rho'\rangle), (\CAR~x),\cred{\sigma} )\!:\!S,\, \heap,\, s, \, \cred{\sigma}$ }     &
{\sc car-1-clo} \\
%% \hline

%% $\heap(\rho(x)) \mbox{ is } \langle s, \rho'\rangle$ &\makecell{ $\rho, S,
%%   \heap, (\CAR~x), \cred{\sigma}$  $\rightsquigarrow \rho', (\rho, x,
%%   (\CAR~x), \cred{\sigma})\!:\!S, \heap, s, \cred{(\clazy \cup \acar)\sigma }$}      &
%% {\sc car-clo}
%% \\
\hline
\makecell[t]{$\cred{GC(\rho_1, S, \heap_1, (+~x~y), \sigma) = (\rho, S, \heap)}$,\\
  $\rho(x) \mbox{ is } \bot$  or $\rho(y) \mbox{ is }
  \bot$} & $\rho, S,
  \heap, (+~x~y), \cred{\sigma} \rightsquigarrow \bang$   &
{\sc prim-bang} 
\\
\hline

\makecell[t]{$\cred{GC(\rho_1, S, \heap_1, (+~x~y), \sigma) = (\rho,
    S, \heap)}$, \\ 
$\heap(\rho(x)), \heap(\rho(y)) \in \mathbb{N}$}
 &

\makecell[t]{$\rho, (\rho', \ell, e, \cred{\sigma'})\!:\!S, \heap,
  (+~x~y), \cred{\sigma}$  $\rightsquigarrow$ \\ $\rho', S, \heap[\ell
    := \heap(\rho'(x)) + \heap(\rho'(y))], e, \cred{\sigma'}$}     
 &
{\sc prim-add} \\
\hline
\makecell[t]{$\cred{GC(\rho_1, S, \heap_1, (+~x~y), \sigma) = (\rho, S, \heap)}$,\\$\heap(\rho(x)) \mbox{ is } \langle s, \rho'\rangle$} &\makecell[t]{$\rho, S,
  \heap, (+~x~y), \cred{\sigma}$  $\rightsquigarrow$ \\ $ \rho', (\rho, \rho(x),
  (+~x~y), \cred{\sigma})\!:\!S, \heap, s, \cred{\clazy\sigma}$}      &
{\sc prim-1-clo} \\
\hline
\makecell[t]{$\cred{GC(\rho_1, S, \heap_1, (+~x~y), \sigma) = (\rho, S, \heap)}$,\\$\heap(\rho(y)) \mbox{ is } \langle s, \rho'\rangle $} & \makecell[t]{$\rho,
  S, \heap, (+~x~y), \cred{\sigma}$  $\rightsquigarrow$ \\ $ \rho', (\rho,\rho(y),
  (+~x~y), \cred{\sigma})\!:\!S, \heap, s, \cred{\clazy\sigma}$}      &
{\sc prim-2-clo} \\
\hline
\makecell[t]{$\cred{GC(\rho_1, S, \heap_1, (f~\myvec{x}), \sigma) =
    (\rho, S, \heap)}$,\\ $\mathit{f}~\mbox{defined as}$
$~(\DEFINE~(f~\myvec{y})~e_{\mathit{f}})$}  & \makecell[t]{$\rho, S, \heap,
  (f~\myvec{x}), \cred{\sigma}$  $\rightsquigarrow$\\$ [\myvec{y} \mapsto
    \rho(\myvec{x})], S, \heap, e_{\mathit{f}}, \cred{\sigma}$}      &
{\sc funcall} \\
\hline
\makecell[t]{$\cred{GC(\rho_1, S, \heap_1, (\LET~x\leftarrow
  s~\IN~e), \sigma) = (\rho, S, \heap)}$,\\$\ell$ is a new location}& \makecell[t]{$\rho, S, \heap, (\LET~x\leftarrow
  s~\IN~e), \cred{\sigma}$  $ \rightsquigarrow$ \\ $ \rho\oplus[x
    \mapsto \ell], S, \heap[\ell := \langle s, \lfloor\rho\rfloor_{FV(s)}\rangle], e, \cred{\sigma}$} &
{\sc let} \\ 
\hline
\makecell[t]{$\cred{GC(\rho_1, S, \heap_1, (\pi:\SIF~\psi:x~e_1~e_2),
    \sigma) = (\rho, S, \heap)}$, \\
  $\rho(x) \mbox{ is } \bot$} & $\rho, S,
  \heap, (\pi:\SIF~\psi:x~e_1~e_2), \cred{\sigma} \rightsquigarrow \bang$   &
{\sc if-bang} 
\\
\hline
\makecell[t]{$\cred{GC(\rho_1, S, \heap_1, (\pi:\SIF~\psi:x~e_1~e_2),
    \sigma) = (\rho, S, \heap)}$, \\$\heap(\rho(x)) \ne 0$}  & \makecell{$\rho, S, \heap, (\pi:\SIF~\psi:x~e_1~e_2),
  \cred{\sigma}$  $\rightsquigarrow \rho, S, \heap,  e_1, \cred{\sigma}$} & {\sc if-true} \\
\hline
\makecell[t]{$\cred{GC(\rho_1, S, \heap_1, (\pi:\SIF~\psi:x~e_1~e_2), \sigma) = (\rho, S, \heap)}$,\\$\heap(\rho(x)) = 0$} & \makecell{$\rho, S, \heap, (\pi:\SIF~\psi:x~e_1~e_2),
  \cred{\sigma}$   $\rightsquigarrow
\rho, S, \heap,  e_2, \cred{\sigma}$} & {\sc if-false} \\
\hline
\makecell[t]{$\cred{GC(\rho_1, S, \heap_1, (\pi:\SIF~\psi:x~e_1~e_2), \sigma) = (\rho, S, \heap)}$,\\$\heap(\rho(x)) = \langle s, \rho' \rangle $} & \makecell[t]{$\rho, S, \heap,
  (\pi:\SIF~\psi:x~e_1~e_2), \cred{\sigma}$ $\rightsquigarrow$ \\ $
\rho', (\rho, \rho(x), (\SIF~x~e_1~e_2),  \cred{\sigma})\!:\!S, \heap, s,
\cred{\clazy\sigma}$}
&
{\sc if-clo} \\
\hline
\makecell[t]{$\cred{GC(\rho_1, S, \heap_1, (\SRETURN~x), \sigma) = (\rho, S, \heap)}$,\\
  $\rho(x) \mbox{ is } \bot$} & $\rho, S,
  \heap, (\SRETURN~x), \cred{\sigma} \rightsquigarrow \bang$   &
{\sc return-bang} 
\\
\hline
\makecell[t]{$\cred{GC(\rho_1, S, \heap_1, (\SRETURN~x), \sigma) =
    (\rho, S, \heap)}$,\\ $\heap(\rho(x))~\mbox{is in WHNF with
    value}~v$} & \makecell[t]{$\rho,
  (\rho', \ell, e, \cred{\sigma'})\!:\!S, \heap, (\SRETURN~x), \cred{\sigma}$  $\rightsquigarrow$\\$ \rho', S, \heap[\ell := v], e, 
  \cred{\sigma'}$} &
{\sc return-whnf}\\
\hline
\makecell[t]{$\cred{GC(\rho_1, S, \heap_1, (\SRETURN~x), \sigma) = (\rho, S, \heap)}$,\\$\heap(\rho(x)) = \langle s, \rho' \rangle $} & \makecell[t]{$\rho, S, \heap,
  (\SRETURN~x), \cred{\sigma}$  $
  \rightsquigarrow$ \\
$\rho',~ (\rho, \rho(x), (\SRETURN~x), \cred{\sigma})\!:\!S, \heap,  s,
  \cred{\sigma}$} &
{\sc return-clo} \\
\hline
\end{tabular}

\caption {Minefield semantics. The differences with the small-step
  semantics have been highlighted by shading. \label{fig:minefield-semantics}}
\end{center}

\end{figure*}

\end{document}

%%  LocalWords:  dependences subscripting