### 2016-05-02 paper-for-archive/data-liveness-analysis.tex

\renewcommand{\arraystretch}{.9}
\begin{tabular}{|l||c@{\ }|c@{\ }|c@{\ }|c@{\ }|c@{\ }|c@{\ }|c@{\ }|c@{\ }|c@{\ }|c@{\ }|c@{\ }|c@{\ }|} \hline
 \multicolumn{1}{|r||}{Program} & \rot{\tt fibheap} & \rot{\tt sudoku} & \rot{\tt nperm} & \rot{\tt paraffins} & \rot{\tt lcss} & \rot{\tt huffman} & \rotmore{\tt knightstour} & \rot{\tt nqueens} & \rot{\tt deriv} & \rot{\tt treejoin} & \rot{\tt lambda} & \rot{\tt gc\_bench} \\ \hline
 \#CFG Nonterminals & 621 & 1422 & 662 & 1174 & 642 & 499 & 660 & 404 & 328 & 615 & 669 & 390 \\ \hline
 \#CFG Rules & 1176 & 2009 & 866 & 1773 & 1206 & 818 & 883 & 643 & 468 & 1328 & 1088 & 450 \\ \hline
 \#DFA States & 1761 & 4283 & 1546 & 3346 & 1666 & 1414 & 1519 & 889 & 809 & 1803 & 1703 & 571 \\ \hline
 \#DFA Transitions & 2829 & 7690 & 2522 & 6086 & 2726 & 2528 & 2420 & 1170 & 1435 & 2797 & 2580 & 788 \\ \hline
 DFA Gen Time (sec) & 37.28 & 655.41 & 0.94 & 13.22 & 8.66 & 4.00 & 10.97 & 0.36 & 0.61 & 903.14 & 11.01 & 0.10 \\ \hline
\end{tabular}\\
 (a) Data for Liveness Analysis